---
title: "Curriculum Vitae"
author: "Emilio M. Bruna"
date: "`r format(Sys.time(), '%B %Y')`"

web: www.BrunaLab.org
github: embruna
twitter: BrunaLab
phone: "(352) 514-3935"
email: embruna@ufl.edu

bibliography: ./emb_cv_modular/EMB_publications.bib
csl: brunacv.csl

updated: yes

header-includes:
# - \usepackage{etaremune} # for reverse-numbered list of publications etc
# see https://ctan.math.utah.edu/ctan/tex-archive/macros/latex/contrib/etaremune/etaremune.pdf
- \usepackage[default]{sourcesanspro}
# - \usepackage{enumerate} #for reverse numbering
# - \usepackage{enumitem} #for reverse numbering
- \usepackage{soul} # highlights text
- \usepackage[dvipsnames]{xcolor} #allows access to other colors
- \sethlcolor{Goldenrod} # use goldenrod to match highklighting on preview
- \usepackage{fancyhdr} # modify header into left, center, right
- \pagestyle{fancy} # cues the use of fancy headers
- \fancyfoot{} # cues the use of fancy footer
- \fancyfoot[R]{`r format(Sys.time(), '%B %Y')`} # month & year on footer right
- \fancyhead[R]{Emilio M. Bruna, p. \thepage} # name and page on header right
# - \fancyhead[L]{\hl{Highlighting indicates items in the 2021-2025 PTR period}} # name and page on header left
- \usepackage{setspace} # manipulate line spacing 
- \usepackage{parskip} # manipulate indents and spacing before/after paragraphs 
- \usepackage{xcolor} # changes colors of text
- \AtBeginDocument{\let\maketitle\relax} # this removes default title section
  # to center to top level (#) headings (my name) and change color of font
  # more info # https://www.ctan.org/pkg/sectsty
- \usepackage{sectsty} \sectionfont{\centering\color{darkmidnightblue}}
# centersub-headings (##) & sub-sub-headings (###) & change font color
- \usepackage{sectsty} \subsectionfont{\color{darkmidnightblue}}
- \usepackage{sectsty} \subsubsectionfont{\color{darkmidnightblue}}
# http://latexcolor.com/ to choose colors
# define the colors used above with rgb codes
- \definecolor{darkcerulean}{rgb}{0.03, 0.27, 0.49}
- \definecolor{darkmidnightblue}{rgb}{0.0, 0.2, 0.4}
- \usepackage{xhfill} # modification of line color and width
- \pdfmapfile{=SourceSansPro.map} # this is to deal with error https://tex.stackexchange.com/questions/173204/texlive-font-installation-what-am-i-doing-wrong

geometry: margin=1in
linkcolor: darkmidnightblue
urlcolor: darkmidnightblue
fontawesome: yes
# 
output:
  pdf_document: default
  df_print: paged
  toc: no
  # word_document: default
  # html_document: null
  

---


```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = FALSE,message=FALSE,warning=FALSE)
library(tidyverse)
library(gridExtra)
library(kableExtra)
# library(citr)
library(RefManageR)

```
<!-- heading is line with centered text here: -->
<!-- https://tex.stackexchange.com/questions/15119/draw-horizontal-line-left-and-right-of-some-text-a-single-line -->
<!-- \newcommand*\ruleline[1]{\par\noindent\raisebox{.8ex}{\makebox[\linewidth]{\hrulefill\hspace{1ex}\raisebox{-.8ex}{#1}\hspace{1ex}\hrulefill}}} -->
<!-- \ruleline{\textbf{\textcolor{darkmidnightblue}{\large{RESEARCH \& TEACHING INTERESTS}}}} -->

<!-- this next line suppresses the header (line, name,  page number) on 1st page -->
<!-- \begin{document}. -->
\thispagestyle{empty}

\vspace{-0.3cm}  <!--space between name and address -->
# EMILIO MIGUEL BRUNA
\vspace{-0.3cm}  <!--space between name and address -->
University of Florida\hfill embruna@ufl.edu  
Dept. of Wildlife Ecology & Conservation &\hfill +1 (352) 846-0634  
Center for Latin American Studies\hfill&nbsp;[www.BrunaLab.org](www.brunalab.org)  
Gainesville, FL 32611-0430 USA\hfill&nbsp;[Github.com/BrunaLab](https://github.com/BrunaLab)
\vspace{-0.5cm}  <!--there was too much space, so reduced it with minus: -Xcm -->
<!-- ___   -->
<!-- \par\noindent\rule{\textwidth}{0.4pt} -->
\vspace{0.2cm}  
<!-- \textbf{\textcolor{darkmidnightblue}{RESEARCH \& TEACHING INTERESTS:}} Tropical ecology and conservation, plant population ecology, plant-animal interactions, scientometrics and bibliometrics, science & science policy in Latin America   -->

<!-- text size: https://www.overleaf.com/learn/latex/Font_sizes%2C_families%2C_and_styles -->
<!-- xrfill adjusts line position and width -->
\vspace{0.1cm}
\textbf{\textcolor{darkmidnightblue}{\large{RESEARCH \& TEACHING INTERESTS }}}\xrfill[0.5ex]{1pt}

Tropical ecology and conservation, plant population ecology, demographic modeling, plant-animal interactions, scientometrics and bibliometrics, science & science policy in Latin America

<!-- \textbf{\textcolor{darkmidnightblue}{LANGUAGES:}} Fluent in Spanish, English, & Portuguese   -->
<!-- \vspace{-0.6cm}   -->

<!-- ```{r, child=c('one.Rmd', 'two.Rmd')} -->
<!-- ``` -->

<!-- ```{r, child='one.Rmd'} -->
<!-- ``` -->

<!-- text size: https://www.overleaf.com/learn/latex/Font_sizes%2C_families%2C_and_styles -->
<!-- xrfill adjusts line position and width -->
\vspace{0.2cm}
\textbf{\textcolor{darkmidnightblue}{\large{EDUCATION }}}\xrfill[0.5ex]{1pt}

<!-- ### EDUCATION -->
<!--to indent a block of stuff treat it as a 'group' and then set leftskip -->
<!--be sure to indicate where the group ends with \endgroup -->

\begingroup 
\setlength{\leftskip}{.2cm} 
\setlength{\rightskip}{1cm} 
<!-- ### EDUCATION \hrulefill\ -->
<!-- TEXT \hrulefill\ -->
<!-- Here is some text\xfill{1pt} \par -->
<!-- Here is some text\xfill{8pt} \par -->
<!-- Here is some text\xhrulefill{darkmidnightblue}{2pt}\par -->
<!-- Here is some text\xhrulefill{darkmidnightblue}{1ex}\par -->
<!-- HOW TO MAKE LINE COLOR SAME'? -->
<!-- https://tex.stackexchange.com/questions/32597/vertically-centered-horizontal-rule-filling-the-rest-of-a-line -->
<!-- \textbf{\textcolor{darkmidnightblue}{EDUCATION }}\xrfill[0.5ex]{0.5pt} -->

<!-- ### EDUCATION \hrulefill -->



**Ph.D. in Population Biology**, University of California, Davis  \hfill 2001  
<!-- &nbsp;&nbsp;&nbsp;&nbsp;\hspace{6mm}  _Dissertation Co-Advisors: Susan P. Harrison and Sharon Y. Strauss_   -->
**M.S. in Biology**, University of California, San Diego  \hfill 1995  
<!-- &nbsp;&nbsp;&nbsp;&nbsp;\hspace{6mm} _Thesis Advisor: Ted J. Case_   -->
**B.S. in Ecology, Behavior, & Evolution**, Revelle College,  University of California, San Diego  \hfill 1994  

\endgroup 

<!-- ```{r ed, echo=FALSE,message = FALSE,warning=FALSE} -->

<!-- ed<-read_csv("./cv_files/education.csv") %>% -->
<!-- replace_na(list(program=" ", school=" ",year=" ",degree=" ")) -->
<!-- ed$program<-gsub('","',", ",ed$program) -->
<!-- ed$school<-gsub('","',", ",ed$school) -->
<!-- kbl(ed, -->
<!--     digits = 1, -->
<!--     format = "latex", -->
<!--     align="l", -->
<!--     col.names = NULL, -->
<!--     longtable = T, -->
<!--     row.names = FALSE, -->
<!--     booktabs=T, -->
<!--     linesep = "") %>% -->

<!--   kable_styling(bootstrap_options = c("hover"), -->
<!--                 latex_options = c("scale_down"), -->
<!--                 full_width = F, -->
<!--                 font_size = 10, -->
<!--                 position = "left") %>% -->
<!--   # row_spec(0:0, bold = T) %>% -->
<!--   # # column_spec(2, width = "35em") %>% -->
<!--   sub("\\\\toprule", "", .) %>% -->
<!--   sub("\\\\bottomrule", "", .) -->
<!-- ``` -->
<!-- text size: https://www.overleaf.com/learn/latex/Font_sizes%2C_families%2C_and_styles -->
<!-- xrfill adjusts line position and width -->
\vspace{0.2cm}
\textbf{\textcolor{darkmidnightblue}{\large{PROFESSIONAL APPOINTMENTS }}}\xrfill[0.5ex]{1pt}

<!-- ### PROFESSIONAL APPOINTMENTS -->
\begingroup 
\setlength{\leftskip}{.2cm} 
\setlength{\rightskip}{1cm} 
\textbf{Professor \& Distinguished Teaching Scholar}, University of Florida  
Department of Wildlife Ecology and Conservation \& Center for Latin American Studies \hfill 2013-Present  

\textbf{Secretary}, Ecological Society of America  \hfill 2022-2025  

\textbf{President}, Association for Tropical Biology \& Conservation   \hfill 2020-2021  

\textbf{Director}, Florida-Brazil Linkage Institute, University of Florida  \hfill 2012-2021  

**Editor-in-Chief**, _Biotropica_   \hfill 2013-2019

**Board of Directors**, Dryad Digital Data Repository   \hfill 2014-2020  

**Adjunct Professor**, Universidade Federal de Uberlândia, Brazil   \hfill 2005-2020  

**Associate Professor**, University of Florida  \hfill 2007-2013  

**Assistant Professor**, University of Florida  \hfill 2002-2007  

**NSF Minority Postdoctoral Fellow**, Biological Dynamics of Forest Fragments Project   \hfill 2001-2002  
Instituto Nacional de Pesquisas da Amazônia-Smithsonian Institution  

\endgroup
<!-- ```{r positions, echo=FALSE,message = FALSE,warning=FALSE} -->

<!-- positions<-read_csv("./cv_files/positions.csv") %>% -->
<!-- replace_na(list(dates=" ", position=" ")) -->

<!-- # positions<-as.data.frame(positions) -->
<!-- # positions[1,1] <- cell_spec(positions[1, 1], "latex", bold = T)  -->
<!-- kbl(positions, -->
<!--     digits = 1, -->
<!--     format = "latex", -->
<!--     align="l", -->
<!--     col.names = NULL, -->
<!--     longtable = T, -->
<!--     row.names = FALSE, -->
<!--     booktabs=T, -->
<!--     linesep = "") %>% -->

<!--   kable_styling(bootstrap_options = c("hover"), -->
<!--                 latex_options = c("scale_down"), -->
<!--                 full_width = F, -->
<!--                 font_size = 10, -->
<!--                 position = "left") %>% -->
<!--   # row_spec(0:0, bold = T) %>% -->
<!--   # # column_spec(2, width = "35em") %>% -->
<!--   sub("\\\\toprule", "", .) %>% -->
<!--   sub("\\\\bottomrule", "", .) -->
<!-- ``` -->


<!-- text size: https://www.overleaf.com/learn/latex/Font_sizes%2C_families%2C_and_styles -->
<!-- xrfill adjusts line position and width -->
\vspace{0.2cm}

\textbf{\textcolor{darkmidnightblue}{\large{AWARDS \& FELLOWSHIPS }}}\xrfill[0.5ex]{1pt}
\vspace{-0.4cm}

### AWARDS  
\vspace{-0.2cm}  
\begingroup 
\setlength{\leftskip}{.3cm} 
\setlength{\rightskip}{1cm} 
Fellow of the American Association for the Advancement of Science \hfill 2022   
Alpha Zeta (Faculty Affiliate UF Chapter) \hfill 2020  
Roche Teaching Scholar, UF College of Agricultural & Life Sciences \hfill 2020  
Sigma Xi (Full Member UF Chapter) \hfill 2016  
University of Florida Academy of Distinguished Teaching Scholars \hfill 2014  
Outstanding Paper Award, US Chapter of the International Association for Landscape Ecology \hfill 2012  
Graduate Advisor & Teacher of the Year, UF College of Agricultural & Life Sciences \hfill 2011  
University of Florida Research Foundation Professor \hfill 2009  
Julie S. Denslow Prize for the Outstanding article in _Biotropica_ \hfill 2007  
UF International Educator of the Year (Untenured Faculty category) \hfill 2005  
Diversity Scholar, American Institute for Biological Sciences \hfill 2003  
John L. Harper Early Career Researcher Award, British Ecological Society \hfill 2002  
Ernest C. Mort Leadership Excellence Award, Revelle College, UC San Diego \hfill 1994  

\endgroup

<!-- &nbsp;\quad\quad\quad\quad _for the year's outstanding contribution to the field, published in any journal_   -->
<!-- &nbsp;\quad\quad\quad\quad _for the outstanding paper published in the journal Biotropica_   -->
<!-- &nbsp;\quad\quad\quad\quad _for the outstanding paper by a young investigator in the Journal of Ecology_ -->

<!-- \endgroup -->
\vspace{-0.2cm}
<!-- \begingroup  -->
<!-- \setlength{\leftskip}{.2cm}   -->
### FELLOWSHIPS
\vspace{-0.2cm}
\begingroup 
\setlength{\leftskip}{.3cm} 
\setlength{\rightskip}{1cm} 
Edward P. Bass Distinguished Visiting Environmental Scholar, Yale University \hfill 2024  
*Ciência Sem Fronteiras* Special Visiting Researcher Fellowship, CNPq Brazil \hfill 2012  
Fulbright Scholar, Brazil \hfill 2009  
Visiting Researcher Fellowship, CNPq Brazil \hfill 2009  
NSF Minority Postdoctoral Research Fellowship \hfill 2001  
NSF International Research Fellowship *(declined)* \hfill 2001  
Ford Foundation Dissertation Year Fellowship \hfill 2000  
UC Davis Graduate Studies Fellowship \hfill 1999  
UC Davis John F. Steindler Graduate Fellowship \hfill 1998  
NSF Minority Predoctoral Research Fellowship \hfill 1995  
Howard Hughes Senior Honors Thesis Fellowship \hfill 1993  
UCSD Howard Hughes Summer Research Fellowship \hfill 1993  

\endgroup

<!-- text size: https://www.overleaf.com/learn/latex/Font_sizes%2C_families%2C_and_styles -->
<!-- xrfill adjusts line position and width -->
\vspace{0.2cm}
\textbf{\textcolor{darkmidnightblue}{\large{GRANTS IN SUPPORT OF RESEARCH, TEACHING, \& TRAINING }}}\xrfill[0.5ex]{1pt}
<!-- ## GRANTS IN SUPPORT OF RESEARCH, TRAINING, & PROGRAM DEVELOPMENT (>$5.5 million)   -->

\begingroup 
\setlength{\leftskip}{.2cm}  

1. National Science Foundation. _"OPUS: Is habitat fragmentation detrimental or beneficial? Promoting our understanding of a contentious ecological question via synthesis."_ ($245,385, PI: EM Bruna)  

1. UF Faculty Enhancement Opportunity Program _"Application of artificial intelligence for the study of tropical plant demography"_ ($31,159.03, Spring 2025).

1. Maxwell-Hanrahan Foundation. _"Support for the ATBC Field Grant Program"_ (PIs EM Bruna, P Sampaio, L Lohmann: $10,000 for 2022-2023, renewed 2023-2026 for $45,000)

1. Maxwell-Hanrahan Foundation. _"Support for the UF TCD Seed Grant Program"_ ($10,000; PIs EM Bruna, B Loiselle: $10,000 for 2022-2023, renewed 2023-2026 for $45,000)

1. Sloan Foundation. _"Reinventing Field Courses"_ ($74,060; PI: EM Bruna, Co-PIs: R McCleery, P Antonenko, N Attias, D Barton, 2020-2021)

2. National Science Foundation. _"SG: Synergistic effects of forest fragmentation and droughts on tropical plant population dynamics"_ ($199,904; PI: EM Bruna, Co-PIs: M. Uriarte, 2020-2022).

3. National Science Foundation. _"REU Supplement for SG: Synergistic effects of forest fragmentation and droughts on tropical plant population dynamics"_ ($14,400; PI: EM Bruna, Co-PIs: M Uriarte, 2021).

3. National Science Foundation. _"The role of the matrix in fragmented landscapes: experimental tests and novel theory"_ ($548,058; PI: R Fletcher, Co-PIs: EM Bruna, RD Holt, and B Reichert, 2018-2023)  

4. UF Informatics Institute SEED Funds: _"Cross-National Patterns of Productivity & Collaboration by Latin American Scientists"_ ($20,000, PI: EM Bruna, Co-PIs: S Perz, C McCarty).

5. Mindlin Foundation. _"Who are the Gatekeepers of Discovery? A crowdsourced project to ID & map the editors of academic journals"_ ($3633, 2017-2018).

5. National Geographic Global Exploration Fund. _"Experimental climate change in the Amazon: How will rising CO2 influence plants, insects, and their interactions?"_ ($19,970, PI: EM Bruna, Co-PIs H Vasconcelos, F Mundim, D Lapola).

5. John Wiley and Sons. _"Funds to support the Biotropica Editorial Office"_ ($199,945; 2014-2019).

5. UF Center for Latin American Studies Faculty Travel Award ($1500, 2014)

5. UF Center for Latin American Studies Speakers Program. _"Herton Escobar: Science Correspondent, Estado de São Paulo Newspaper"_ ($2000, including matching funds provided by the College of Journalism and Communications, the Tropical Conservation & Development Program, and Office of the VP for Research, 2014).

5. UF IFAS Dean For Research, UF VP for Research, and FAPESP-Brazil. _"Support for 4 UF-FAPESP Collaborative Research Grants"_ ($160,000, 2013-2015).

5. UF Center for Latin American Studies Faculty Research Award: _"Cross-national patterns of productivity and collaboration by Latin American ecologists"_ ($5450, 2012-2013).

5. National Science Foundation: _"Doctoral Dissertation Enhancement Proposal: International: Carbon sequestration and tree biodiversity in Amazonian floodplain forests."_  ($11,000; OISE-0929183 to E. Bruna prepared by and for doctoral student Christine Lucas, 2010-2012).

5. David and Lucile Packard Foundation. _"Nitrogen dynamics in Brazil's Cerrado"_ ($142,198, Grant 2008-32691 to E. Bruna and co-PIs M Mack and H Vasconcelos, 2009-2011).

5. National Science Foundation. _"IGERT in Spatial Ecology & Evolution: Quantitative Training in Biology, Statistics, & Mathematics."_ ($2,958,113; PI: B Bolker; Co-PIs: E Bruna, C Osenberg, M Martcheva, & M Christman, 2008-2013).

5. UF International Center. _"International Speakers Program: Dr. Claudio Padua, 2008 UF Distinguished Alumnus"_ ($1500, with $3500 in matching funds from Wildlife Ecology and Conservation, Latin American Studies, the Tropical Conservation & Development Program, and Institute for Agricultural & Life Sciences).

5. US Department of Education Fund for the Improvement of Postsecondary Education (FIPSE). _"Energy, environment, and sustainable development: the central role of the United States and Brazil"_ ($202,329; PI: J. Kahn, Washington & Lee University; E. Bruna is UF Project Director; 2006-2010)

5. National Science Foundation Collaborative Grant. _"Mechanisms influencing seedling recruitment and establishment in a fragmented Amazonian landscape"_ DEB-0614149 to E. Bruna ($230,888) and DEB-0614339 to M Uriarte, Columbia Univ. ($251,831; 2006-10).

5. National Science Foundation Minority Career Advancement Award. _"Linking population and ecosystem ecology: how does nitrogen deposition influence plant demography in tropical savannas?"_ DEB-0542287 ($174,499; 2006-08).

5. Joseph Jones Ecological Research Center. _"Productivity and Biodiversity Patterns: Control in a SE Coastal Plain Landscape (Support for the graduate studies of Gwen Iacona)."_ ($19,860; 2005-2007).

5. National Science Foundation Collaborative Grant. _"Elucidating mechanisms of coexistence in a community of ant-plant mutualists"_.  DEB-0453631 to E Bruna ($245,096) & DEB-0452720 to B Inouye, FSU ($260,000; 2005-09).

5. National Science Foundation Grant OISE-0437369.  _"U.S.-Brazil Collaborative Research: Seed predation and herbivory in Neotropical savannas: are there demographic consequences for plants?"_ ($25,012; 2005-2008).

5.  Conselho Nacional de Desenvolvimento Científico e Tecnológico, Brazil. _"Top-down effects on the structure of Cerrado plant communities"_ (R$ 22,875.74, Co-PI, 2005-2007)

5. Fundação de Amparo a Pesquisa do Estado de Minas Gerais, Brazil. _"Diversity and conservation of invertebrates in the "Triângulo Mineiro": a case study using ants as bioindicators"_ (R$12,924.13; Senior Collaborator, 2005-2007)

5. Fundação de Amparo a Pesquisa do Estado de Minas Gerais, Brazil (Fapemig). _"Effect of herbivory by leaf-cutter ants on Cerrado vegetation"_ (R$25,229.74, Co-PI, 2004-2006)

5. School for Natural Resources and the Environment New Faculty Support Program. _"Seed predation and herbivory in Neotropical savannas: are there demographic consequences for plants?"_ ($5,000; 2004)

5. Alfred P. Sloan Foundation.  _"Increasing the number of minority PhDs studying Ecology at UF."_ ($32,440 and four-year Teaching Assistantships from the College of Agricultural and Life Sciences; 2004).

5. National Science Foundation Research Starter Grant DEB-0309819.  _"The effect of forest fragmentation on ant-plant mutualisms."_ $50,585; 2003-2005).

5. USAID Subcontract from IIEB.  _"Sustainable Communities and Landscapes: A proposed 	program to sustain natural ecosystems and enhance local livelihoods in Brazil's Amazon and Atlantic Forest regions"_ ($199,145; 10/01/03-9/30/07; PI: Marianne Schmink; EB and 5 other Co-PIs).

5. Instituto Internacional de Educação do Brasil-SUNY International Development Group.  _"Training of young scientists in Amazonia: A new internship program at the Biological Dynamics of Forest Fragments Project."_ (R$21,000; co-PI Patricia Delamônica, 2001).

5. National Science Foundation Minority Postdoctoral Research Fellowship DBI-0109226. _"The effect of forest fragmentation on ant-plant mutualisms"_ ($190,200; 2001).

5. NSF Dissertation Improvement Grant INT 98-06351. _"Effect of habitat fragmentation on the growth, reproduction, and population dynamics of an Amazonian understory herb"_ ($17,600; 1998-2000).

5. Grants for Graduate Research: Smithsonian Institution, Manomet Observatory Kathleen S. Anderson Award, Wilson Ornithological Society Paul A. Stewart Award, and UC Davis (M.E. Mathias Award, Graduate Studies Research Grant, Jastro-Shields Research Award, and Center for Population Biology Summer Research Grant (Total: $11860; 1996-1997).

5. Grants for Undergraduate Research: Explorer's Club Grant, UCSD Undergraduate Scholastic Research Grant (Total: $1955; 1994-1995).

\endgroup

<!-- ### GRANT PROPOSALS UNDER REVIEW or IN PREPARATION -->

<!-- 1. UF 10x10x10 Initiative: The "interconnected future of Florida & Latin America: Entrepreneurship, the Environment, & Transnational Migration" (EM Bruna and R. Stepp, UF Center for Latin American Studies). -->

<!-- 1. National Science Foundation. "The genomic basis and demographic consequences of male-biased local adaptation in a moss" ($2,267,331; PI: S. McDaniel, Co-PIs: K. Krippen, EM Bruna, G. John, B. P. Durham).  -->


<!-- --- -->
<!-- nocite: '@*' -->
<!-- ... -->

<!-- text size: https://www.overleaf.com/learn/latex/Font_sizes%2C_families%2C_and_styles -->
<!-- xrfill adjusts line position and width -->
\vspace{0.2cm}
\textbf{\textcolor{darkmidnightblue}{\large{PUBLICATIONS \& SCHOLARLY COMMUNICATION  }}}\xrfill[0.5ex]{1pt}
<!-- ## PUBLICATIONS & SCHOLARLY COMMUNICATION -->
\vspace{-0.4cm}

### PREPRINTS AND MANUSCRIPTS IN REVIEW or IN REVISION 

<!-- \begin{enumerate}[{start=7,label=[\arabic*]\addtocounter{enumi}{-2}}] -->

<!-- \item  -->

<!-- Tropical Fragmentation Paper Victor Arroyo-Rodriguez et al -->
1. Manning, M. G. Roberts,P. Hahn, J. Kahler, E. M. Bruna, M. Acevedo.  Sequencing foundational computational instruction in the life sciences.\textit{In review}, \textit{\textbf{Ecology and Evolution}

1. Ykeuti Meiga, AY,  ALJ Desbiez, RAP Dornas, A Giroux, EM Bruna, A Bertassoni, N Attias, V Hull, G Soresini, MH Alves, D Valle. Beyond deforestation: the impact of agricultural intensification on wildlife.  \textit{In review}, \textit{\textbf{Journal for Nature Conservation}}.

1. Scott, ER, M Uriartem EM Bruna. Context-dependent consequences of including delayed effects of climate in Integral Projection Models.  \textit{In revision}, \textit{\textbf{Journal of Ecology}}.

1. Bruna, EM Declining contribution of the United States federal government to global research corpus. \textit{Preprint:} https://doi.org/10.31222/osf.io/4hfe9_v1

1. Galmán, A.  \textit{et. al.} Beyond the mean: how plant growth-defense strategies influence intrapopulation variation in herbivory. \textit{In review}, \textit{\textbf{Ecology Letters}}.

1. Mundim, FM., EVieira-Neto, GF Schneider, H Alborn, EM Bruna. Simultaneous and sequential herbivory dynamics: exploring plant defenses strategies amidst drought challenges. \textit{In review}, \textit{\textbf{Frontiers in Plant Science}}. 

1. Resende Rodrigues, D, YR Bovolenta, EM Bruna, JA Pimenta, E Bianchini. Coexistence of understory tree species via architecture differentiation in the Atlantic Forest of southern Brazil. \textit{In review}, \textit{\textbf{Tropical Ecology}}.


<!-- \end{enumerate}   -->

<!-- very helpful for bibliography: https://inbo.github.io/tutorials/tutorials/r_citations_markdown/ -->
<!-- maybe to print out by reference groups? -->
<!-- https://forums.zotero.org/discussion/23604/a-mini-guide-to-sort-bibliography-by-reference-type -->
<!-- \endgroup -->

<!-- TODO: -->
<!-- 1) indent the automated refs -->
<!-- 2) put field in zotero for noes (eg theis chapter, recipient of..., etc), -->
<!-- 3) make the extra field with notes bold for cv -->
<!-- 3) find out which refs ones are missing -->



<!-- ### PEER-REVIEWED PUBLICATIONS (AUTOMATED) -->

<!-- <div id="refs"></div> # this allows you to put text after -->
<!-- --- -->
<!-- nocite: '@*' # this puts eveything in .bib without needing to ccite it -->
<!-- --- -->

 
### PEER-REVIEWED PUBLICATIONS \& BOOK CHAPTERS}

**_Note regarding coauthorship on publications by lab members:_** _I am not a coauthor on papers by lab members, including publications that result from side projects or thesis chapters, as I consider mentorship regarding study development and design, analyses, and manuscript preparation fundamental to my role as their advisor. A list of the publications by lab members on which I am not a coauthor can be found below._ 

1. Radhamoni, H, M Spicer; M Duguid, L Comita EM Bruna, A Andrade, J Beck, J Burton, H Chuquillanqui, D Cicuzza, M Dovciak, F Gilliam, K Greig, C Hale, C Hulshof, N Izlar, S Johnson, J LaManna, D Matlaga, D Melly, A Menzmer, K Miller, G Moulatlet, J Myers, J Sah, M Vaz, A Wassel, S Wessel, J Yin, A Zuluaga, S Queenborough. \textit{In press}. Herbaceous plants in forest ecosystems: One hundred fundamental and applied questions. \textit{\textbf{Journal of Vegetation Science}}.  

1. R J Fletcher, Jr, TAH Smith, M Jones, F Schellenberg, R Payne, N Kortessis, EM Bruna, and RD Holt. 2026. Landscape quality drives ecological change in habitat fragmentation experiments. \textit{\textbf{Nature Ecology \& Evolution}}.  

1. DS Mason, KM Robertson, PG Hahn, EM Bruna, JL  Willis, and MA Lashley. 2026. Fire promotes passive directed seed dispersal by birds. \textit{\textbf{Current Biology}} 36(1):28–37. https://doi.org/10.1016/j.cub.2025.11.014 **_Data:_** [Dryad](https://doi.org/10.5061/dryad.tqjq2bw77)

1. Smith, TAH, RD Holt, EM Bruna, \& RJ Fletcher Jr. 2025. Isolating the role of the matrix at patch and landscape scales. \textit{\textbf{Journal of Animal Ecology}} 94:1800–1810. https://doi.org/10.1111/1365-2656.70089 **_Data:_** [Dryad](https://doi.org/10.5061/dryad.kprr4xhh6)

1. Altomare de Paula, M., EM Bruna, KC Ferreira Neves, ER Borges, I Oliveras Menor, JA do Prado Júnior, HL Vasconcelos. \textit{In press.} The phylogenetic diversity of plant communities in response to anthropogenic disturbances and non-native species in a Neotropical savanna. \textit{\textbf{Annals of Botany}}.  

1. JL Hyde, C Swanson, SA Bohlman, S Athayde, EM Bruna, and DR Valle 2025. A lack of open data standards for large infrastructure project hampers social-ecological research in the Brazilian Amazon. \textit{\textbf{PeerJ}} 13:e19926 https://doi.org/10.7717/peerj.19926

1. Bruna, EM 2025. Is there really such a thing as _Tropical_ Biology? \textit{\textbf{Biotropica}} 57:e13387. [doi.org/10.1111/btp.13387](https://doi.org/10.1111/btp.13387)  

1. Powers, JS, EM Bruna, and LG Lohmann. 2025. Looking Back and Moving Forward: Reflections on 60 Years of the Association for Tropical Biology and Conservation. \textit{\textbf{Biotropica}} 57(5): e70074. https://doi.org/10.1111/btp.70074.

1. Carrasco‐Rueda, F, CL Cardelús, Y Chirango, CJ Garnica‐Díaz, Y Godwill, V Gowda, B Maas, F Montaño‐Centellas, R Ostertag, A Roopsind, Sheherazade, G Vargas G, Y Arn Teh, EM Bruna, NJ Cordeiro, GGO Dossa, E Effiom, LG Lohmann, M Mujawamariya, I Oliveras Menor, P Parolin, JM Posada, RA Zahawi, LS Santiago. 2025. ATBC Code of Conduct. \textit{\textbf{Biotropica}} 57(5): e70090. https://doi.org/10.1111/btp.70090.

1. Brodie, JF, F Carrasco, S Chatterjee, Y Chirango, CJ Garnica‐Díaz, V Gowda, B Maas, F Montaño‐Centellas, R Ostertag, P Hariharan, G Vargas, EM Bruna, CL Cardelus, NJ Cordeiro, GGO Dossa, E Effiom, LG Lohmann, IO Menor, M Mujawamariya, P Parolin, JM Posada, M Soto Rueda, Y Arn Teh, B Yadok, RA Zahawi, LS Santiago 2025. ATBC Code of Ethics. \textit{\textbf{Biotropica}} 57(5): e70091. https://doi.org/10.1111/btp.70091.  

1. RJ Fletcher Jr, TAH. Smith, S Troy, N Kortessis, EC Turner, EM Bruna, and RD Holt. 2024. Matrix effects in ecology, evolution, and conservation. \textit{\textbf{Annual Review of Ecology, Evolution, and Systematics}} 55:423-447 [doi.org/10.1146/annurev-ecolsys-102722-025653](https://doi.org/10.1146/annurev-ecolsys-102722-025653)

1. Ferreira, AM, AP Faria, HL Vasconcelos, EM Bruna, VE Costa, AS Franco Pinheiro Moreira. 2024. Does anthropogenic nitrogen deposition have effects on grass physiology of natural areas? \textit{\textbf{South African Journal of Botany}} 165: 339-348. [doi.org/10.1016/j.sajb.2023.12.044](https://doi.org/10.1016/j.sajb.2023.12.044)

1. The Herbivory Variability Network (Robinsion, ML _et al._). 2023. Plant size, latitude, and phylogeny explain within-population variability in herbivory. _**Science**_ 382:679-683. [DOI: 10.1126/science.adh8830](https://www.science.org/doi/full/10.1126/science.adh8830)

1. Bruna, EM, M Uriarte, M Rosa Darrigo, P Rubim, CF Jurinitz, ER Scott, O Ferreira da Silva, \& W John Kress. 2023. Demography of the understory herb \textit{Heliconia acuminata} (Heliconiaceae) in an experimentally fragmented tropical landscape. \textit{\textbf{Ecology}} 104(12): https://e4174 doi.org/10.1002/ecy.4174 

1. Bruna, EM 2023. Fundamental errors of data collection \& validation undermine claims of 'Ideological Intensification' in STEM. \textit{\textbf{Bioscience}} 73(11): 775-777. [doi.org/10.1093/biosci/biad047](https://doi.org/10.1093/biosci/biad047)  **_Code:_** [Zenodo](https://zenodo.org/records/7896249)

1. Fletcher, RJ, TAH Smith, N Kortessis, EM Bruna, and RD. Holt. 2023. Landscape experiments unlock the multi-scale relationships among habitat loss, fragmentation, and patch-size effects. \textit{\textbf{Ecology}} 104(5): e4037. [doi.org/10.1002/ecy.4037](https://doi.org/10.1002/ecy.4037)

1. Halpern, BS, C Boettiger, MC Dietze, *et al.* 2023. Priorities for synthesis research in ecology and environmental Science. \textit{\textbf{Ecosphere}} 14(1): e4342. [doi.org/10.1002/ecs2.4342](https://doi.org/10.1002/ecs2.4342)

1. 	Scott, ER, M Uriarte, \& EM Bruna. 2022. Delayed effects of climate on vital rates lead to demographic divergence in Amazonian forest fragments. \textit{\textbf{Global Change Biology}} 28(2):463-479. [doi.org/10.1111/gcb.15900](https://doi.org/10.1111/gcb.15900) \textit{\textbf{Recipient of the Postdoctoral Excellence Award from the Ecological Society of America's Plant Population Ecology Section}}

1. Audrey C. Smith, Leandra Merz, Jesse B. Borden, Chris K. Gulick, Akhil R. Kshirsagar, Emilio M. Bruna. 2022. Assessing the effect of article processing charges on the geographic diversity of authors using Elsevier's 'Mirror Journal' system. \textit{\textbf{Quantitative Science Studies}} 2(4):1123-1143. [doi.org/10.1162/qss_a_00157](https://doi.org/10.1162/qss_a_00157) **_Data & Code:_** [Zenodo](https://doi.org/10.5281/zenodo.5500293) **_Featured in [Nature](https://www.nature.com/articles/d41586-022-00342-w), [NRC News](https://www.nrc.nl/nieuws/2022/02/22/systeem-van-open-access-helpt-wetenschappers-uit-arme-landen-niet-a4092976)_**. 

1. Russell, AE, TM Aide, E Braker, EM Bruna, CN Ganong, RD Hardin, KD Holl, SC Hotchkiss, J Klemens, EK Kuprewicz, SK Macey, D McClearn, G Middendorf, R Ostertag, JS Powers, SE Russo, JL Stynoski, U Valdez, CG Wills. 2022. Integrating tropical research into biology education is urgently needed. \textit{\textbf{PLoS Biology}} 20(6): e3001674. [doi.org/10.1371/journal.pbio.3001674](https://doi.org/10.1371/journal.pbio.3001674) 

1. Mundim, FM. EHM. Vieira-Neto, H Alborn, \& EM Bruna. 2021. Disentangling the influence of resource variability and simultaneous attacks on whole-plant tolerance to herbivory. \textit{\textbf{Journal of Ecology}} 109:2729- 2739. [PhD Thesis Chapter]. [doi.org/10.1111/1365-2745.13684](https://doi.org/10.1111/1365-2745.13684) _**Selected as a 2021 UF-IFAS High Impact Publication**_.  

1. Bruna, EM, R Chazdon, TM Errington and BA Nosek. 2021. A proposal to advance theory and promote collaboration in tropical biology by supporting replications. \textit{\textbf{Biotropica}} 53:6-10. [doi.org/10.1111/btp.12912](https://doi.org/10.1111/btp.12912)

1. Barbosa Silva, LV, HL Vasconcelos, MC Mack, AdS Ferreira, and EM Bruna. 2020. Effects of experimental nitrogen enrichment on soil properties and litter decomposition in a Neotropical savanna. **_Austral Ecology_** 45:1093-1102. [doi.org/10.1111/aec.12936](https://doi.org/10.1111/aec.12936) [Ph. D. Thesis Chapter].  
1. Fournier, A, ME Boone, FR Stevens, and EM Bruna. 2020. refsplitr: Author name disambiguation, author georeferencing, and mapping of coauthorship networks with Web of Science data. **_Journal of Open Source Software_** 5(45):2028. [doi.org/10.21105/joss.02028](https://doi.org/10.21105/joss.02028)

1. Beckman, NG, CE Aslan, HS Rogers, O Kogan, JL Bronstein, JM Bullock, F Hartig, J HilleRisLambers, Y Zhou, D Zurell, JF Brodie, EM Bruna, RS Cantrell, RR Decker, E Efiom, EC Fricke, K Gurski, A Hastings, JS Johnson, BA Loiselle, MN Miriti, MG Neubert, L Pejchar, JR Poulsen, G Pufal, OH Razafindratsima, ME Sandor, K Shea, S Schreiber, EW Schupp, RS Snell, C Strickland, and J Zambrano. 2020. Advancing an interdisciplinary framework to study seed dispersal ecology. **_AoB Plants_** 12:plz048.  

3. Costa, AN, A Bartimachi, HL Vasconcelos, EM Bruna, and EHM Vieira-Neto. 2020. Annual litter production in a Brazilian Cerrado woodland savanna. **_Southern Forests_** 82:65-69.  

4. Powers, JS, EM Bruna, A Campos-Arceiz, P Delamonica Sampaio, SJ DeWalt, LG Lohmann, and RA Zahawi. 2020. Tropical biology and conservation in the time of the covid-19 pandemic. **_Biotropica_** 52:399-399.  

5. Aslan, C, NG Beckman, HS Rogers, J Bronstein, D Zurell, F Hartig, K Shea, L Pejchar, M Neubert, J Poulsen, J HilleRisLambers, M Miriti, B Loiselle, E Effiom, J Zambrano, G Schupp, G Pufal, J Johnson, JM Bullock, J Brodie, E Bruna, RS Cantrell, R Decker, E Fricke, K Gurski, A Hastings, O Kogan, O Razafindratsima, M Sandor, S Schreiber, R Snell, C Strickland, and Y Zhou. 2019. Employing plant functional groups to advance seed dispersal ecology and conservation. **_AoB Plants_** 11:plz006.  

7. Brooks, ME, K Kristensen, MR Darrigo, P Rubim, M Uriarte, E Bruna, and BM Bolker. 2019. Statistical modeling of patterns in annual reproductive rates. **_Ecology_** 100 (7):e02706 [Ph.D. Thesis Chapter].  

8. Costa, AN, HL Vasconcelos, EHM Vieira-Neto, and EM Bruna. 2019. Adaptive foraging of leaf-cutter ants to spatiotemporal changes in resource availability in Neotropical savannas. **_Ecological Entomology_** 44:227-238.  

9. Fletcher, RJ, Jr, JA Sefair, C Wang, CL Poli, TAH Smith, EM Bruna, RD Holt, M Barfield, AJ Marx, and MA Acevedo. 2019. Towards a unified framework for connectivity that disentangles movement and mortality in space and time. **_Ecology Letters_** 22:1680-1689. 

1.	Bruna, EM. 2019. _Biotropica_: the next 50 years. \textit{\textbf{Biotropica}} 51(4): 461–62.

10. Bruna, EM. 2018. Editorial board members are a non-random sample of ecological experts. **_Nature Ecology and Evolution_** 2(2):202.  

10. Costa, AN, EM Bruna, and HL Vasconcelos. 2018. Do an ecosystem engineer and environmental gradient act independently or in concert to shape juvenile plant communities? Tests with the leaf-cutter ant _Atta laevigata_ in a Neotropical savanna. **_PeerJ_** 6:e5612. **_Data:_** [Dryad](https://doi:10.5061/dryad.vf1gq0q)  **_Code:_** [Zenodo](http://doi.org/10.5281/zenodo.1418482)  [Ph. D. Thesis Chapter].  

11. Brito Darosci, AA, EM Bruna, JC Motta-Junior, CdS Ferreira, JG Blake, and CB Rodrigues Munhoz. 2017. Seasonality, diaspore traits and the structure of plant-frugivore networks in Neotropical savanna forest. **_Acta Oecologica_** 84:15-22. [Ph. D. Thesis Chapter].  

12. Brudvig, LA, SJ Leroux, CH Albert, EM Bruna, KF Davies, RM Ewers, DJ Levey, R Pardini, and J Resasco. 2017. Evaluating conceptual models of landscape change. **_Ecography_** 40:74-84.  

13. Costa, AN, HL Vasconcelos, and EM Bruna. 2017. Biotic drivers of seedling establishment in Neotropical savannas: Selective granivory and seedling herbivory by leaf-cutter ants as an ecological filter. **_Journal of Ecology_** 105:132-141. **_Data:_** [Dryad](https://doi.org/10.5061/dryad.1hj56) [Ph. D. Thesis Chapter]. 

14. Espin, J, S Palmas, F Carrasco-Rueda, K Riemer, PE Allen, N Berkebile, KA Hecht, K Kastner-Wilcox, MM Nunez-Regueiro, C Prince, C Rios, E Ross, B Sangha, T Tyler, J Ungvari-Martin, M Villegas, TT Cataldo, and EM Bruna. 2017. A persistent lack of international representation on editorial boards in environmental biology. **_PLoS Biology_** 15:e2002760. **_Code:_** [Zenodo](http://doi.org/10.5281/zenodo.1067800) **_Data:_** [Dryad](https://doi.org/10.5061/dryad.mh189)  

15. Mundim, FM, HT Alborn, EHM Vieira-Neto, and EM Bruna. 2017. A whole-plant perspective reveals unexpected impacts of above- and belowground herbivores on plant growth and defense. **_Ecology_** 98:70-78.  

16. Resasco, J, EM Bruna, NM Haddad, C Banks-Leite, and CR Margules. 2017. The contribution of theory and experiments to conservation in fragmented landscapes. **_Ecography_** 40:109-118.  

17. Rocha-Ortega, M, A Bartimachi, J Neves, EM Bruna, and HL Vasconcelos. 2017. Seed removal patterns of pioneer trees in an agricultural landscape. **_Plant Ecology_** 218:737-748.   

16.	Pitman, N. C. A., J. Widmer, EM Bruna, C. N. Jenkins, G. Stocks, L. Seales, F. Paniagua, and E. Maehr. 2017. El sorprendente liderazgo del Ecuador en la producción científica de la región andino-amazónica, pp. 27-30 in D. Romo et al. (eds.), Los Secretos del Yasuní: Avances en investigación en la Estación de Biodiversidad Tiputini, Universidad San Francisco de Quito, Quito.  

18. Carvalho, KS, MAF Carneiro, IC Nascimento, AK Saha, and EM Bruna. 2016. Lower ant diversity on earth mounds in a semi-arid brazilian ecosystem: Natural or a sign of degradation? **_Sociobiology_** 63(4):1022-1030.    

19. Mundim, FM, and EM Bruna. 2016. Is there a temperate bias in our understanding of how climate change will alter plant- herbivore interactions? A meta- analysis of experimental studies. **_American Naturalist_** 188:S74-S89. [Ph. D. Thesis Chapter] **_Data:_** [Dryad](https://doi.org/10.5061/dryad.dn048) _**Selected as WEC Department's 2016 High Impact Publication**_. 

20. Sobrinho, MS, M Tabarelli, IC Machado, JC Sfair, EM Bruna, and AV Lopes. 2016. Land use, fallow period and the recovery of a caatinga forest. **_Biotropica_** 48:586-597.    

21. Vieira-Neto, EHM, HL Vasconcelos, and EM Bruna. 2016. Roads increase population growth rates of a native leaf-cutter ant in Neotropical savannahs. **_Journal of Applied Ecology_** 53:983-992. **_Data:_** [Dryad](https://doi.org/10.5061/dryad.1s4f2) [Ph. D. Thesis Chapter].  

7.	Parker, TH, Nakagawa, S, Gurevitch, J and IIEE (Improving Inference in Evolutionary Biology and Ecology) workshop participants. 2016. Promoting transparency in evolutionary biology and ecology. **_Ecology Letters_**, 19: 726-728.   

9.	Whitlock, MC, JL Bronstein, EM Bruna, AM Ellison, CW Fox, MA McPeek, AJ Moore, MAF Noor, MD Rausher, LH Rieseberg, M G Ritchie, and RG Shaw. 2016. A balanced data archiving policy for long-term studies. **_Trends in Ecology and Evolution_** 31(2) 84-85 [Response to article].  

10. Fine, PVA and Bruna, EM. 2016. Neotropical White-sand Forests: Origins, Ecology and Conservation of a Unique Rain Forest Environment. **_Biotropica_**, 48: 5-6. https://doi.org/10.1111/btp.12305

21.	Hoeksema, JD and EM Bruna. 2015. Context-dependent outcomes of mutualistic interactions. Pages 181-202 in Mutualism, JL Bronstein (ed.). Oxford University Press. 

22. Bruna, EM, TJ Izzo, BD Inouye, and HL Vasconcelos. 2014. Effect of mutualist partner identity on plant demography. **_Ecology_** 95:3237-3243. **_Cover Article_**.  _Code_ https://doi.org/10.5281/zenodo.11650 **_Data:_** [Dryad](https://doi.org/10.5061/dryad.5kc45)  

23. Cho, AH, SA Johnson, CE Schuman, JM Adler, O Gonzalez, SJ Graves, JR Huebner, DB Marchant, SW Rifai, I Skinner, and EM Bruna. 2014. Women are underrepresented on the editorial boards of journals in environmental biology and natural resource management. **_PeerJ_** 2:e542. **_Named to the 2015 _PeerJ_ Picks Collection (Highlighting 20 articles of the 471 published in 2014)_**.  **_Data:_** [Dryad](https://doi.org/10.5061/dryad.6jn86) _Code_: [Github](https://github.com/embruna/Editorial-Board-Gender)

24. Smith, MJ, C Weinberger, EM Bruna, and S Allesina. 2014. The scientific impact of nations: Journal placement and citation performance. **_PLoS ONE_** 9:e109195.  

25. Braunger Vasconcelos, P, GM Araújo, and EM Bruna. 2014. The role of roadsides in conserving cerrado plant diversity. **_Biodiversity and Conservation_** 23:3035-3050.  

15.	Bruna, EM 2014 On identifying rising stars in biology. **_Bioscience_** 64(3):169.  

26. Girão, LC, AV Lopes, M Tabarelli, and EM Bruna. 2014. Mudanças em atributos reprodutivos de árvores reduzem a diversidade funcional em uma paisagem fragmentada de floresta Atlântica pp. 145-174 in Serra Grande: Uma Floresta de ideas.  M. Tabarelli, A. V. Neto, I. R. Leal, and A. V. Lopes (Eds.).  Editora Universitároa da UFPE. Recife. [translation of Girão _et al_. 2007, _PLoS ONE_].  

26. Araújo, GM, AF Amaral, EM Bruna, and HL Vasconcelos. 2013. Fire drives the reproductive responses of herbaceous plants in a Neotropical swamp. **_Plant Ecology_** 214:1479-1484.  

27. Côrtes, MC, M Uriarte, MR Lemes, R Gribel, W John Kress, PE Smouse, and EM Bruna. 2013. Low plant density enhances gene dispersal in the Amazonian understory herb _Heliconia acuminata_. **_Molecular Ecology_** 22:5716-5729. [Ph. D. Thesis Chapter].  **_Data:_** [Dryad](https://doi.org/10.5061/dryad.b003f)

28. Jurinitz, CF, AA de Oliveira, and EM Bruna. 2013. Abiotic and biotic influences on early-stage survival in two shade-tolerant tree species in Brazil's Atlantic forest. **_Biotropica_** 45:728-736.  [Ph. D. Thesis Chapter].  

29. Lucas, CM, EM Bruna, and CMN Nascimento. 2013. Seedling co-tolerance of multiple stressors in a disturbed tropical floodplain forest. **_Ecosphere_** 4(1):3.  [Ph. D. Thesis Chapter].  

30. Copeland, SM, EM Bruna, LV Barbosa Silva, MC Mack, and HL Vasconcelos. 2012. Short-term effects of elevated precipitation and nitrogen on soil fertility and plant growth in a Neotropical savanna. **_Ecosphere_** 3.  [Ph. D. Thesis Chapter].  **_Data:_** [Dryad](https://doi.org/10.5061/dryad.dg380p9q)


31. Iacona, GD, LK Kirkman, and EM Bruna. 2012. Experimental test for facilitation of seedling recruitment by the dominant bunchgrass in a fire-maintained savanna. **_PLoS ONE_** 7.  [Ph. D. Thesis Chapter].  

32. Mundim, FM, EM Bruna, EHM Vieira-Neto, and HL Vasconcelos. 2012. Attack frequency and the tolerance to herbivory of Neotropical savanna trees. **_Oecologia_** 168:405-414.  [Ph. D. Thesis Chapter].  

33. Passmore, HA, EM Bruna, SM Heredia, and HL Vasconcelos. 2012. Resilient networks of ant-plant mutualists in Amazonian forest fragments. **_PLoS ONE_** 7(8):e40803. **_Data:_** [Dryad](https://doi.org/10.5061/dryad.pf2r0)  

34. Vasconcelos, HL, and EM Bruna. 2012. Arthropod responses to the experimental isolation of Amazonian forest fragments. **_Zoologia_** 29:515-530. **_Invited review_**.  

36.	Palumbo, MJ, SA Johnson, FM Mundim, AC Wolf, S Arunachalam, O Gonzalez, A Lau, JL Ulrich, A Washuta, EM Bruna. 2012. Harnessing smartphones for ecological education, research, and outreach.  **_Bulletin of the Ecological Society of America_** 93(4): 390-393.  

35. Bruna, EM, and AS de Andrade. 2011. Edge effects on growth and biomass partitioning of an Amazonian understory herb (_Heliconia acuminata_; Heliconiaceae). **_American Journal of Botany_** 98:1727-1734. **_Data:_** [Dryad](https://doi.org/10.5061/dryad.553hc134)  

36. Bruna, EM, TJ Izzo, BD Inouye, M Uriarte, and HL Vasconcelos. 2011. Asymmetric dispersal and colonization success of Amazonian plant-ants queens. **_PLoS ONE_** 6:e22937. **_Data:_** [Dryad](https://doi.org/10.5061/dryad.h6t7g)  

37. Ferreira, AV, EM Bruna, and HL Vasconcelos. 2011. Seed predators limit plant recruitment in Neotropical savannas. **_Oikos_** 120:1013-1022.  

39.	Uriarte, M. Anciães, M. T.B. da Silva, P. Rubim, E. Johnson, and EM Bruna.  2011. Disentangling the drivers of reduced long-distance seed dispersal by birds in an experimentally fragmented  landscape.  **_Ecology_** 92(4): 924-93. **_2012 Outstanding Paper in Landscape Ecology, Int. Assoc. for Landscape Ecology - US Chapter_**.  

38. Gagnon, PR, EM Bruna, P Rubim, MR Darrigo, RC Littell, M Uriarte, and WJ Kress. 2011. Growth of an understory herb is chronically reduced in Amazonian forest fragments. **_Biological Conservation_** 144:830-835.  

39. Laurance, WF, JLC Camargo, RCC Luizao, SG Laurance, SL Pimm, EM Bruna, PC Stouffer, GB Williamson, J Benitez-Malvido, HL Vasconcelos, KS Van Houtan, CE Zartman, SA Boyle, RK Didham, A Andrade, and TE Lovejoy. 2011. The fate of Amazonian forest fragments: A 32-year investigation. **_Biological Conservation_** 144:56-67. **_Highlighted as 'Exceptional' by the Faculty of 1000_**.  

40. Pitman, NCA, J Widmer, CN Jenkins, G Stocks, L Seales, F Paniagua, and EM Bruna. 2011. Volume and geographical distribution of ecological research in the Andes and the Amazon, 1995-2008. **_Tropical Conservation Science_** 4:64-81.    

1. Girão, LC, AV Lopes, M Tabarelli, and EM. Bruna. 2011. Changes in Tree Reproductive Traits Reduce Functional Diversity in a Fragmented Atlantic Forest Landscape. pp. 71-94 In: _Reproductive Physiology in Plants_ (P Stewart and S Globig, Eds). CRC Press. (reprint of Girão _et al_. 2007, _PLoS ONE_)

41. Bruna, EM, JF Guimarães, CT Lopes, R Pacheco, KG Facure, FG Lemos, and HL Vasconcelos. 2010. Mammalia, Estação Ecologica do Panga, a Cerrado protected area in Minas Gerais state, Brazil. **_CheckList: Journal of Species Lists and Distribution_** 6:668-675.  

42. Fiske, IJ, and EM Bruna. 2010. Alternative spatial sampling in studies of plant demography: Consequences for estimates of population growth rate. **_Plant Ecology_** 207:213-225. [Ph. D. Thesis Chapter].    

43. Fortini, LB, EM Bruna, DJ Zarin, SS Vasconcelos, and IS Miranda. 2010. Altered resource availability and the population dynamics of tree species in Amazonian secondary forests. **_Oecologia_** 162:923-934.    

44. Iacona, GD, LK Kirkman, and EM Bruna. 2010. Effects of resource availability on seedling recruitment in a fire-maintained savanna. **_Oecologia_** 163:171-180.  [Ph. D. Thesis Chapter].  

45. Quitete Portela, RDC, EM Bruna, and FA Maes dos Santos. 2010. Demography of palm species in Brazil's Atlantic Forest: A comparison of harvested and unharvested species using matrix models. **_Biodiversity and Conservation_** 19:2389-2403. [Ph. D. Thesis Chapter].   

46. Quitete Portela, RdC, EM Bruna, and FA Maes dos Santos. 2010. Are protected areas really protecting populations? A test with an Atlantic Rain Forest palm. **_Tropical Conservation Science_** 3:361-372.  [Ph. D. Thesis Chapter].  

47. Ribeiro, MBN, EM Bruna, and W Mantovani. 2010. Influence of post-clearing treatment on the recovery of herbaceous plant communities in Amazonian secondary forests. **_Restoration Ecology_** 18:50-58.  [Ph. D. Thesis Chapter].  

48. Uriarte, M, EM Bruna, P Rubim, M Anciaes, and I Jonckheere. 2010. Effects of forest fragmentation on the seedling recruitment of a tropical herb: assessing seed vs. safe-site limitation. **_Ecology_** 91:1317-1328.

16.	Bruna, EM 2010. Journals can advance tropical biology and conservation by requiring data archiving.  **_Biotropica_** 42(4): 399-401.

49. Bruna, EM, IJ Fiske, and MD Trager. 2009. Habitat fragmentation and plant populations: Is what we know demographically irrelevant? **_Journal of Vegetation Science_** 20:569-576.   

50. Callis, KL, LR Christ, J Resasco, DW Armitage, JD Ash, TT Caughlin, SF Clemmensen, SM Copeland, TJ Fullman, RL Lynch, C Olson, RA Pruner, EHM Vieira-Neto, R West-Singh, and EM Bruna. 2009. Improving Wikipedia: educational opportunity and professional responsibility. **_Trends in Ecology \& Evolution_** 24:177-179.    

51. Cardoso, E, MIB Cruzeiro Moreno, E. M., and HL Vasconcelos. 2009. Mudanças fitofisionômicas no Cerrado: 18 anos de sucesão ecológica na Estação Ecológica do Panga, Uberlândia, MG. **_Caminhos de Geografia_** 10:254-268.    

52. Côrtes, M, V Gowda, WJ Kress, EM Bruna, and M Uriarte. 2009. Characterization of 10 microsatellite markers for the understory Amazonian herb _Heliconia acuminata_. **_Molecular Ecology Resources_** 9 1261-1264. **_Data:_** GenBank [Ph. D. Thesis Chapter].  

53. Dattilo, W, TJ Izzo, BD Inouye, HL Vasconcelos, and EM Bruna. 2009. Recognition of host plant volatiles by _Pheidole minutula_ Mayr (Myrmicinae), an Amazonian ant-plant specialist. **_Biotropica_** 41:642-646.   

54. Izzo, TJ, EM Bruna, HL Vasconcelos, and BD Inouye. 2009. Cooperative colony founding alters the outcome of interspecific competition between Amazonian plant-ants. **_Insectes Sociaux_** 56:341-345.  

55. Kainer, KA, ML DiGiano, AE Duchelle, LHO Wadt, E Bruna, and JL Dain. 2009. Partnering for greater success: Local stakeholders and research in tropical biology and conservation. **_Biotropica_** 41:555-562.  

56. Vasconcelos, HL, R Pacheco, RC Silva, PB Vasconcelos, CT Lopes, AN Costa, and EM Bruna. 2009. Dynamics of the leaf-litter arthropod fauna following fire in a Neotropical woodland savanna. **_PLoS ONE_** 4:e7762.  

57. Bruna, EM, MR Darrigo, AM Furuya Pacheco, and HL Vasconcelos. 2008. Interspecific variation in the defensive responses of ant mutualists to plant volatiles. **_Biological Journal of the Linnean Society_** 94:241-249.  

58. Costa, AN, HL Vasconcelos, EHM Vieira-Neto, and EM Bruna. 2008. Do herbivores exert top-down effects in Neotropical savannas? Estimates of biomass consumption by leaf-cutter ants. **_Journal of Vegetation Science_** 19:849-854. [Ph. D. Thesis Chapter].  

59. Fiske, IJ, EM Bruna, and BM Bolker. 2008. Effects of sample size on estimates of population growth rates calculated with matrix models. **_PLoS ONE_** 3:e3080.  [Ph. D. Thesis Chapter].  

60. Morris, WF, CA Pfister, S Tuljapurkar, CV Haridas, CL Boggs, MS Boyce, EM Bruna, DR Church, T Coulson, DF Doak, S Forsyth, J-M Gaillard, CC Horvitz, S Kalisz, BE Kendall, TM Knight, CT Lee, and ES Menges. 2008. Longevity can buffer plant and animal populations against changing climatic variability. **_Ecology_** 89:19-25.    

61. Stocks, G, L Seales, F Paniagua, E Maehr, and EM Bruna. 2008. The geographical and institutional distribution of ecological research in the tropics. **_Biotropica_** 40:397-404. **_Data:_** [Dryad](https://doi.org/10.5061/dryad.9097)

65.	Vasconcelos, HL, AN Costa, EHM Vieira-Neto, and EM Bruna.  2008.  Efeitos da herbivoria por formigas  do gênero Atta Fabricius, 1804 (Hymenoptera: Formicidae) sobre a vegetação do Cerrado. Pp. 301-311 In: Insetos Sociais: da biologia a aplicação. Vilela, E.F.; Santos, I.A.; Schoereder, J.H.; Lino Neto, J.; Campos, L.A.O. \& Serrão, J.E. (Eds.).  Editora da Univ. Federal de Viçosa. Viçosa, MG.  

62. Girao, LC, AV Lopes, M Tabarelli, and EM Bruna. 2007. Changes in tree reproductive traits reduce functional diversity in a fragmented Atlantic forest landscape. **_PLoS ONE_** 2:e908.   

63. De Athayde, SF, GM Da Silva, J Kaiabi, M Kaiabi, HR De Souza, K Ono, and EM Bruna. 2006. Participatory research and management of Arumã (_Ischnosiphon gracilis_ Rudge Koern., Marantaceae) by the Kaiabi people in the Brazilian Amazon. **_Journal of Ethnobiology_** 26:36-59. **_Cover article_**.  

64. Boyce MS, Haridas CV, Lee CT, \& The Nceas Stochastic Demography Working Group. 2006. Demography in an increasingly variable world. **_Trends in Ecology \& Evolution_** 21(3):141-8. doi: 10.1016/j.tree.2005.11.018.

64. Kainer, KA, M Schmink, H Covert, JR Stepp, EM Bruna, JL Dain, S Espinosa, and S Humphries. 2006. A graduate education framework for Tropical Conservation \& Development. **_Conservation Biology_** 20:3-13.   

65. Trager, MD, and EM Bruna. 2006. Effects of plant age, experimental nutrient addition and ant occupancy on herbivory in a Neotropical myrmecophyte. **_Journal of Ecology_** 94:1156-1163.  

66. Vasconcelos, HL, EHM Vieira-Neto, FM Mundim, and EM Bruna. 2006. Roads alter the colonization dynamics of a keystone herbivore in Neotropical savannas. **_Biotropica_** 38:661-665. **_Recipient of the Julie S. Denslow Prize from the Association for Tropical Biology and Conservation_**.    

67. Bruna, EM, and MBN Ribeiro. 2005. Regeneration and population structure of _Heliconia acuminata_ in Amazonian secondary forests with contrasting land-use histories. **_Journal of Tropical Ecology_** 21:127-131.  

68. Bruna, EM, and MK Oli. 2005. Demographic effects of habitat fragmentation on a tropical herb: Life-table response experiments. **_Ecology_** 86:1816-1824.  

69. Bruna, EM, and MBN Ribeiro. 2005. The compensatory responses of an understory herb to experimental damage are habitat-dependent. **_American Journal of Botany_** 92:2101-2210.    

71. Bruna, EM, HL Vasconcelos, and S Heredia. 2005. The effect of habitat fragmentation on communities of mutualists: Amazonian ants and their host plants. **_Biological Conservation_** 124:209-216.  

72. Doak, DF, WF Morris, C Pfister, BE Kendall, and EM Bruna. 2005. Correctly estimating how environmental stochasticity influences fitness and population growth. **_American Naturalist_** 166:E14-E21.  **_Highlighted as a 'Must Read' by the Faculty of 1000_**.  

73. Lapola, DM, EM Bruna, CG de Willink, and HL Vasconcelos. 2005. Ant-tended hemiptera in Amazonian myrmecophytes:  Patterns of abundance and implications for mutualism function. **_Sociobiology_** 46:433-442.   

78.	Bruna EM \& W. J. Kress.  2005.  Forest fragments and plant reproduction in Amazonian Brazil.  Pp. 141-146 in G. A. Krupnick \& W. J. Kress, eds. Plant conservation: a natural history approach.  U. of Chicago Press.  

18.	Bruna, EM and K. A. Kainer.  2005.  A delicate balance in Amazonia.  **_Science_** 307:1044-1045. [Response to published article].  

74. Bruna, EM, WJ Kress, F Marques, and OF da Silva. 2004. _Heliconia acuminata_ reproductive success is independent of local floral density. **_Acta Amazonica_** 34:467-471. **_Cover Article_**.  

75. Bruna, EM, DM Lapola, and HL Vasconcelos. 2004. Interspecific variation in the defensive responses of obligate plant-ants: Experimental tests and consequences for herbivory. **_Oecologia_** 138:558-565.  

76. Nakazono, EM, EM Bruna, and RCG Mesquita. 2004. Experimental harvesting of the non-timber forest product _Ischnosiphon polyphyllus_ in central Amazonia. **_Forest Ecology and Management_** 190:219-225.  

80.	Lapola, D. M., E. M Bruna, and H. L. Vasconcelos.  2004.  Amizade tênue: relações mutualísticas entre plantas e formigas na Amazônia.  **_Ciência Hoje_** 34(204): 28-33.  

83.	Bruna, EM  2004.  Biological impacts of deforestation and fragmentation.  Pages 85-90 in The Encyclopaedia of Forest Sciences.  J. Burley, J Evans, and J Youngquist, (eds.).  Elsevier Press, London.  

77. Bruna, EM. 2003. Are plant populations in fragmented habitats recruitment limited? Tests with an Amazonian herb. **_Ecology_** 84:932-947.  

78. Lapola, DM, EM Bruna, and HL Vasconcelos. 2003. Contrasting responses to induction cues by ants inhabiting _Maieta guianensis_ (Melastomataceae). **_Biotropica_** 35:295-300.    

79. Bruna, EM. 2002. Effects of forest fragmentation on _Heliconia acuminata_ seedling recruitment in central Amazonia. **_Oecologia_** 132:235-243.  

80. Bruna, EM, and WJ Kress. 2002. Habitat fragmentation and the demographic structure of an Amazonian understory herb ( _Heliconia acuminata_). **_Conservation Biology_** 16:1256-1266.  

81. Bruna, EM, O Nardy, SY Strauss, and SP Harrison. 2002. Experimental assessment of _Heliconia acuminata_ growth in a fragmented Amazonian landscape. **_Journal of Ecology_** 90:639-649.  **_Recipient of the _John L. Harper Prize_ from the British Ecological Society_**.  

82. Laurance, WF, TE Lovejoy, HL Vasconcelos, EM Bruna, RK Didham, PC Stouffer, C Gascon, RO Bierregaard Jr, SG Laurance, and E Sampaio. 2002. Ecosystem decay of Amazonian forest fragments, a 22 year investigation. **_Conservation Biology_** 16:605-618.    

83. Hoeksema, JD, and EM Bruna. 2000. Pursuing the big questions about interspecific mutualism: A review of theoretical approaches. **_Oecologia_** 125:321-330.  

84. Bruna, EM. 1999. Seed germination in rainforest fragments. **_Nature_** 402:139.  

85. Harrison, S, and E Bruna. 1999. Habitat fragmentation and large-scale conservation: What do we know for sure? **_Ecography_** 22:225-232.  

86. Bruna, EM, RN Fisher, and TJ Case. 1996. New evidence of habitat segregation between two cryptic species of Pacific skinks (_Emoia cyanura_ and _E. impar_). **_Copeia_** 1996:998-1004.  

87. Bruna, EM, RN Fisher, and TJ Case. 1996. Morphological and genetic evolution appear decoupled in Pacific skinks (Squamata: Scincidae: _Emoia_). **_Proceedings of the Royal Society of London Series B Biological Sciences_** 263:681-688. **_Cover article_**.  **_Data:_** GenBank

88. Bruna, EM, RN Fisher, and TJ Case. 1995. Cryptic species of Pacific skinks (_Emoia_): Further support from mitochondrial DNA sequences. **_Copeia_** 1995:981-983. **_Data:_** GenBank


### ANNOUNCEMENTS, BOOK REVIEWS, \& CORRESPONDENCE

1. Bruna, EM, Powers, JS and Lohmann, LG 2022. 2022 ATBC Honorary Fellows. **_Biotropica_** 54:1521-1522. https://doi.org/10.1111/btp.13170 [Essay]  

1.	KR Zamudio, A Kellner, C Serejo, MR de Britto, CB Castro, PA Buckup, DO Pires, M Couri, A Brilhante Kury, I Azevedo Cardoso, ML Monné, J Pombal Jr, C Mello Patiu, V Padula, A Dias Pimenta, R Ventura, E  Hajdu, J Zanol, EM Bruna, J Fitzpatrick, LA Rocha. 2018. Lack of support for Science and Education Fails Brazilian Society - Again. **_Science_** 361(6409): 1322-1323. [Correspondence] 

1.	Bruna, EM. 2017. Invasive new US visa rules hurt Americans too. **_Americas Quarterly_**. http://www.americasquarterly.org/content/academic-collaboration [Invited essay]  

1.	Harms, K, J Benitez-Malvido, \& EM Bruna. 2017. 2016 Alwyn Gentry Awards and Luis F. Bacardi Award Recipients. \textit{\textbf{Biotropica}} [Essay].  

1.	Bruna, EM 2016. Biotropica renames its "Outstanding Paper" awards for two pioneering scientists. \textit{\textbf{Biotropica}} 48(6): 929 [Essay].  

1.	EM Bruna, M. Peña-Claros, Bryan Finega, and B. A. Kaplin. 2016. A New Data Archiving Policy for _Biotropica._ \textit{\textbf{Biotropica}} 48(2): 139-140. [Editorial].  

1.	V. Christianini, P. S. Oliveira, EM Bruna, H. L. Vasconcelos. 2014. Fauna in decline: Meek shall inherit. \textit{\textbf{Science}} 354(6201):1129. [Correspondence]  

1.	Bruna, EM. 2014. Biotropica to become an 'online-only' journal in 2015. \textit{\textbf{Biotropica}} 46(6): 782. [Editorial].  

1.	Bruna, EM. 2014. 2014 Alwyn Gentry Awards, Luis F. Bacardi Award, and ATBC Honorary Fellows. \textit{\textbf{Biotropica}} 46(6): 779-781. [Essay].  

1.	Bruna, EM. 2014. A word from the Editor-in-Chief. \textit{\textbf{Biotropica}} 46(4): 377 [Editorial].

1.	Bruna, EM 2010.  Announcement: _Biotropica_ Award for Excellence in Tropical Biology and Conservation. \textit{\textbf{Biotropica}} 42(4): 518-519 [Essay].  

1.	Bruna, EM  2003.  "For the Love of Nature".  **_Science: Next Wave Online Magazine_**. [Invited essay].

1.	Bruna, EM 2002. A biologist's guide to fragmenting the Amazon (including why it's a bad idea).  **_Conservation Biology_**, 16(5): 1446-1448. [Book review].  


### SOFTWARE \& TECHNICAL MANUALS

1. Fournier, A., F. Stevens, M. Boone, and EM Bruna. 2025. refsplitr. author name disambiguation, author georeferencing, and mapping of coauthorship networks with Web of Science data. R package version 1.2.0. https://docs.ropensci.org/refsplitr _(Package maintained by EM Bruna)_

1. HerbVar Steering Committee 2024. HerbVar: Project Manual and Field protocols (v0.9.0). Zenodo. https://doi.org/10.5281/zenodo.14232308


### SCHOLARLY PRODUCTIVITY in ONLINE OUTLETS

1.	Johnson-Ulrich, L and EM Bruna. 2019. Emilio Bruna: Seed germination is lower in fragmented rainforest. **_Project Biodiversify_** [projectbiodiversify.org/emilio-bruna/](projectbiodiversify.org/emilio-bruna/) [Teaching materials]

1. _Jean Theodore Descourtilz (1796-1855) and the Ornithologie Bresilienne_: Website created with graduate students in response to request from the UF Harn Museum of Art to analyze an aqcuisition of chromolithographs https://descourtilz.wordpress.com/

1. Over 30 Wikipedia pages edited with students as part of graduate and undergraduate courses in Plant Animal Interactions and Tropical Ecology. _Complete list at http://en.wikipedia.org/wiki/User:Embruna _   


### PUBLICATIONS \& BOOK CHAPTERS WRITTEN BY MENTEES DURING THEIR LAB TENURE 

**_See co-authorship philosophy above; ^GS^Graduate Student,^PD^Postdoctoral Researcher, ^VS^Visiting Student_** 
1. Sridhar, H, J Chanam, \& ^GS^P Hariharan. 2026. Spark, Scaffold, Substrate: The Multiple Hidden Roles of Natural History in Ecology and Evolutionary Biology. **_The American Naturalist_**. 208(3).

1. Ramesh, V, ^GS^P Hariharan, PR Gupte _et al._ 2025. Grassland bird species decline with colonial-era landscape change in a tropical montane ecosystem. _Global Change Biology_ 31(7): e70358. https://doi.org/10.1111/gcb.70358.

1. Ramesh, V, ^GS^P Hariharan, PR Gupte, AV Mohan, VA Akshay, A Rajan, C Das, I Lockwood, VV Robin, MW Tingley, \& R DeFries. 2025. Grassland bird species decline with colonial-era landscape change in a tropical montane system. **_Global Change Biology_**. 31(7), e70358. https://doi.org/10.1111/gcb.70358

1. Ramesh, V, ^GS^P Hariharan, VA Akshay, P Choksi, S Khanwilkar, R DeFries and VV Robin. 2023. Using passive acoustic monitoring to examine
the impacts of ecological restoration on faunal biodiversity in the Western Ghats. **_Biological Conservation_**. 282, 110071. https://doi.org/10.1016/j.biocon.2023.110071

1.	^VS^Costa, AN and ^GS^Vieira-Neto, EHM. 2016. Species turnover regulates leaf-cutter ant densities in environmental gradients across the Brazilian Cerrado. **_Journal of Applied Entomology_** 140(2016):474-478. 

2.	M. Tabarelli, ^GS^AV Neto, IR Leal, and AV Lopes (Eds.).  2014. Serra Grande: Uma Floresta de ideas.  Editora Universitária da UFPE. Recife. 672 pp.

3.	Cruz Neto O, ^GS^Aguiar AV, Twyford AD, Neaves LE, Pennington RT, and Lopes AV. 2014. Genetic and ecological outcomes of _Inga vera_ subsp _affinis_ (Leguminosae) Tree plantations in a fragmented tropical landscape. **_PLoS ONE_** 96: e99903.

4.	^GS^Lucas CM, Schoengart J, Sheikh P, Wittmann F, Piedade MTF, and McGrath DG. 2014. Effects of land-use and hydroperiod on aboveground biomass and productivity of secondary Amazonian floodplain forests. **_Forest Ecology and Management_** 319:116-127. [Ph.D. thesis chapter].

5.	Soares SM, Machado IC, ^GS^Aguiar AV, and Lopes AV. 2014. Dioecy in the Caatinga, a Brazilian tropical  dry forest: typical reproductive traits of a low frequent sexual system. **_Plant Systematics and Evolution_** 300:1299-1311.

6.	Tabarelli M, ^GS^Aguiar AV, Robeiro MC, and Metzger JP. 2012. A conversão da Floresta Atlântica em paisagens antrópicas: lições para a conservação da diversidade biológica das florestas tropicais. **_Interciencia_** 37:88-92.

7.	^GS^Lucas CM, Mekdece F, Nascimento CMN, Holanda A-SS, Braga J, Dias S, Sousa S, Rosa PS, and Suemitsu C. 2012. Effects of short-term and prolonged saturation on seed germination of Amazonian floodplain forest species. **_Aquatic Botany_** 99:49-55. [Ph.D. thesis chapter].

8.	Horn, M, SB Correa, P Parolin, BJA Pollux, JT Anderson, ^GS^C Lucas, P Widmann, A Tiju, M Galetti, and M. Goulding. 2011. Seed dispersal by fishes in tropical and temperate waters: the growing evidence.  **_Acta Oecologia_** 37:561-577.

9.	Tabarelli M, ^GS^AV Aguiar, MC Ribeiro, JPMetzger JP, and Peres CA. 2010. Prospects for biodiversity conservation in the Atlantic Forest: lessons from aging human-modified landscapes. **_Biological Conservation_** 143:2328-2340.

10.	^PD^Gagnon, P.R., Passmore, H.A., Platt, W.J., Paine, C.E.T., Myers, J.A. and Harms, K.E. 2010. Does pyrogenicity protect burning plants? **_Ecology_** 91: 3481-3486 **_Cover Article_**. 

11.	^GS^Trager, M. D., S. Bhotika, J. A. Hostetler, G. V. Andrade, M. A. Rodriguez-Cabal, C. S. McKeon, C. W. Osenberg, and B. M. Bolker. 2010. Benefits for plants in ant-plant protective mutualisms: A meta-analysis. **_PLoS ONE_** 5:e14308.

12.	Parolin P, ^GS^C. Lucas, M.T.F. Piedade, \& F. Wittmann. 2010. Drought responses of flood-tolerant trees in Amazonian Floodplains. **_Annals of Botany_** 105(1):129-39. [Invited article]

13.	^GS^Aguiar AV, and Tabarelli M. 2010. Edge effects and seedling bank depletion: the role played by the early successional palm _Attalea oleifera_ (Arecaceae) in the Atlantic Forest. **_Biotropica_** 42:158-166.

14.	Duchelle, AE, K. Biedenweg, ^GS^C Lucas, J Radachowsky, D Wojcik, M Londres, D Alvira, W Bartels, A Virapongse, and KA Kainer. 2009. Graduate students and knowledge exchange with local stakeholders: Possibilities and preparation.  **_Biotropica_** 41(5): 578-585.

15.	^GS^Lucas, CM. 2008. Within flood season variation in fruit consumption and seed dispersal by two characin fishes of the Amazon. **_Biotropica_** 40(5): 581-589.

16.	Hartter, J, ^GS^C Lucas, A Gaughan, and LL Aranda. 2008. Detecting tropical dry forest succession in a shifting cultivation mosaic of the Yucatán Peninsula, Mexico.  **_Applied Geography_** 28: 134-149.


<!-- ### OTHER PUBLICLY ARCHIVED DATASETS \& CODE  -->
<!-- 1. Heliconia acuminata Demographic Data:   -->

<!-- > Seed set: http://dx.doi.org/10.6084/m9.figshare.1273926   -->
<!-- > Leaf Length vs. Area: http://dx.doi.org/10.6084/m9.figshare.92691   -->
<!-- > Leaf number, leaf area, shoot number, and height of reproductive _H. acuminata_: https://zenodo.org/records/5041931   -->

<!-- 1. Emilio M. Bruna, Audrey C. Smith, Leandra Merz, Jesse Borden, Chris K. Gulick, \& Akhil R. Kshirsagar. (2021). Data and Code from: Assessing the effect of article processing charges on the geographic diversity of authors using Elsevier's 'Mirror Journal' system. (v1.0.0). Zenodo. https://doi.org/10.5281/zenodo.5500293 -->

<!-- 2. Costa et al. (2018) Data from: The relative influence of an environmental gradient and an ecosystem engineer on seedling communities in a Neotropical savanna. Dryad Digital Repository https://doi:10.5061/dryad.vf1gq0q  -->

<!-- 2.	Emilio M. Bruna. (2018, September 13). embruna/Costa-etal-PeerJ-2018: First release of code used in Costa et al 2018 PeerJ (Version v1.0.0). Peerj. Zenodo. http://doi.org/10.5281/zenodo.1418482  -->

<!-- 3.	Espin, J., et al. (2017) Data from: A persistent lack of international representation on editorial boards in environmental biology. Dryad Digital Repository. https://doi.org/10.5061/dryad.mh189 -->

<!-- 4.	Bruna, EM 2017. Code (v1.0) used in 'A persistent lack of international representation on editorial boards in environmental biology, Espin et al.' Zenodo http://doi.org/10.5281/zenodo.1067800 -->

<!-- 5.	Vieira-Neto EHM, Vasconcelos HL, Bruna EM (2016) Data from: Roads increase population growth rates of a native leaf-cutter ant in Neotropical savannahs. Journal of Applied Ecology https://doi.org/10.5061/dryad.1s4f2 -->

<!-- 6.	Costa AN, Vasconcelos HL, Bruna EM (2016) Data from: Biotic-drivers of seedling establishment in Neotropical savannas: selective granivory and seedling herbivory by leaf-cutter ants as an ecological filter. Journal of Ecology https://doi.org/10.5061/dryad.1hj56 -->

<!-- 7.	Mundim FM, Bruna EM (2016) Data from: Is there a temperate bias in our understanding of how climate change will alter plant-herbivore interactions? A meta-analysis of experimental studies. The American Naturalist https://doi.org/10.5061/dryad.dn048 -->

<!-- 8.	Bruna EM, Izzo TJ, Inouye BD, Vasconcelos HL (2014) Data from: Effect of mutualist partner identity on plant demography. Ecology https://doi.org/10.5061/dryad.5kc45 -->

<!-- 9.	Bruna E.M. 2014. Code (v1.0) used in "Effect of mutualist partner identity on plant demography. Ecology". Zenodo: doi:10.5281/zenodo.11650. Available for improvement at https://github.com/embruna/ant-plant-demography.  -->

<!-- 10.	Cho AH, Johnson SA, Schuman CE, Adler JM, Gonzalez O, Graves SJ, Huebner JR, Marchant DB, Rifai SW, Skinner I, Bruna EM (2014) Data from: Women are underrepresented on the editorial boards of journals in environmental biology and natural resource management. PeerJ _Data:_ https://doi.org/10.5061/dryad.6jn86  -->

<!-- 11.	Cho et al. 2014. Code used in "Women are underrepresented on the editorial boards of journals in environmental biology and natural resource management. PeerJ":: https://github.com/embruna/Editorial-Board-Gender  -->

<!-- 12.	Corrêa Côrtes M. et al. 2013. Low plant density enhances gene dispersal in the Amazonian understory herb Heliconia acuminata. Molecular Ecology Dryad Digital Repository:  https://doi.org/10.5061/dryad.b003f -->

<!-- 13.	Passmore, H. A et al. 2012. Data from: Resilient networks of ant-plant mutualists in Amazonian forest fragments. PLoS ONE.  https://doi.org/10.5061/dryad.pf2r0  -->

<!-- 14.	Copeland, S. M. et al. 2012. Data from: Effects of experimentally elevated precipitation and nitrogen on soil fertility and plant growth in a Neotropical savanna. Ecosphere. Dryad Digital Repository: https://doi.org/10.5061/dryad.dg380p9q -->

<!-- 15.	Bruna EM, et al. 2011. Data from: Asymmetric dispersal and colonization success of Amazonian plant-ants queens. PLoS ONE. Dryad Digital Repository. https://doi.org/10.5061/dryad.h6t7g  -->

<!-- 16.	Bruna, EM and A. Segalin de Andrade. 2011. Data from: Phenotypic plasticity in plant responses to edge effects: experimental test with an Amazonian understory herb. American Journal of Botany. Dryad Digital Repository: https://doi.org/10.5061/dryad.553hc134 -->

<!-- 17.	Microsatellite sequences from Cortes, MC et al. 2009. Characterization of 10 microsatellite markers for the understorey Amazonian herb Heliconia acuminata. Molecular Ecology Resources. Genbank. -->

<!-- 18.	Stocks G. et al. 2008. Data from: The geographical and institutional distribution of ecological research in the tropics. Biotropica. Dryad Digital Repository: htttps://doi.org/10.5061/dryad.9097  -->

<!-- 19.	Sequences used in Mitochondrial DNA sequences from Bruna, EM et al. 1996. Morphological and genetic evolution appear decoupled in Pacific skinks (Emoia). Proc. Roy. Soc. Lond. B: Genbank. -->



<!-- ### MANUSCRIPTS IN PREPARATION (N=11, drafts available upon request).  -->

<!-- 1. Mundim,F. M. E. Vieira-Neto, G. F. Schneider, H. Alborn, EM Bruna. _In preparation._ Simultaneous and Sequential herbivory dynamics: exploring plant defenses strategies amidst drought challenges. **_Ecology_** -->

<!-- 1.	E. S. Parsons and EM Bruna. Assessing the magnitude of abiotic edge effects in fragmented landscapes: revisiting the paradigm with data. In preparation for **_Biological Conservation_**[MA Thesis chapter]. -->

<!-- 1. Kortessis, N., R. J. Fletcher, EM Bruna, T. A. H. Smith, and R. D. Holt. Habitat degradation, dispersal, and matrix effects: Interactions between processes at different spatial scales determine species persistence. In preparation for **_The American Naturalist_**. -->

<!-- 1. Smith, T. A. H., EM Bruna, and R. J. Fletcher. Autotomy in a changing world: How landscape structure alters the prevalence and consequences of limb loss. In preparation for **_Functional Ecology_** -->

<!-- 1.	E. H. M. Vieira-Neto, F. M. Mundim, and EM Bruna. Closed and open demographic models give contrasting patterns of inter-population dynamics in a long-lived superorganism. In preparation for **_Ecology_**. -->

<!-- 1.	Goring, S., K.S. Whitney, EM Bruna, A. L. Jacob, and T. Poisot. Ten Simple Rules to Make Scientific Content More Accessible. Preprint at: DOI: [10.22541/au.150844289.92609826](https://doi.org/10.22541/au.150844289.92609826) -->

<!-- 1.	Tolentino,M., M. Anciães, EM Bruna, and M. Uriarte. **_In revision*. Local and landscape factors mediate reproductive success of _Heliconia acuminata_ in a fragmented landscape. **_Landscape Ecology_**.  -->

<!-- 1.	Corrêa Côrtes, M., M. Uriarte, W. J. Kress, J. B. Illian, and EM Bruna. *In revision*. Why are the spatial distribution of individual plants and genotypes decoupled? A framework to assess the contribution of ecological interactions and environmental filtering. **_Ecology Letters_** [Ph.D. Thesis Chapter]. -->

<!-- 1.	Bruna, EM and W. Hahn. In preparation. Is Latin American ecology really on the rise? Cross-national patterns of productivity and collaboration by Latin American scientists. In preparation for **_PLoS Biology_**. -->

<!-- 1.	Rubim, P., D. M. S. Matos, M. Uriarte, and EM Bruna. Local density influences the growth of seedlings of an Amazonian understory herb. In preparation for **_Biotropica_** [Ph.D. Thesis Chapter]. -->

<!-- 1.	Loiselle, B. A., EM Bruna, R. Buschbacher, and S. Athayde. In prep. How can Brazilian Studies and partnerships with Brazilian institutions shape the perception of Tropical Conservation & Development issues at US universities? Lessons from the University of Florida. Brazilian Studies in the United States: The road ahead. Yale University, FGV, and The Ministry of Education in Brazil. -->

<!-- 1.	Rubim, P. R., M. Uriarte, O. Ferreira, and EM Bruna. Neighborhood effects and the reproductive effort of a tropical perennial herb. In preparation for **_Ecology_** [Ph.D. Thesis Chapter]. -->


<!-- ### WEB PAGES AND OTHER FORMS OF DIGITAL COMMUNICATION -->

<!-- 1. Over 30 Wikipedia pages edited with students as part of graduate and undergraduate courses in Plant Animal Interactions and Tropical Ecology. Complete at: http://en.wikipedia.org/wiki/User:Embruna -->

<!-- 2. Biotropica.org: The Editor"s Blog. A forum for sharing information about the journal Biotropica, its ongoing initiatives, announcements debates on recent publications, and advice for authors. >80,000 views since launched in August 2013, http://biotropica.org/ -->

<!-- 3. BrunaLab.org / -Between the 23s- (lab web page and blog): In addition to hosting publications, data, and other information on current lab activities, we provide resources for the broader community on topics (e.g., resources for beginners on making figures in R or version control with GitHub). We also maintain a compendium on smartphone applications for fieldwork, which is responsible for most of the sites 113,000 page views since January 2013. http://brunalab.org/ -->

<!-- 4. -Jean Theodore Descourtilz (1796-1855) and the Ornithologie Bresilienne.- A website created by TCD graduate students in my Tropical Ecology class to provide background information and an analysis of artwork acquired by the Harn Museum of Art on their behalf. https://descourtilz.wordpress.com/ -->


<!-- text size: https://www.overleaf.com/learn/latex/Font_sizes%2C_families%2C_and_styles -->
<!-- xrfill adjusts line position and width -->
\vspace{0.2cm}
\textbf{\textcolor{darkmidnightblue}{\large{COURSE DEVELOPMENT \& TEACHING EXPERIENCE }}}\xrfill[0.5ex]{1pt}

<!-- ## COURSE DEVELOPMENT & TEACHING EXPERIENCE -->
\vspace{-0.4cm}

### Undergraduate Courses, University of Florida 
\vspace{-0.2cm}
\begingroup 
\setlength{\leftskip}{.2cm}  
The Future of Rain Forests (General Education & UF Quest Program, 2021-2023,2025-2026)  
Ecology and Conservation of Tropical Wildlife (biannually 2003-11 & 2012-2014,2018)  

\endgroup 

\vspace{-0.2cm}
### Graduate Courses, University of Florida 
\vspace{-0.2cm}
\begingroup 
\setlength{\leftskip}{.2cm}  
Data Collection & Management for Latin American Studies (2019-2024, 2026)  
Plant-Animal Interactions (biannually 2004-2010)  
Fundraising for Tropical Conservation & Development (2013 and biannually 2016-2020)  
Scientific Publishing (2014-2015, 2017-2018)  
Ecological Principles for Development & Conservation Professionals (2010-2012)  
Tropical Forest Community Ecology (2000)  
Design and Methods of Research in Latin American Studies (2008, 2010)  
Research Methods in Tropical Conservation & Development (2003-2007)  
Seminar in Wildlife Ecology and Conservation (2003-2004)  
Social Media Workshop (2016)  

\endgroup 
\vspace{-0.2cm}
### Guest Lectures, Participation in Field Courses, & courses taught as Visiting Faculty
\vspace{-0.2cm}
\begingroup 
\setlength{\leftskip}{.2cm}  
**University of Florida:** _Data Management Workshop - LAS Annual Conference_ (2026), _UF SACNAS CV Workshop_ (2023, 2024, 2025), _The Professional Scientist_ (WIS 6934: 2021, 2022), _Diverse Perspectives in Conservation_ (WIS4934: 2021, 2022, 2025), _Human Footprint on the Landscape_ (GEO3352: 2019), _Facets of Sustainability_ (IDS2154: 2011, 2013), _Environmental Planning & Design_ (EES 6051: 2016, 2018), _The Amazon_ (LAS6290: 2009, 2011), _Design & Methods of Research in Latin American Studies_ (Graduate, 2019, 2020), _Discovering Research & Communicating Science_ (IDH3931: 2015), _Wildlife Techniques_ (WIS4945C: 2005), _Development Theory & Practice_ (LAS6938: 2010), _Wildlife Population Modeling_ (WIS6466: 2003, 2006, 2008, 2010), _Principles of Wildlife Ecology & Conservation_ (WIS6934: 2008), _Luso-Brazilian Civilization_ (POW3500: 2004)  **Yale**: Tropical Forest Ecology & Conservation (2024), **Michigan State University** _Decolonizing EEB_ (2024), **INPA:** _Population Ecology_ (2002, 2003), Community Ecology (2002), **OTS**: _3M Field Course_ (2001), _Amazonian Ecosystems Field Course_ (2001), **Antioch College**: Semester Abroad in Brazil (1999, 2001), **Univ. Federal do Amazonas**: _Fragmented Landscapes Field Course_ (2000, 2002), **Univ. Federal de Uberlândia**: _Mechanisms of Species Coexistence_ (Graduate Course, 2006), _Data Management & Archiving_ (Graduate Course, 2012, 2014), _Scientific Publishing_ (Graduate Course, 2011)  
 

\endgroup 

<!-- THIS WILL PUT THINGS CENTERED WITH LINE AROUND -->
<!-- \newcommand*\ruleline[1]{\par\noindent\raisebox{.8ex}{\makebox[\linewidth]{\hrulefill\hspace{1ex}\raisebox{-.8ex}{#1}\hspace{1ex}\hrulefill}}} -->
<!-- \ruleline{\textbf{\textcolor{darkmidnightblue}{STUDENT AND POSTDOCTORAL MENTORSHIP}}}    -->

<!-- THIS WILL PUT THINGS LEFT WITH LINE EXTENDING RIGHT -->
<!-- Raised Rule Command: -->
<!-- Arg 1 (Optional) - How high to raise the rule -->
<!-- Arg 2            - Thickness of the rule -->
<!-- \newcommand{\raisedrule}[2][0em]{\leaders\hbox{\rule[#1]{1pt}{#2}}\hfill}   -->
<!-- <!-- into your document's preface, and then you can accomplish what you were hoping to with: --> 

<!-- text size: https://www.overleaf.com/learn/latex/Font_sizes%2C_families%2C_and_styles -->
<!-- xrfill adjusts line position and width -->
\vspace{0.2cm}
\textbf{\textcolor{darkmidnightblue}{\large{STUDENT \& POSTDOCTORAL MENTORSHIP }}}\xrfill[0.5ex]{1pt}

\vspace{-0.4cm}

<!-- ## STUDENT & POSTDOCTORAL MENTORSHIP -->

### GRADUATE STUDENTS ADVISED
```{r GradStudentsA, echo=FALSE,message = FALSE,warning=FALSE}
# library(flextable)
# https://cran.r-project.org/web/packages/flextable/vignettes/format.html#borders
# df <- grad_advised<-read_csv("./cv_files/grad_advised.csv") %>% filter(role=="advisor") %>% select(-role)
# table1<-df %>% regulartable() %>% autofit() %>% 
# width(j=~Degree,width=.3) %>% width(j=~Year,width=.9) %>% width(j=~Instititution,width=.1)
# table1 <- fontsize(table1, part = "header", size = 11) 
# table1 <- border_remove(x = table1)
# table1
```
\vspace{-0.2cm}
```{r GradStudents, echo=FALSE,message = FALSE,warning=FALSE}
grad_advised<-read_csv("./cv_files/grad_advised.csv") %>% filter(role=="advisor") %>% select(-role)
kbl(grad_advised,digits = 1,
                  format = "latex",
                  align="l",
    col.names = NULL,
    longtable = T,
                  row.names = FALSE,
                  booktabs=T,
                  linesep = "") %>%
  kable_styling(bootstrap_options = c("hover"),
                latex_options = c("scale_down"),
                full_width = F,
                font_size = 10,
                position = "left") %>%
  sub("\\\\toprule", "", .) %>%
  sub("\\\\bottomrule", "", .)
```
\vspace{-0.2cm}
_Students under my supervision have been awarded over $75,000 in grants and awards in support of their research, teaching, and outreach and been the recipients of the following fellowships to support their studies:_ 

*	NSF Predoctoral Fellowships (M Trager, S Copeland)
*	NSF Doctoral Dissertation Enhancement Grant (C Lucas)
*	Inter-American Foundation Grassroots Development Fellowship (C Lucas)
*	Fulbright Fellowship (A Neto) 
*	Science Without Borders Doctoral Fellowship (E V Neto, F Mundim) 
*	FLAS Fellowship (S Parsons)
*	UF NSF IGERT Fellowship (C Lucas)
*	UF Alumni Fellowship (M Trager) 
*	UF Jennings Scholarship (A Neto, M Trager, F Mundim, G Iacona) 
*	UF Doris Lowe & Earl & Verna Lowe Scholarship (P Hariharan, C Lucas, E V Neto) 
* UF Grinter Fellowship (P Hariharan)
* UF William C and Bertha M Cornett Fellowship (P Hariharan)

_Research by my students has also been recognized for its excellence:_ 

* Priyanka Hariharan: ATBC Navjot Sodhi Conservation Research Award (2024), UF CALS International Student Outstanding Achievement Award (2022)
* Thomas Smith: Best Student Presentation, IALE-NA Annual Meeting (2022)
*	Fabiane Mundim: UF IFAS High-Impact Publication (2021), UF WEC High-Impact Publication (2016), Honorable Mention, OTS Outstanding Paper Competition (2012)
*	Ana Cecilia Silva: Honorable Mention, Charles Wood Thesis Award, UF-LATAM (2015)
*	Christine Lucas: Marianne Schmink Outstanding Dissertation Award, UF-TCD (2012)
*	Ernane Vieira Neto: E. Lucy Braun Outstanding Poster Award, ESA Annual Meeting (2011)
*	Shannon Moore: Best Presentation, UF High School Student Science Training Program Symposium (2010)
*	Matthew Trager: Outstanding Research Award - MS Category, Wildlife Graduate Student Association (2005)


### DOCTORAL COMMITTEE SERVICE
\vspace{-0.2cm}
```{r DocCmte, echo=FALSE,message = FALSE,warning=FALSE}
grad_advised<-read_csv("./cv_files/grad_advised.csv") %>% filter(role=="doc_cmte") %>% select(-role)
kbl(grad_advised,digits = 1,
                  format = "latex",
                  align="l",
                  col.names = NULL,
                  row.names = FALSE,
                  booktabs=T,
                  longtable = T, 
                  linesep = "") %>%
  kable_styling(bootstrap_options = c("hover"),
                latex_options = c("scale_down"),
                full_width = F,
                font_size = 10,
                position = "left") %>%
  sub("\\\\toprule", "", .) %>%
  sub("\\\\bottomrule", "", .)
```
<!-- **Wildlife Ecology & Conservation**: David Mason (in progress), Flavia Montaño Centellas (2018), Margo Stoddard (2017), Rajeev Pillay (2016), Rosalyn Johnson (2016), Gonzalo Rivas (2015), Oscar Gonzalez (2015), Dan Thornton (2010), Ivan Diaz (2008), Christa Zweig (2008), Traci Darnell (2006)  **Forest Resources & Conservation**: Andrea Irons (in progress), Demian Gomez Damiano (2020), Jacy Hyde (2019), Christie Klimas (2010)  **Biology (*formerly Depts. of Botany & Zoology*)**: Maria Beatriz Cortez (in progress), Johanna R. Jantzen (2020), Trevor Caughlin (2013), Kristine Callis (2013), Mollie Brooks (2012), Silvia Alvarez-Clare (2012), C. Seabird McKeon (2010), Joseph Veldman (2010), Jennifer Schafer (2010), Alex Jahn (2009), Silvia Lomáscolo (2008), Bonifacio Mostacedo (2007)  **Interdisciplinary Ecology**: Joanna Tucker (2010), Matt Trager	(2009)  **Entomology**: Jade Badon (2018), Pablo Allen (2017), David Serrano (2006)  **Anthropology**: Scott Macrae (2017)  **Columbia University**: Marina Cortes (Ecology, Evolution, Behavior (2012) -->



### MASTERS COMMITTEE SERVICE
\vspace{-0.2cm}
```{r MastersCmte, echo=FALSE,message = FALSE,warning=FALSE}
grad_advised<-read_csv("./cv_files/grad_advised.csv") %>% filter(role=="masters_cmte") %>% select(-role)
kbl(grad_advised,digits = 1,
                  format = "latex",
                  align="l",
        longtable = T, 
    col.names = NULL,
                  row.names = FALSE,
                  booktabs=T,
                  linesep = "") %>%
  kable_styling(bootstrap_options = c("hover"),
                latex_options = c("scale_down"),
                full_width = F,
                font_size = 10,
                position = "left") %>%
  sub("\\\\toprule", "", .) %>%
  sub("\\\\bottomrule", "", .)
```
<!-- **Wildlife Ecology & Conservation**: Lauren Trotta (2016), Brittany Bankovich (2014), Nathan Marcy (2010), Kristen Aaltonen (2009), Mariano Rodriguez Cabal (2007)  **Forest Resources & Conservation**: Jennifer Holm (2007), Christie Klimas (2006)  **Biology (*formerly Depts. of Botany & Zoology*)**: Tania Kim (2006), Silvia Alvarez Clare (2004)  **Interdisciplinary Ecology**: Jamie Cotta (2006)  **Entomology**: Teresa Cooper (2006)  **Latin American Studies**: Crisne Lebron (2015), Carly Voight (2010), Laura Fonseca (2009), Jennifer Davies (2003), **Anthropology**: Autumn Rose (in progress). -->

### PROPOSAL REVIEW, THESIS REVIEW, OR THESIS DEFENSE COMMITTEE
\vspace{-0.2cm}
```{r external, echo=FALSE,message = FALSE,warning=FALSE}
other<-read_csv("./cv_files/grad_advised.csv") %>% filter(role=="other") %>% select(-role,-Program)
kbl(other,digits = 1,
                  format = "latex",
                  align="l",
        longtable = T, 
    col.names = NULL,
                  row.names = FALSE,
                  booktabs=T,
                  linesep = "") %>%
  kable_styling(bootstrap_options = c("hover"),
                latex_options = c("scale_down"),
                full_width = F,
                font_size = 10,
                position = "left") %>%
  sub("\\\\toprule", "", .) %>%
  sub("\\\\bottomrule", "", .)
```

### POSTDOCTORAL, SABBATICAL, & POSTGRADUATE SCHOLARS 
\vspace{-0.2cm}
```{r lab_members, echo=FALSE,message = FALSE,warning=FALSE}
lab_members<-read_csv("./cv_files/lab_members.csv") 
lab_members$cat<-as.factor(lab_members$cat)
lab_members$cat <- ordered(lab_members$cat, levels = c("postdoc","sab" ,"postgrad", "vis_grad","ug","hs"))
lab_members2<-lab_members %>% 
  filter(cat=="postdoc"|cat=="sab"|cat=="postgrad") %>% 
  arrange(order(cat)) %>% 
  select(-cat) %>% 
  drop_na()
  # replace_na(list(Name=" ", Institution=" ", Position=" ", Year=" "))
kbl(lab_members2,digits = 1,
    format = "latex",
    align="l",
    longtable = T, 
    col.names = NULL,
    row.names = FALSE,
    booktabs=T,
    linesep = "") %>%
  kable_styling(bootstrap_options = c("hover"),
                latex_options = c("scale_down"),
                full_width = F,
                font_size = 10,
                position = "left") %>%
  sub("\\\\toprule", "", .) %>%
  sub("\\\\bottomrule", "", .)
```

<!-- **Doctoral Defense Committee:** Helen Belan (U. Federal de Uberlândia, Brazil, 2020), Alan Nilo da Costa (U. Federal de Uberlândia, Brazil, 2013). **Doctoral Dissertation Examiner:** Talita Reis (UNICAMP, Brazil, 2015), Mauricio Bonesso Sampaio (UNICAMP, Brazil, 2012), Maria Rosa Darrigo (UNICAMP, Brazil, 2011), Sjirk Geerts (Stellenbosch University, South Africa, 2011), Rafael Soares de Arruda (INPA, Brazil, 2008), Thiago Junqueira Izzo (INPA, Brazil, 2005). **Doctoral Proposal Defense**: Laura Vivian Barbosa (U. Federal de Uberlândia, Brazil, 2014) **Master's Thesis Defense Committee**: Jonas B. Marvalhas (U. Federal de Uberlândia, Brazil 2013), Ernane Vieira Neto (U. Federal de Uberlândia, Brazil, 2008), Alan Nilo da Costa (U. Federal de Uberlândia, Brazil, 2007), Daniel Mascia Vieira (Universidade de Brasília, Brazil, 2002) -->

<!-- Carine Emer			INPA, Brazil				M.S. Thesis Examiner, 2011 -->
<!-- Flávio de Oliveira Rodrigues	INPA, Brazil				M.S. Thesis Examiner, 2011 -->
<!-- Waldete Castro Lourenço		INPA, Brazil				M.S. Thesis Examiner, 2009 -->
<!-- Simone B. Fontoura		INPA, Brazil				M.S. Thesis Examiner, 2007 -->
<!-- Paula M. Pedrosa			INPA, Brazil				M.S. Thesis Examiner, 2004 -->

<!-- Carine Emer 			INPA, Brazil				M.S. Proposal Examiner, 2009 -->
<!-- Simone Benedet Founatura	INPA, Brazil				M.S. Proposal Examiner, 2006 -->
<!-- Thiago Junqueira Izzo		INPA, Brazil				M.S. Proposal Examiner, 2002 -->

<!-- Melina Alcalá			UFSCar, Brazil				M.S. Qualifying Exam, 2009 -->
<!-- Raquel Negrão Baldoni		UFSCar, Brazil				M.S. Qualifying Exam, 2009 -->

<!-- Pedro B. Vasconcelos		U. Federal de Uberlândia, Brazil		B.S. Thesis Defense, 2009 -->
<!-- Fabiane M. Mundim 		U. Federal de Uberlândia, Brazil		B.S. Thesis Defense, 2007 -->
### UNDERGRADUATE THESIS STUDENTS, HIGH SCHOOL RESEARCHERS, & VISITING GRADUATE STUDENTS 
\vspace{-0.2cm}
```{r lab_members2, echo=FALSE,message = FALSE,warning=FALSE}
lab_members<-read_csv("./cv_files/lab_members.csv") 
lab_members$cat<-as.factor(lab_members$cat)
lab_members$cat <- ordered(lab_members$cat, levels = c("postdoc","sab" ,"postgrad", "vis_grad","ug","hs"))
lab_members3<-lab_members %>% 
  filter(cat=="ug"|cat=="hs"|cat=="vis_grad") %>% 
  arrange(order(cat)) %>% 
  select(-cat) %>% 
  drop_na()
  # replace_na(list(Name=" ", Institution=" ", Position=" ", Year=" "))
kbl(lab_members3,digits = 1,
                  format = "latex",
                  align="l",
        longtable = T, 
    col.names = NULL,
                  row.names = FALSE,
                  booktabs=T,
                  linesep = "") %>%
  kable_styling(bootstrap_options = c("hover"),
                latex_options = c("scale_down"),
                full_width = F,
                font_size = 10,
                position = "left") %>%
  sub("\\\\toprule", "", .) %>%
  sub("\\\\bottomrule", "", .)
```

\begingroup 
\setlength{\leftskip}{.2cm}

*__Field Research Interns:__* _2007_: Wesley Dáttilo da Cruz, Univ. Estadual do Norte Fluminense (Brazil), Jefferson José Valsko da Silva, Universidade do Norte Fluminense (Brazil). _2005-2007_: Renata Pacheco do Nacimento and Cauê Thomé Lopes, Univ. Federal de Uberlândia (Brazil). _2005-2006_: Polyanna Custodio Duarte, Univ. Federal de Uberlândia (Brazil). _2006_: Waldete Castro Lourenço, Univ. Federal do Amazonas (Brazil), Elisabete Marques da Costa, Univ. Federal de Pernambuco (Brazil). _2005_: Bruno Turbiani, Universidade Estadual Paulista (Brazil). _2004_: Cristina Escate, Universidade Federal do Amazonas (Brazil), Simone Benedet, Universidade do Vale do Rio do Sinos (Brazil). _2003_: Cris Follman Jurinitz, Univ. Federal do Rio Grande do Sul (Brazil), Denise Cruz, Univ. Federal do Rio de Janeiro (Brazil). _2003-2004_: David M. Lapola, Universidade Estadual Paulista (Brazil). _2002_: Maria Beatriz Nogueira, Universidade de São Paulo (Brazil). _2001-2002_: Sylvia Heredia, Universidad de los Andes (Colombia). _2001_: Obed Garcia, INPA (Brazil), Olavo Nardy, UNESP (Brazil).

\endgroup


<!-- text size: https://www.overleaf.com/learn/latex/Font_sizes%2C_families%2C_and_styles -->
<!-- xrfill adjusts line position and width -->
\vspace{0.2cm}
\textbf{\textcolor{darkmidnightblue}{\large{SERVICE TO THE UNIVERSITY \& PROFESSIONAL ORGANIZATIONS }}}\xrfill[0.5ex]{1pt}

<!-- ## INSTITUIONAL AND PROFESSIONAL SERVICE -->
\vspace{-0.4cm}
### UNIVERSITY OF FLORIDA  
\vspace{-0.2cm}
\begingroup 
\setlength{\leftskip}{.2cm}

UF Quest Advisory Board (2026-present)  
SEED Grant Committee, Artificial Intelligence & Informatics Research Institute (2026)  
Search Committee, UF QUEST Program Director (2023)   
Selection Committee, UF Undergraduate Faculty Mentor/Advisor of the Year (2022)  
Selection Committee, UF Undergraduate Teacher of the Year  (2022)  
International Outreach Advisory Council, Graduate School (2014-2020)  
Fulbright Fellows Campus Interview Committee  (2019)  
Catalyst Fellow, UF Innovation Academy Catalyst Challenge (2018)  
External Advisory Committee, Institute of Museum & Library Services Grant (2015-2018)  
Selection Committee, UF Teacher/Scholar of the Year (2015)  
Samuel P. Harn Museum of Art Faculty Council (2012-2015)  
Internal Advisory Board, Innovation through Institutional Integration (I3) program (2010-2014)  
Administrative Council, QSE3 IGERT (2008-2014)  
UF Research Foundation Professorship Selection Committee (2010)  
Selection Committee, UF International Educator of the Year (2006)  

\endgroup 

\vspace{-0.2cm}
### UF COLLEGE OF AGRICULTURAL & LIFE SCIENCES
\vspace{-0.2cm}
\begingroup 
\setlength{\leftskip}{.2cm}
IFAS International Programs Advisory Team (2020-2022)  
UF-CALS Outstanding Thesis Selection Committee (2020)  
Commencement Marshal (2018, 2015, 2014, 2004)   
York Distinguished Lecture Selection Committee (2015)  
Search & Screen Committee, CALS Quantitative Sciences Faculty Positions (2013)  
IFAS Faculty Professional Development Advisory Committee (2012-2013)  
Coordinating Committee, School for Natural Resources & the Environment (2012)  
Selection Committee, CALS Teaching and Advising Awards (2012)  
CALS Undergraduate Scholarships & Awards Committee (2007-2009)  
Unifas Focus Group - Teaching Section (2007)  
Unifas Advisory Committee (2006-2007)  
SNRE Seed Grant and New Faculty Support Grant Programs (2005)  


\endgroup 

\vspace{-0.2cm}
### UF TENURE MENTORSHIP & PEER TEACHING EVALUATION COMMITTEES
\vspace{-0.2cm}
\begingroup 
\setlength{\leftskip}{.2cm}
Tenure Mentorship Committee, Dr. Nia Morales (Wildlife Ecology & Conservation, 2021-2025)  
Tenure Mentorship Committee, Dr. Matt Hallet (Wildlife Ecology & Conservation, 2022-2025)  
Tenure Mentorship Committee, Dr. Luke Flory (Agronomy, 2012)  
Chair, Tenure Mentorship Committee, Dr. Robert Fletcher (Wildlife Ecology & Conservation, 2009-2012)  
Chair, Peer Teaching Evaluation Committee, Dr. Ben Baiser (Wildlife Ecology & Conservation, 2025)  
Peer Teaching Evaluation Committee, Dr. Raelene Crandall (Forest Fisheries, & Geomatic Sciences, 2021-2022)  
Chair, Peer Teaching Evaluation Committee, Dr. Ethan White (Wildlife Ecology & Conservation, 2019-2020)  
Chair, Peer Teaching Evaluation Committee, Dr. Morgan Ernest (Wildlife Ecology & Conservation, 2017-2018)  
Peer Teaching Evaluation Committee, Dr. Bill Pine (Wildlife Ecology & Conservation, 2015)  
Chair, Peer Teaching Evaluation Committee, Dr. Lyn Branch (Wildlife Ecology & Conservation, 2011)  
Chair, Peer Teaching Evaluation Committee, Dr. Bill Pine (Wildlife Ecology & Conservation, 2010)  
Chair, Peer Teaching Evaluation Committee, Dr. Robert Fletcher (Wildlife Ecology & Conservation, 2009)  
Peer Teaching Evaluation for Dr. Michelle Mack (Botany, 2006)    
Peer Teaching Evaluation Committee, Dr. Madan Oli (Wildlife Ecology & Conservation, 2003)  

\endgroup 

\vspace{-0.2cm}
### UF DEPARTMENT OF WILDLIFE ECOLOGY & CONSERVATION
\vspace{-0.2cm}
\begingroup 
\setlength{\leftskip}{.2cm}
Research Committee (2020-present, 2005-2007)  
Search & Screen Committee, Research Asst. Professor, Zoo Wildlife Conservation Biology (2020)  
Communications Committee (2014-2020)  
Space Committee (2019)  
Search & Screen Committee, Asst. Professor of Population Ecology (2017)  
Seminar Series Coordinator (2015)  
Graduate Program Committee (2005-2012)  
International Programs Committee (2006-2010)  
Steering Committee, PSTC Compton Fellowship Program (2002-2010)  
Search & Screen Committee, Department Head (2006)  
Seminar Series Coordinator (2003-2004)  
Search & Screen Committee, Asst. Professor of Landscape Ecology (2004)  
Search & Screen Committee, Asst. Professor of Conservation Genetics, WEC (2004)  

\endgroup 

\vspace{-0.2cm}

### UF CENTER FOR LATIN AMERICAN STUDIES
\vspace{-0.2cm}
\begingroup 
\setlength{\leftskip}{.2cm}
Search Committee, TCD Program Director (2026)    
Graduate Admissions Committee (2026)    
Search & Screen Committee, Asst. Professor of Environmental Justice (2023)  
Post-Tenure Review Committee (2023)  
LAS Faculty Advisory Council (2020-2023)  
LAS Bylaws Revision Committee (2020-2021)  
Chair, LAS Awards & Scholarship Committee (2019)  
LAS Merit Evaluation Committee (2015-2017)  
Chair, Search & Screen Committee, Lecturer in Latin American Studies (2015)  
Search & Screen Committee, LAS Communications Coordinator (2015)  
Bacardi Seminar Series Coordinator (2015)  
Curriculum Committee, Masters Program in Sustainable Development Practice (2010-2012)  
LAS Graduate Admissions & Fellowship Committee (2007)  

\endgroup 

\vspace{-0.2cm}
### UF TROPICAL CONSERVATION & DEVELOPMENT PROGRAM
\vspace{-0.2cm}
\begingroup 
\setlength{\leftskip}{.2cm}
Chair, Field Research Grants Committee (2026, 2024, 2020-2021, 2015, 2008-2010, 2004)  
Academic Advisor (2021-2022, 2008-11)  
Steering Committee (2018-2024,2010-16, 2002-08)  
Chair, Marianne Schmink Outstanding Dissertation Award Selection Committee (2022, 2013, 2011, 2010)  
Chair, Fellowship Committee (2022, 2012)  
Advisory Committee, Pan-Amazon Community of Practice Project (2018-2020)  
Faculty Advisor, TCD Graduate Student Organization (2008-2016)  
Organizing Committee, TCD Strategic Planning (2007-2008)  
Advisory Council, Amazon Conservation Leadership Initiative (2004-2005)  
Fellowship Committee (2003-2005)  
Field Research Grants Committee (2003-2004)  

\endgroup  

\vspace{-0.2cm}
### ASSOCIATION FOR TROPICAL BIOLOGY & CONSERVATION 
\vspace{-0.2cm}
\begingroup 
\setlength{\leftskip}{.2cm}  

Search Committee, ATBC Executive Director (2024-2025)  
Search Committee, _Biotropica_ Editor-in-Chief (2024-2025)  
Past-President (2022)   
Chair, Nominations Committee (2022)  
President (2021)  
Annual Meeting Program & Organizing Committees (2021)    
President-Elect (2020)  
Student Field Grants Selection Committee (2020)  
Crowdfunding Workshop, Annual Meeting (2018)  
Annual Meeting Scientific Committee (2017, 2012)  
Councillor (2009-2011)  
Conservation Committee (2004-2006)  
Annual Meeting Symposium Co-organizer (2004)  

\endgroup 

\vspace{-0.2cm}
### ECOLOGICAL SOCIETY OF AMERICA   
\vspace{-0.2cm}
\begingroup 
\setlength{\leftskip}{.2cm}  

Secretary (2022-2025)  
Governing Board (2022-2025)  
Teller's Committee (2022-2025)    
Governance & Leadership Committee (2022-2025)   
Historical Records Committee (2022-2025)   

\endgroup 
\vspace{-0.2cm}
### OTHER UNIVERSITIES & PROFESSIONAL ORGANIZATIONS   
\vspace{-0.2cm}
\begingroup 
\setlength{\leftskip}{.2cm}

External Tenure & Promotion Evaluations (N=26, 2011-present)  
<!-- 2024 (3) 2023 (3), 2022 (4), 2021 (2), 2020 (3), 2018 (2), 2017 (1), 2016 (3), 2015 (1), 2011 (1)   -->
ASN/SSE/SSB Committee on Development a Professional Code of Ethics (2018-2019)  
External Advisor, 2nd Planning Workshop, Brazilian Biodiversity & Ecosystem Services Synthesis Center (2018)  
Discussant, Research & Writing Conference, Florida Education Fund/McKnight Fellowship Program (2016)  
Co-host & Co-organizer, UC Davis Population Biology Workshop (2000)  
Student Representative, UC Davis Population Biology GRT Executive Committee (1996-1997)  
UCSD Selection Committee, UC Tropical Biology Education Abroad Program (1994)  

\endgroup 

\vspace{0.2cm}
\textbf{\textcolor{darkmidnightblue}{\large{EDITORIAL POSITIONS \& SERVICE AS A REVIEWER}}}\xrfill[0.5ex]{1pt}

\begingroup 
\setlength{\leftskip}{.2cm}

**Editor-in-Chief**: _Biotropica_ (2013-2019; _>2400 decisions_)  

**Associate Editor:** _Biotropica_ (2010-2013)  

**Subject Editor:** _PLOS Ecosystems_ (2025-Present), _Biotropica_ (2007-2019), _Ecology_ (2008-2012), _Ecological Monographs_ (2008-2012), _Ecological Applications_ (2008-2012).

__*Ad hoc* journal referee (No. of reviews):__ *Trends in Ecology and Evolution* (1), *Nature Human Behavior* (1),*Nature Ecology & Evolution* (3), *Information, Communication,& Society* (1), _Acta Amazonica_ (1), *Ecology* (5), *Ecology and Evolution* (1),  *Methods in Ecology & Evolution* (1), *Conservation Biology* (5), *Biological Conservation* (7), *Ecological Applications* (1), *Oecologia* (4), *Journal of Tropical Ecology* (2), *Journal of Ecology* (6), *New Phytologist* (1), *American Naturalist* (5), *Plant Ecology* (3), *Proc. Roy. Soc. Lond. B* (1), *Journal of Applied Ecology* (1), *Biotropica* (14), *Ecography* (5), *Ciência y Cultura* (2), *Oikos* (1), *Revista Chilena de Historia Natural* (1), *Biochemical Systematics & Ecology* (1), *Ecology Letters* (2), *Insectes Sociaux* (1), *American J. of Botany* (1), *Scientometrics* (1), *F1000* (1), *Desenvolvimento e Meio Ambiente* (1), *PLoS Biology* (2), *PLoS ONE* (1), *Theoretical Ecology* (1), *Conservation Science & Practice* (1), *Nature Communications* (1)

__US National Science Foundation:__ Committee of Visitors, NSF Office of International Science & Engineering (2023), Proposal Review Panels: DEB Ecology (2005), DEB Population & Evolutionary Processes (2007), DEB Population & Community Ecology (2011, 2013). Ad-hoc proposal review: Division of Environmental Biology Core Programs (9), UMEB Program (2), RUI Program (1), AMERICAS Dissertation Improvement Grants (1), NSF Global Scientists & Engineers Program (1), International Research Fellowship Program (1),  Dimensions of Biodiversity Program (1)

__Other reviewer service__: Red Line Editorial (2019, technical review of a K-12 book), Mindlin Foundation (2018), Cornell University Press (2017), Instituto Serrapliheira (2017), HHMI Professors Program (2017), UF FAES/CRIS Project Proposals (2), U. of Chicago Press (1), Alfred P. Sloan Foundation Minority Ph.D. Program (3). National Geographic Society (1), U. of California Press (1)  

\endgroup 


<!-- text size: https://www.overleaf.com/learn/latex/Font_sizes%2C_families%2C_and_styles -->
<!-- xrfill adjusts line position and width -->
\vspace{0.2cm}
\textbf{\textcolor{darkmidnightblue}{\large{SEMINARS, OUTREACH, \& MEDIA APPEARANCES }}}\xrfill[0.5ex]{1pt}
\vspace{-0.4cm}
<!-- ## SEMINARS & OUTREACH ACTIVITIES -->

### INVITED SEMINARS
\vspace{-0.2cm}
```{r talks, echo=FALSE,message = FALSE,warning=FALSE}
talks<-read_csv("./cv_files/talks.csv", col_types = cols(.default = "c")) %>% 
  select(-year)  %>% 
  replace_na(list(year2=""))
kbl(talks, 
      format    = "latex", 
      align = "l",
      longtable = T, 
      row.names = FALSE,
      col.names = NULL,
      booktabs=T,
      linesep = "") %>%
  kable_styling(bootstrap_options = c("hover"),
                latex_options = c("scale_down"),
                full_width = F,
                font_size = 10,
                position = "left") %>%
  sub("\\\\toprule", "", .) %>%
  sub("\\\\bottomrule", "", .)
```

### INTERVIEWS & MEDIA APPEARANCES  
\begingroup 
\setlength{\leftskip}{.2cm}
2026\hspace{4mm}  [Daily Campus (UConn)](https://dailycampus.com/2026/02/09/how-habitat-fragmentation-and-climate-change-impact-tropical-plants/)   
2024\hspace{4mm}  [Science News](https://www.sciencenews.org/article/invasive-ant-lion-dinner-trees-ecosystem), [Revista Piauí](https://piaui.folha.uol.com.br/materia/ciencia-recalcula-sua-rota/), _Science_     
2023\hspace{4mm}  [_Nature_](https://www.nature.com/articles/d41586-023-00711-z), [Vox](https://www.vox.com/policy-and-politics/23593369/ron-desantis-florida-schools-higher-education-woke)   
2022\hspace{4mm}  [NRC (The Netherlands)](https://www.nrc.nl/nieuws/2022/02/22/systeem-van-open-access-helpt-wetenschappers-uit-arme-landen-niet-a4092976), [Nature](https://www.nature.com/articles/d41586-022-00342-w), WIRED Magazine, [Fox News Business](https://www.foxbusiness.com/politics/university-professors-unhappy-desantis-accountability), [The Scientist](https://www.the-scientist.com/news-opinion/florida-faculty-await-details-on-new-tenure-law-69971),  
&nbsp;\hspace{10mm} [Miami Herald](https://www.miamiherald.com/news/local/education/article263966316.html), [The Independent Florida Alligator](https://www.alligator.org/article/2022/09/blue-and-gold-macaw-uf-campus), [Chronicle for Higher Education](https://www.chronicle.com/article/it-could-have-been-worse-higher-ed-reacts-to-ben-sasse-at-u-of-florida),  
&nbsp;\hspace{10mm} [The Hill](https://thehill.com/changing-america/enrichment/education/3274661-texas-and-florida-take-steps-to-limit-professor-tenure-at-state-schools/), Voice of America, ['How Ecology Works' Podcast](https://www.chemistryworld.com/news/most-scientific-editors-still-come-from-us-or-uk/3008426.article)  
2021\hspace{4mm} Chronicle for Higher Education, [The Florida Independent Alligator](https://www.alligator.org/article/2021/01/what-happens-if-my-classmate-has-covid-19), [Associated Press](https://apnews.com/article/brazil-rio-de-janeiro-trees-abf6a15a804ea1485fbcec0eeed3d8d7), ABC's Nightline  
2020\hspace{4mm} [Washington Post](https://www.washingtonpost.com/nation/2020/11/11/kyle-lamb-data-florida-coronavirus-desantis/), [South Florida Sun Sentinel](https://www.sun-sentinel.com/opinion/commentary/fl-op-com-grimm-ohio-sports-blogger-florida-covid-20201113-wikb2iv3gbd2vcby4hcevwyd6i-story.html)   
2019\hspace{4mm} [USA Today](https://www.usatoday.com/story/news/nation/2019/10/18/amazon-rainforest-still-burning-more-fires-future/4011238002/), CBC News Network, [The Atlantic](https://www.theatlantic.com/science/archive/2019/04/amnh-under-fire-hosting-brazils-jair-bolsonaro/587068/), [The Verge](https://www.theverge.com/2019/9/10/20858834/amazon-fire-brazil-forest-harm-recovery-century), MSNBCs Velshi & Ruhle  
&nbsp;\hspace{10mm} [WIRED Magazine](https://www.wired.com/story/the-horrifying-science-of-the-deforestation-fueling-amazon-fires/), [Mashable](https://mashable.com/article/amazon-rainforest-fires-tipping-point-deforestation), BBC World Service, [BBC News](https://www.bbc.com/mundo/noticias-america-latina-49504502), The Miradas Podcast, [Phys.org](https://phys.org/news/2019-10-amazon-hasnt-outbreaks-month-future.html), [Infobae](https://www.infobae.com/america/america-latina/2019/09/10/las-areas-quemadas-del-amazonas-podrian-tardar-siglos-en-recuperarse-por-completo/)    
2018\hspace{4mm} Popular Science, Associated Press, [National Geographic](https://www.nationalgeographic.com/science/article/news-museu-nacional-fire-rio-de-janeiro-natural-history), Scholastic Classroom Magazines,  
&nbsp;\hspace{10mm} [Minnesota Public Radio's Climate Cast](https://www.mprnews.org/story/2018/11/15/why-brazils-new-president-may-bring-more-climate-disaster-death), Gainesville Sun, Mongabay.com, [NBC News](https://www.nbcnews.com/news/latino/brazil-s-president-elect-jair-bolsonaro-may-smother-rainforest-scientists-n940326), [La Repubblica](https://www.repubblica.it/ambiente/2018/11/27/news/amazzonia_un_milione_di_campi_di_calcio_perduti_in_un_anno-212784490/),  
&nbsp;\hspace{10mm} [Editage Insights](https://www.editage.com/insights/science-journals-might-be-more-international-but-not-their-editorial-boards)     
2017\hspace{4mm} [Nature](https://www.nature.com/articles/548271a), [The Wire (India)](https://thewire.in/science/science-journals-might-international-not-editorial-boards), [Chemistry World](https://www.chemistryworld.com/news/most-scientific-editors-still-come-from-us-or-uk/3008426.article), Gainesville Sun, [Indianapolis  Star](https://www.indystar.com/story/opinion/columnists/suzette-hackney/2017/03/24/hackney-robotics-kids-kindness-triumphs-hate/99578116/), [Phys.org](https://phys.org/news/2017-12-science-international-scientific-editorial-boards.html)   
2016\hspace{4mm} [_Nature_](https://www.nature.com/articles/nature.2016.20830), [Chronicle of Higher Education](https://www.chronicle.com/article/in-donald-trump-phd-scientists-find-catharsis-amid-gags/), [DailyDot](https://www.dailydot.com/unclick/donald-trump-phd-twitter-makes-science-great-again/), [Mongabay.com](https://news.mongabay.com/2016/01/special-issue-of-biotropica-makes-the-case-for-protecting-south-americas-white-sand-forests/) [Mongabay.com](https://news.mongabay.com/2016/04/building-roads-agricultural-expansion-brazil-aiding-spread-insects-harm-crops/)  
2015\hspace{4mm} [WAMC Public Radio](https://www.wamc.org/2015-01-30/is-there-a-pubradiovoice-that-sounds-like-america)  
2014\hspace{4mm} NPRs Morning Edition - "Joe's Big Idea", TV Acre - Rede Globo Rio Branco, Bom Dia Amazônia, [SciDevNet](https://www.scidev.net/global/news/armadillo-mascot-world-cup/)  
&nbsp;\hspace{10mm} TV Gazetta 'Entrevista com Rogério Wenceslau' (Rede Record Rio Branco), Estado de São Paulo  
&nbsp;\hspace{10mm} Nature.com, WUFT, Science/Science Insider, Howler Soccer Magazine's Dummy Podcast  
2013\hspace{4mm} SciDevNet, WSLR Florida-Caribe Radio Show  
2006\hspace{4mm} The Scientist Magazine, Science Minority Scientists Network  
2005\hspace{4mm} Science's Next Wave, Televisão Universitária, Uberlândia, Folha de São Paulo  
2004\hspace{4mm} _Science_ Magazine's Minority Scientists Network  
2002\hspace{4mm} NPR's Weekend Edition Saturday, BBC World Service Radio 4, Austrian Public Television's Modern Times.  

\endgroup   

<!-- text size: https://www.overleaf.com/learn/latex/Font_sizes%2C_families%2C_and_styles -->
<!-- xrfill adjusts line position and width -->
\vspace{0.2cm}
\textbf{\textcolor{darkmidnightblue}{\large{PROFESSIONAL DEVELOPMENT }}}\xrfill[0.5ex]{1pt}

<!-- ## PROFESSIONAL DEVELOPMENT ACTIVITIES   -->
\vspace{-0.2cm}
### TRAINING PROGRAMS
\begingroup 
\setlength{\leftskip}{.2cm}  
2013\hspace{4mm} University of Florida Science Communication Academy  
2004\hspace{4mm} California Academy of Sciences: The Ant Course: Taxonomy & Systematics (Costa Rica)  
1996\hspace{4mm} Organization for Tropical Studies: Tropical Plant Systematics 96-9 (Costa Rica)  
1993\hspace{4mm} University of California Education Abroad Program: Tropical Biology Quarter (Costa Rica) 

\endgroup

\vspace{-0.2cm}

### MEMBERSHIPS IN PROFESSIONAL SOCIETIES  

\begingroup 
\setlength{\leftskip}{.2cm}  
Ecological Society of America, Association for Tropical Biology & Conservation, British Ecological Society, American Association for the Advancement of Science, International Society for Scientometrics & Infometrics  

\endgroup

<!-- text size: https://www.overleaf.com/learn/latex/Font_sizes%2C_families%2C_and_styles -->
<!-- xrfill adjusts line position and width -->
\vspace{0.2cm}
\textbf{\textcolor{darkmidnightblue}{\large{LANGUAGES }}}\xrfill[0.5ex]{1pt}  

\begingroup 
\setlength{\leftskip}{.2cm}  

Fluent in Spanish, English, & Portuguese  

\endgroup
