---
title: "Curriculum Vitae"
author: "Theresa M. Laverty"
date: "`r format(Sys.time(), '%B %Y')`"

web: www.LavertyLab.org
github: theresalaverty
#twitter: BrunaLab
phone: "(609) 515-3953"
email: tlaverty@nmsu.edu

bibliography: ./emb_cv_modular/TML_publications.bib
csl: brunacv.csl

updated: yes

header-includes:
# - \usepackage{etaremune} # for reverse-numbered list of publications etc
# see https://ctan.math.utah.edu/ctan/tex-archive/macros/latex/contrib/etaremune/etaremune.pdf
- \usepackage[default]{sourcesanspro}
# - \usepackage{enumerate} #for reverse numbering
# - \usepackage{enumitem} #for reverse numbering
- \usepackage{soul} # highlights text
- \usepackage[dvipsnames]{xcolor} #allows access to other colors
- \sethlcolor{Goldenrod} # use goldenrod to match highklighting on preview
- \usepackage{fancyhdr} # modify header into left, center, right
- \pagestyle{fancy} # cues the use of fancy headers
- \fancyfoot{} # cues the use of fancy footer
- \fancyfoot[R]{`r format(Sys.time(), '%B %Y')`} # month & year on footer right
- \fancyhead[R]{Theresa M. Laverty, p. \thepage} # name and page on header right
# - \fancyhead[L]{\hl{Highlighting indicates items in the 2021-2025 PTR period}} # name and page on header left
- \usepackage{setspace} # manipulate line spacing 
- \usepackage{parskip} # manipulate indents and spacing before/after paragraphs 
- \usepackage{xcolor} # changes colors of text
- \AtBeginDocument{\let\maketitle\relax} # this removes default title section
  # to center to top level (#) headings (my name) and change color of font
  # more info # https://www.ctan.org/pkg/sectsty
- \usepackage{sectsty} \sectionfont{\centering\color{darkmidnightblue}}
# centersub-headings (##) & sub-sub-headings (###) & change font color
- \usepackage{sectsty} \subsectionfont{\color{darkmidnightblue}}
- \usepackage{sectsty} \subsubsectionfont{\color{darkmidnightblue}}
# http://latexcolor.com/ to choose colors
# define the colors used above with rgb codes
- \definecolor{darkcerulean}{rgb}{0.03, 0.27, 0.49}
- \definecolor{darkmidnightblue}{rgb}{0.0, 0.2, 0.4}
- \usepackage{xhfill} # modification of line color and width
- \pdfmapfile{=SourceSansPro.map} # this is to deal with error https://tex.stackexchange.com/questions/173204/texlive-font-installation-what-am-i-doing-wrong

geometry: margin=1in
linkcolor: darkmidnightblue
urlcolor: darkmidnightblue
fontawesome: yes
# 
output:
  pdf_document: default
  df_print: paged
  toc: no
  # word_document: default
  # html_document: null
  

---


```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = FALSE,message=FALSE,warning=FALSE)
library(tidyverse)
library(gridExtra)
library(kableExtra)
# library(citr)
library(RefManageR)

```
<!-- heading is line with centered text here: -->
<!-- https://tex.stackexchange.com/questions/15119/draw-horizontal-line-left-and-right-of-some-text-a-single-line -->
<!-- \newcommand*\ruleline[1]{\par\noindent\raisebox{.8ex}{\makebox[\linewidth]{\hrulefill\hspace{1ex}\raisebox{-.8ex}{#1}\hspace{1ex}\hrulefill}}} -->
<!-- \ruleline{\textbf{\textcolor{darkmidnightblue}{\large{RESEARCH \& TEACHING INTERESTS}}}} -->

<!-- this next line suppresses the header (line, name,  page number) on 1st page -->
<!-- \begin{document}. -->
\thispagestyle{empty}

\vspace{-0.3cm}  <!--space between name and address -->
# THERESA MARIE LAVERTY
\vspace{-0.3cm}  <!--space between name and address -->
New Mexico State University\hfill tlaverty@nmsu.edu  
Department of Fish, Wildlife and Conservation Ecology\hfill +1 (609) 515-3953  
2980 South Espina Street, Knox Hall 132\hfill&nbsp;[www.LavertyLab.org](www.lavertylab.org)  
Las Cruces, NM 88003 USA\hfill&nbsp;[Github.com/LavertyLab](https://github.com/LavertyLab)
\vspace{-0.5cm}  <!--there was too much space, so reduced it with minus: -Xcm -->
<!-- ___   -->
<!-- \par\noindent\rule{\textwidth}{0.4pt} -->
\vspace{0.2cm}  
<!-- \textbf{\textcolor{darkmidnightblue}{RESEARCH \& TEACHING INTERESTS:}} Tropical ecology and conservation, plant population ecology, plant-animal interactions, scientometrics and bibliometrics, science & science policy in Latin America   -->

<!-- text size: https://www.overleaf.com/learn/latex/Font_sizes%2C_families%2C_and_styles -->
<!-- xrfill adjusts line position and width -->
\vspace{0.1cm}
\textbf{\textcolor{darkmidnightblue}{\large{RESEARCH \& TEACHING INTERESTS }}}\xrfill[0.5ex]{1pt}

Vertebrate ecology and conservation, population and community ecology, food web ecology, dryland ecology and management, demographic modeling, bibliometics 

<!-- \textbf{\textcolor{darkmidnightblue}{LANGUAGES:}} Fluent in Spanish, English, & Portuguese   -->
<!-- \vspace{-0.6cm}   -->

<!-- ```{r, child=c('one.Rmd', 'two.Rmd')} -->
<!-- ``` -->

<!-- ```{r, child='one.Rmd'} -->
<!-- ``` -->

<!-- text size: https://www.overleaf.com/learn/latex/Font_sizes%2C_families%2C_and_styles -->
<!-- xrfill adjusts line position and width -->
\vspace{0.2cm}
\textbf{\textcolor{darkmidnightblue}{\large{EDUCATION }}}\xrfill[0.5ex]{1pt}

<!-- ### EDUCATION -->
<!--to indent a block of stuff treat it as a 'group' and then set leftskip -->
<!--be sure to indicate where the group ends with \endgroup -->

\begingroup 
\setlength{\leftskip}{.2cm} 
\setlength{\rightskip}{.2cm} 
<!-- ### EDUCATION \hrulefill\ -->
<!-- TEXT \hrulefill\ -->
<!-- Here is some text\xfill{1pt} \par -->
<!-- Here is some text\xfill{8pt} \par -->
<!-- Here is some text\xhrulefill{darkmidnightblue}{2pt}\par -->
<!-- Here is some text\xhrulefill{darkmidnightblue}{1ex}\par -->
<!-- HOW TO MAKE LINE COLOR SAME'? -->
<!-- https://tex.stackexchange.com/questions/32597/vertically-centered-horizontal-rule-filling-the-rest-of-a-line -->
<!-- \textbf{\textcolor{darkmidnightblue}{EDUCATION }}\xrfill[0.5ex]{0.5pt} -->

<!-- ### EDUCATION \hrulefill -->



**Ph.D. in Fish, Wildlife, and Conservation Biology**, Colorado State University  \hfill 2019  
_Bats, elephants, and their food: A conservation perspective on trophic interactions in the Namib Desert_ \newline  
Dissertation Advisor: Joel Berger

**B.A. in Ecology and Evolutionary Biology**, Princeton University  \hfill 2010  
_Niche overlap in the diets of three Amazonian caiman species_ \newline
Thesis Advisor: Andrew P. Dobson 

\endgroup 

<!-- ```{r ed, echo=FALSE,message = FALSE,warning=FALSE} -->

<!-- ed<-read_csv("./cv_files/education.csv") %>% -->
<!-- replace_na(list(program=" ", school=" ",year=" ",degree=" ")) -->
<!-- ed$program<-gsub('","',", ",ed$program) -->
<!-- ed$school<-gsub('","',", ",ed$school) -->
<!-- kbl(ed, -->
<!--     digits = 1, -->
<!--     format = "latex", -->
<!--     align="l", -->
<!--     col.names = NULL, -->
<!--     longtable = T, -->
<!--     row.names = FALSE, -->
<!--     booktabs=T, -->
<!--     linesep = "") %>% -->

<!--   kable_styling(bootstrap_options = c("hover"), -->
<!--                 latex_options = c("scale_down"), -->
<!--                 full_width = F, -->
<!--                 font_size = 10, -->
<!--                 position = "left") %>% -->
<!--   # row_spec(0:0, bold = T) %>% -->
<!--   # # column_spec(2, width = "35em") %>% -->
<!--   sub("\\\\toprule", "", .) %>% -->
<!--   sub("\\\\bottomrule", "", .) -->
<!-- ``` -->
<!-- text size: https://www.overleaf.com/learn/latex/Font_sizes%2C_families%2C_and_styles -->
<!-- xrfill adjusts line position and width -->
\vspace{0.2cm}
\textbf{\textcolor{darkmidnightblue}{\large{PROFESSIONAL APPOINTMENTS }}}\xrfill[0.5ex]{1pt}

<!-- ### PROFESSIONAL APPOINTMENTS -->
\begingroup 
\setlength{\leftskip}{.2cm} 
\setlength{\rightskip}{.2cm} 
\textbf{Assistant Professor}, New Mexico State University  \hfill 2023-Present  
Department of Fish, Wildlife and Conservation Ecology  

\textbf{Postdoctoral Scholar}, University of Chicago  \hfill 2021-2022  
Department of Ecology and Evolution 

\textbf{Postdoctoral Scholar}, Colorado State University  \hfill 2019-2021  
Department of Fish, Wildlife and Conservation Ecology  

\endgroup
<!-- ```{r positions, echo=FALSE,message = FALSE,warning=FALSE} -->

<!-- positions<-read_csv("./cv_files/positions.csv") %>% -->
<!-- replace_na(list(dates=" ", position=" ")) -->

<!-- # positions<-as.data.frame(positions) -->
<!-- # positions[1,1] <- cell_spec(positions[1, 1], "latex", bold = T)  -->
<!-- kbl(positions, -->
<!--     digits = 1, -->
<!--     format = "latex", -->
<!--     align="l", -->
<!--     col.names = NULL, -->
<!--     longtable = T, -->
<!--     row.names = FALSE, -->
<!--     booktabs=T, -->
<!--     linesep = "") %>% -->

<!--   kable_styling(bootstrap_options = c("hover"), -->
<!--                 latex_options = c("scale_down"), -->
<!--                 full_width = F, -->
<!--                 font_size = 10, -->
<!--                 position = "left") %>% -->
<!--   # row_spec(0:0, bold = T) %>% -->
<!--   # # column_spec(2, width = "35em") %>% -->
<!--   sub("\\\\toprule", "", .) %>% -->
<!--   sub("\\\\bottomrule", "", .) -->
<!-- ``` -->


<!-- text size: https://www.overleaf.com/learn/latex/Font_sizes%2C_families%2C_and_styles -->
<!-- xrfill adjusts line position and width -->
\vspace{0.2cm}

\textbf{\textcolor{darkmidnightblue}{\large{AWARDS \& FELLOWSHIPS }}}\xrfill[0.5ex]{1pt}
\vspace{-0.4cm}

### AWARDS  
\vspace{-0.2cm}  
\begingroup 
\setlength{\leftskip}{.3cm} 
\setlength{\rightskip}{.2cm} 
Outstanding Departmental Publication Award, Colorado State University \hfill 2021 \newline
for the 2021 _Ecological Applications_ article   

\endgroup

<!-- &nbsp;\quad\quad\quad\quad _for the year's outstanding contribution to the field, published in any journal_   -->
<!-- &nbsp;\quad\quad\quad\quad _for the outstanding paper published in the journal Biotropica_   -->
<!-- &nbsp;\quad\quad\quad\quad _for the outstanding paper by a young investigator in the Journal of Ecology_ -->

<!-- \endgroup -->
\vspace{-0.2cm}
<!-- \begingroup  -->
<!-- \setlength{\leftskip}{.2cm}   -->
### FELLOWSHIPS
\vspace{-0.2cm}
\begingroup 
\setlength{\leftskip}{.3cm} 
\setlength{\rightskip}{.2cm} 
National Science Foundation Graduate Research Fellowship \hfill 2013-2018  

\endgroup

<!-- text size: https://www.overleaf.com/learn/latex/Font_sizes%2C_families%2C_and_styles -->
<!-- xrfill adjusts line position and width -->
\vspace{0.2cm}
\textbf{\textcolor{darkmidnightblue}{\large{GRANTS IN SUPPORT OF RESEARCH, TEACHING, \& TRAINING }}}\xrfill[0.5ex]{1pt}
<!-- ## GRANTS IN SUPPORT OF RESEARCH, TRAINING, & PROGRAM DEVELOPMENT (>$5.5 million)   -->

\begingroup 
\setlength{\leftskip}{.2cm}  

1. New Mexico Department of Wildlife. _"Acoustic surveys of bat species of greatest conservation need within wind energy-impacted landscapes of New Mexico"_ ($114,066; PI: TM Laverty; 2027-2029).

1. National Science Foundation. _"LTER: Advancing ecological understanding of drylands through integrated research at the Jornada Basin (JRN-8)"_ ($7,649,997 [3% credit]; PI: NP Hanan, Co-PIs: BT Bestelmeyer, S Bestelmeyer, DM Browning, F Garcia-Pichel, TM Laverty, GE Maurer, ME Mauritz-Tozer, GS Okin, N Pietrasiak, SC Reed, AL Romero-Olivares, OE Sala, RL Schooley, ER Vivoni; 2025-2031).

1. Arizona Game and Fish Department. _"Movement ecology of bats with Motus"_ ($100,000; PI: TM Laverty; 2025).

1. Bureau of Land Management. _"Identifying barriers to pronghorn (Antilocapra americana) habitat connectivity in southwestern New Mexico in relation to rangeland management and renewable energy development"_ ($502,843 [30% credit]; PI: JW Cain, Co-Pis: JK Frey, TM Laverty; 2024-2028).

1. National Science Foundation. _"BRC-BIO: The interactive effects of bats and insects on shrub encroachment dynamics in the Chihuahuan Desert"_ ($497,568; PI: TM Laverty, 2024-2027).

2. National Park Service. _"Determining activity patterns within and seasonal movements between protected areas in the southwest borderlands for bats of greatest conservation need"_ ($54,884; PI: TM Laverty; 2024-2026; with an additional $42,500 awarded in 2025).

3. Bureau of Land Management. _"Influence of Wind Energy Development on Medium and Large Mammals"_ ($248,609 [50% credit]; PI: JK Frey, Co-PI: TM Laverty, 2023-2025; with an additional $5,800 awarded in 2024).

3. National Geographic Society. _"Coding for conservation: Quantitative training for the next generation of environmental leaders in Madagascar"_ ($25,000; PI: CE Brook, Team Members: AF Andrianiaina, E Cornelius Ruhs, TM Laverty, O Razafindramatsima; 2022-2023).

3. New Mexico Department of Game and Fish. _"Seasonal abundance, population structure, and diet of long-nosed bats in southwestern New Mexico in relation to contemporaneous food availability"_ ($49,839; PI: K Stoner, Co-PI: TM Laverty; 2020-2022).

4. Grants for Graduate Research: National Science Foundation Graduate Research Fellowship, Bat Conservation International Student Research Scholarship for Global Bat Conservation Priorities, National Geographic Society Young Explorers Grant, Sigma Xi Grant-In-Aid of Research, The Explorers Club Exploration Fund, University of Montana (Fred H. Mass Scholarship, Danny On Memorial Scholarship, George and Mildred Cirica Graduate Student Support Fund [declined due to university transfer]), and Colorado State University (Eugene Decker Fellowship, Hill Memorial Fellowship, WCNR Student Success Scholarship) (Total: $149,450; 2013-2019).

5. Grants for Undergraduate Research: Princeton University (John T. Bonner Senior Thesis Fund and the Dean of the College Fund for undergraduate senior thesis research) (Total: $7,800; 2009).

\endgroup

<!-- ### GRANT PROPOSALS UNDER REVIEW or IN PREPARATION -->

<!-- 1. UF 10x10x10 Initiative: The "interconnected future of Florida & Latin America: Entrepreneurship, the Environment, & Transnational Migration" (EM Bruna and R. Stepp, UF Center for Latin American Studies). -->

<!-- 1. National Science Foundation. "The genomic basis and demographic consequences of male-biased local adaptation in a moss" ($2,267,331; PI: S. McDaniel, Co-PIs: K. Krippen, EM Bruna, G. John, B. P. Durham).  -->


<!-- --- -->
<!-- nocite: '@*' -->
<!-- ... -->

<!-- text size: https://www.overleaf.com/learn/latex/Font_sizes%2C_families%2C_and_styles -->
<!-- xrfill adjusts line position and width -->
\vspace{0.2cm}
\textbf{\textcolor{darkmidnightblue}{\large{PUBLICATIONS \& SCHOLARLY COMMUNICATION  }}}\xrfill[0.5ex]{1pt}
<!-- ## PUBLICATIONS & SCHOLARLY COMMUNICATION -->
<!-- \vspace{-0.4cm} -->

**_Lab personnel: ^GS^Graduate Student, ^UG^Undergraduate Student; ^*^Co-First Author_** 
<!-- ^PD^Postdoctoral Researcher, ^VS^Visiting Student -->

\vspace{-0.4cm}
### PREPRINTS AND MANUSCRIPTS IN REVIEW or IN REVISION 

<!-- \begin{enumerate}[{start=7,label=[\arabic*]\addtocounter{enumi}{-2}}] -->

<!-- \item  -->
1. Cooper MC, LaRue ZBT, Boeing WJ, Laverty TM, Fernald AG, Geli HME, Jones KW. Water conservation attitudes and behaviors in the arid southwestern United States.\textit{In review}, \textit{\textbf{Sustainable Water Resources Management}}.

1. ^GS^Rohan IM, Laverty TM, Frey JK. Terrestrial mammals alter habitat selection in response to turbine visibility at a central New Mexico wind energy facility. \textit{In revision}, \textit{\textbf{Conservation Science and Practice}}.


<!-- \end{enumerate}   -->

<!-- very helpful for bibliography: https://inbo.github.io/tutorials/tutorials/r_citations_markdown/ -->
<!-- maybe to print out by reference groups? -->
<!-- https://forums.zotero.org/discussion/23604/a-mini-guide-to-sort-bibliography-by-reference-type -->
<!-- \endgroup -->

<!-- TODO: -->
<!-- 1) indent the automated refs -->
<!-- 2) put field in zotero for noes (eg theis chapter, recipient of..., etc), -->
<!-- 3) make the extra field with notes bold for cv -->
<!-- 3) find out which refs ones are missing -->



<!-- ### PEER-REVIEWED PUBLICATIONS (AUTOMATED) -->

<!-- <div id="refs"></div> # this allows you to put text after -->
<!-- --- -->
<!-- nocite: '@*' # this puts eveything in .bib without needing to ccite it -->
<!-- --- -->

 
### PEER-REVIEWED PUBLICATIONS \& BOOK CHAPTERS

1. Montauban C, Budinski I, Webala PW, Laverty TM, Tanshi I, Torrent L, Bakwo-Fils E, Taylor PJ, Kane A, Monadjem A. 2026. Underrepresentation of bats in Africa’s protected areas. \textit{\textbf{Conservation Biology}} 40(1):e70108. https://doi.org/10.1111/cobi.70108

2. Andrianiaina AF, Andry S, Kettenburg G, Ranaivoson HC, Lacoste V, Dussart P, Heraud J-M, Laverty TM, Guth S, Young KI, Andrianarimisa A, Brook CE. 2025. Diversity and seasonality of ectoparasite burden on two species of Madagascar fruit bat, *Eidolon dupreanum* and *Rousettus madagascariensis*. \textit{\textbf{Parasites \& Vectors}} 18(1):302. https://doi.org/10.1186/s13071-025-06805-z **_Data & Code:_** [GitHub](https://github.com/brooklabteam/Mada-Ectoparasites)

3. Waag AG, Johnson JS, Schorr RA, Frick WF, Laverty TM, Neubaum DJ, Siemers JL, Treanor JJ, Reyes GA, Halstead BJ. 2025. Using passive integrated transponder (PIT) tags to monitor and research bats: applications, benefits, and limitations. \textit{\textbf{Journal of North American Bat Research}} Special Issue 1:100–118. https://www.eaglehill.us/nabr-pdfs-special/nabr-008-Waag.pdf

4. Laverty TM, Buecher DC, Dalton DC, Davies ML, Gual-Suárez F, Heady III PA, Medellín RA, Ordóñez-García MC, Pérez-Harp S, Rivera-Villanueva AN, ^GS^Stevenson BD, Stoner KE, Wolf SA, Zamora-Gutierrez V, Zavala-Norzagaray AA, Frick WF. 2025. Long-haul flights and migratory routes of a nectar-feeding bat. \textit{\textbf{Ecology}} 106(4):e70067. https://doi.org/10.1002/ecy.70067 **_Data & Code:_** [Zenodo](https://doi.org/10.5281/zenodo.14853138)

5. Laverty TM, Buecher DC, Dalton DC, Davies ML, Gual-Suárez F, Heady III PA, Medellín RA, Ordóñez-García MC, Pérez-Harp S, Rivera-Villanueva AN, ^GS^Stevenson BD, Stoner KE, Wolf SA, Zamora-Gutierrez V, Zavala-Norzagaray AA, Frick WF. 2025. On the move: documenting long-distance flights of a nectar-feeding bat. \textit{\textbf{The Bulletin of the Ecological Society of America}} 106(3):e70016. https://doi.org/10.1002/bes2.70016

6. Monadjem A, Montauban C, Webala PW, Laverty TM, Bakwo-Fils EM, Torrent L, Tanshi I, Kane A, Rutrough AL, Waldien DL, Taylor PJ. 2024. African bat database: curated data of occurrences, distributions and conservation metrics for sub-Saharan bats. \textit{\textbf{Scientific Data}} 11(1):1309. https://doi.org/10.1038/s41597-024-04170-7 **_Data & Code:_** [figshare](https://doi.org/10.6084/m9.figshare.26363308)

7. ^*^Nelson-Olivieri JR, ^*^Layden TJ, Antunez E, Khalighifar A, Lasky M, Laverty TM, Sanchez KA, Shannon G, Starr S, Verahrami AK, Bombaci SP. 2024. Inequalities in noise will affect urban wildlife. \textit{\textbf{Nature Ecology \& Evolution}} 8(1):163–174. https://doi.org/10.1038/s41559-023-02257-9 **_Data:_** [Dryad](https://doi.org/10.5061/dryad.s4mw6m998) **_Code:_** [Zenodo](https://doi.org/10.5281/zenodo.7843664)

8. Wilkins K, Carroll SL, Davis KP, Hauptfeld R, Jones MS, Larson CL, Laverty TM, Pejchar L. 2023. Sexual harassment disproportionately affects ecology and evolution graduate students with multiple marginalized identities in the United States. \textit{\textbf{BioScience}} 73(5):376–387. https://doi.org/10.1093/biosci/biad032

9. Laverty TM, Stoner KE. 2022. In search of bachelorettes: observations of male *Leptonycteris yerbabuenae* with dorsal patches across its range. \textit{\textbf{Therya}} 13(2):163–170. https://doi.org/10.12933/therya-22-1146 **_Media:_** [figshare](https://doi.org/10.6084/m9.figshare.15138702)

10. Mushabati LM, Eiseb SJ, Benda P, Laverty TM. 2022. Effects of lunar phase and temperature on bat activity and species richness at varying altitudes in the Kunene Region, Namibia. \textit{\textbf{African Journal of Ecology}} 60(3):467–480. https://doi.org/10.1111/aje.12968

11. Laverty TM, Berger J. 2022. Indirect effects of African megaherbivore conservation on bat diversity in the world’s oldest desert. \textit{\textbf{Conservation Biology}} 36(2):e13780. https://doi.org/10.1111/cobi.13780 **_Data & Code:_** [figshare](https://doi.org/10.6084/m9.figshare.13376975)

12. Laverty TM, Teel TL, Gawusab AA, Berger J. 2021. Listening to bats: Namibian pastoralists’ perspectives, stories, and experiences. \textit{\textbf{Journal of Ethnobiology}} 41(1):70–86. https://doi.org/10.2993/0278-0771-41.1.70

13. Mans BJ, Kelava S, Pienaar R, Featherston J, de Castro MH, Quetglas J, Reeves WK, Durden LA, Miller MM, Laverty TM, Shao R, Takano A, Kawabata H, Moustafa MAM, Nakao R, Matsuno K, Greay TL, Evasco KL, Barker D, Barker SC.. 2021. Nuclear (18S-28S rRNA) and mitochondrial genome markers of *Carios* (*Carios*) *vespertilionis* (Argasidae) support *Carios* Latreille, 1796 as a lineage embedded in the Ornithodorinae: re-classification of the *Carios* sensu Klompen and Oliver (1993) clade into its respective subgenera. \textit{\textbf{Ticks and Tick-borne Diseases}} 12(4):101688. https://doi.org/10.1016/j.ttbdis.2021.101688

14. ^*^Aubry LM, ^*^Laverty TM, ^*^Ma Z. 2021. Impacts of COVID-19 on ecology and evolutionary biology faculty in the United States. \textit{\textbf{Ecological Applications}} 31(2):e2265. https://doi.org/10.1002/eap.2265

15. Orlova MV, Laverty TM, Reeves WK, ^UG^Gratton EM, Davies ML, Anisimov NV. 2020. New geographical and host record of bat ectoparasite *Steatonyssus* (*Steatonyssus*) *afe*r Radovsky and Yunker, 1963 (Mesostigmata: Gamasina: Macronyssidae). \textit{\textbf{Acarologia}} 60(4):951–958. https://doi.org/10.24349/acarologia/20204411

16. Reeves WK, Laverty TM, ^UG^Gratton EM, Mushabati LM, Eiseb SJ. 2020. New national records for *Cyclopodia greeffi greeffi* (Diptera: Nycteribiidae) from the Kunene Region, Namibia, Africa. \textit{\textbf{Entomological News}} 129(3):327–329. https://doi.org/10.3157/021.129.0311

17. Reeves WK, Mans BJ, Durden LA, Miller MM, ^UG^Gratton EM, Laverty TM. 2020. *Rickettsia hoogstraalii* and a Rickettsiella from the Bat Tick *Argas transgariepinus*, in Namibia. \textit{\textbf{Journal of Parasitology}} 106(5):663–669. https://doi.org/10.1645/20-46

18. Laverty TM, Berger J. 2020. Do bats seek clean water? A perspective on biodiversity from the Namib Desert. \textit{\textbf{Biological Conservation}} 248:108686. https://doi.org/10.1016/j.biocon.2020.108686 **_Data & Code:_** [Mendeley Data](https://doi.org/10.17632/dz6fyzrj5p.1)

18. Laverty TM. 2020. The Namib Desert and its bats. In: Goldstein MI, DellaSala DA, editors. *Encyclopedia of the world’s biomes, vol. 2*. Oxford: Elsevier; p. 181–192. https://doi.org/10.1016/B978-0-12-409548-9.12011-1

19. Orlova MV, Laverty TM, Reeves WK, ^UG^Gratton EM, Davies ML. 2020. The first finding of parasitic mite, *Parasteatonyssus nyctinomi* (Mesostigmata: Gamasina: Macronyssidae), in Namibia. \textit{\textbf{Journal of the South African Veterinary Association}} 91(1):a2002. https://doi.org/10.4102/jsava.v91i0.2002

20. Orlova MV, Laverty TM, Reeves WK, ^UG^Gratton EM, Davies ML. 2020. The first record of the spinturnicid mite *Spinturnix kolenatii* Oudemans, 1910 (Mesostigmata: Gamasina: Spinturnicidae) from the long-tailed serotine bat *Eptesicus hottentotus* A. Smith, 1833 (Chiroptera: Vespertilionidae) in Africa. \textit{\textbf{International Journal of Acarology}} 46(3):160–164. https://doi.org/10.1080/01647954.2020.1731596

21. ^*^Jimenez MF, ^*^Laverty TM, Bombaci SP, Wilkins K, Bennett DE, Pejchar L. 2019. Underrepresented faculty play a disproportionate role in advancing diversity and inclusion. \textit{\textbf{Nature Ecology \& Evolution}} 3(7):1030–1033. https://doi.org/10.1038/s41559-019-0911-5

22. Laverty TM, Teel TL, Thomas REW, Gawusab AA, Berger J. 2019. Using pastoral ideology to understand human-wildlife coexistence in arid agricultural landscapes. \textit{\textbf{Conservation Science and Practice}} 1(5):e35. https://doi.org/10.1111/csp2.35

23. Moldowan PD, Laverty TM, Emmans CJ, Stanley RC. 2016. Diet, gastric parasitism, and injuries of caimans (*Caiman*, _Melanosuchus_, and _Paleosuchus_) in the Peruvian Amazon. \textit{\textbf{South American Journal of Herpetology}} 11(3):176–182. https://doi.org/10.2994/SAJH-D-16-00012.1

24. Laverty TM, Dobson AP. 2013. Dietary overlap between black caimans and spectacled caimans in the Peruvian Amazon. \textit{\textbf{Herpetologica}} 69(1):91–101. https://doi.org/10.1655/HERPETOLOGICA-D-12-00031


<!-- ### ANNOUNCEMENTS, BOOK REVIEWS, \& CORRESPONDENCE -->

<!-- 1. Bruna, EM, Powers, JS and Lohmann, LG 2022. 2022 ATBC Honorary Fellows. **_Biotropica_** 54:1521-1522. https://doi.org/10.1111/btp.13170 [Essay]   -->

<!-- 1.	KR Zamudio, A Kellner, C Serejo, MR de Britto, CB Castro, PA Buckup, DO Pires, M Couri, A Brilhante Kury, I Azevedo Cardoso, ML Monné, J Pombal Jr, C Mello Patiu, V Padula, A Dias Pimenta, R Ventura, E  Hajdu, J Zanol, EM Bruna, J Fitzpatrick, LA Rocha. 2018. Lack of support for Science and Education Fails Brazilian Society - Again. **_Science_** 361(6409): 1322-1323. [Correspondence]  -->

<!-- 1.	Bruna, EM. 2017. Invasive new US visa rules hurt Americans too. **_Americas Quarterly_**. http://www.americasquarterly.org/content/academic-collaboration [Invited essay]   -->

<!-- 1.	Harms, K, J Benitez-Malvido, \& EM Bruna. 2017. 2016 Alwyn Gentry Awards and Luis F. Bacardi Award Recipients. \textit{\textbf{Biotropica}} [Essay].   -->

<!-- 1.	Bruna, EM 2016. Biotropica renames its "Outstanding Paper" awards for two pioneering scientists. \textit{\textbf{Biotropica}} 48(6): 929 [Essay].   -->

<!-- 1.	EM Bruna, M. Peña-Claros, Bryan Finega, and B. A. Kaplin. 2016. A New Data Archiving Policy for _Biotropica._ \textit{\textbf{Biotropica}} 48(2): 139-140. [Editorial].   -->

<!-- 1.	V. Christianini, P. S. Oliveira, EM Bruna, H. L. Vasconcelos. 2014. Fauna in decline: Meek shall inherit. \textit{\textbf{Science}} 354(6201):1129. [Correspondence]   -->

<!-- 1.	Bruna, EM. 2014. Biotropica to become an 'online-only' journal in 2015. \textit{\textbf{Biotropica}} 46(6): 782. [Editorial].   -->

<!-- 1.	Bruna, EM. 2014. 2014 Alwyn Gentry Awards, Luis F. Bacardi Award, and ATBC Honorary Fellows. \textit{\textbf{Biotropica}} 46(6): 779-781. [Essay].   -->

<!-- 1.	Bruna, EM. 2014. A word from the Editor-in-Chief. \textit{\textbf{Biotropica}} 46(4): 377 [Editorial]. -->

<!-- 1.	Bruna, EM 2010.  Announcement: _Biotropica_ Award for Excellence in Tropical Biology and Conservation. \textit{\textbf{Biotropica}} 42(4): 518-519 [Essay].   -->

<!-- 1.	Bruna, EM  2003.  "For the Love of Nature".  **_Science: Next Wave Online Magazine_**. [Invited essay]. -->

<!-- 1.	Bruna, EM 2002. A biologist's guide to fragmenting the Amazon (including why it's a bad idea).  **_Conservation Biology_**, 16(5): 1446-1448. [Book review].   -->


<!-- ### SOFTWARE \& TECHNICAL MANUALS -->

<!-- 1. Fournier, A., F. Stevens, M. Boone, and EM Bruna. 2025. refsplitr. author name disambiguation, author georeferencing, and mapping of coauthorship networks with Web of Science data. R package version 1.2.0. https://docs.ropensci.org/refsplitr _(Package maintained by EM Bruna)_ -->

<!-- 1. HerbVar Steering Committee 2024. HerbVar: Project Manual and Field protocols (v0.9.0). Zenodo. https://doi.org/10.5281/zenodo.14232308 -->


<!-- ### SCHOLARLY PRODUCTIVITY in ONLINE OUTLETS -->

<!-- 1.	Johnson-Ulrich, L and EM Bruna. 2019. Emilio Bruna: Seed germination is lower in fragmented rainforest. **_Project Biodiversify_** [projectbiodiversify.org/emilio-bruna/](projectbiodiversify.org/emilio-bruna/) [Teaching materials] -->

<!-- 1. _Jean Theodore Descourtilz (1796-1855) and the Ornithologie Bresilienne_: Website created with graduate students in response to request from the UF Harn Museum of Art to analyze an aqcuisition of chromolithographs https://descourtilz.wordpress.com/ -->

<!-- 1. Over 30 Wikipedia pages edited with students as part of graduate and undergraduate courses in Plant Animal Interactions and Tropical Ecology. _Complete list at http://en.wikipedia.org/wiki/User:Embruna _    -->

<!-- ### OTHER PUBLICLY ARCHIVED DATASETS \& CODE  -->
<!-- 1. Heliconia acuminata Demographic Data:   -->

<!-- > Seed set: http://dx.doi.org/10.6084/m9.figshare.1273926   -->
<!-- > Leaf Length vs. Area: http://dx.doi.org/10.6084/m9.figshare.92691   -->
<!-- > Leaf number, leaf area, shoot number, and height of reproductive _H. acuminata_: https://zenodo.org/records/5041931   -->

<!-- 1. Emilio M. Bruna, Audrey C. Smith, Leandra Merz, Jesse Borden, Chris K. Gulick, \& Akhil R. Kshirsagar. (2021). Data and Code from: Assessing the effect of article processing charges on the geographic diversity of authors using Elsevier's 'Mirror Journal' system. (v1.0.0). Zenodo. https://doi.org/10.5281/zenodo.5500293 -->

<!-- 2. Costa et al. (2018) Data from: The relative influence of an environmental gradient and an ecosystem engineer on seedling communities in a Neotropical savanna. Dryad Digital Repository https://doi:10.5061/dryad.vf1gq0q  -->

<!-- 2.	Emilio M. Bruna. (2018, September 13). embruna/Costa-etal-PeerJ-2018: First release of code used in Costa et al 2018 PeerJ (Version v1.0.0). Peerj. Zenodo. http://doi.org/10.5281/zenodo.1418482  -->

<!-- 3.	Espin, J., et al. (2017) Data from: A persistent lack of international representation on editorial boards in environmental biology. Dryad Digital Repository. https://doi.org/10.5061/dryad.mh189 -->

<!-- 4.	Bruna, EM 2017. Code (v1.0) used in 'A persistent lack of international representation on editorial boards in environmental biology, Espin et al.' Zenodo http://doi.org/10.5281/zenodo.1067800 -->

<!-- 5.	Vieira-Neto EHM, Vasconcelos HL, Bruna EM (2016) Data from: Roads increase population growth rates of a native leaf-cutter ant in Neotropical savannahs. Journal of Applied Ecology https://doi.org/10.5061/dryad.1s4f2 -->

<!-- 6.	Costa AN, Vasconcelos HL, Bruna EM (2016) Data from: Biotic-drivers of seedling establishment in Neotropical savannas: selective granivory and seedling herbivory by leaf-cutter ants as an ecological filter. Journal of Ecology https://doi.org/10.5061/dryad.1hj56 -->

<!-- 7.	Mundim FM, Bruna EM (2016) Data from: Is there a temperate bias in our understanding of how climate change will alter plant-herbivore interactions? A meta-analysis of experimental studies. The American Naturalist https://doi.org/10.5061/dryad.dn048 -->

<!-- 8.	Bruna EM, Izzo TJ, Inouye BD, Vasconcelos HL (2014) Data from: Effect of mutualist partner identity on plant demography. Ecology https://doi.org/10.5061/dryad.5kc45 -->

<!-- 9.	Bruna E.M. 2014. Code (v1.0) used in "Effect of mutualist partner identity on plant demography. Ecology". Zenodo: doi:10.5281/zenodo.11650. Available for improvement at https://github.com/embruna/ant-plant-demography.  -->

<!-- 10.	Cho AH, Johnson SA, Schuman CE, Adler JM, Gonzalez O, Graves SJ, Huebner JR, Marchant DB, Rifai SW, Skinner I, Bruna EM (2014) Data from: Women are underrepresented on the editorial boards of journals in environmental biology and natural resource management. PeerJ _Data:_ https://doi.org/10.5061/dryad.6jn86  -->

<!-- 11.	Cho et al. 2014. Code used in "Women are underrepresented on the editorial boards of journals in environmental biology and natural resource management. PeerJ":: https://github.com/embruna/Editorial-Board-Gender  -->

<!-- 12.	Corrêa Côrtes M. et al. 2013. Low plant density enhances gene dispersal in the Amazonian understory herb Heliconia acuminata. Molecular Ecology Dryad Digital Repository:  https://doi.org/10.5061/dryad.b003f -->

<!-- 13.	Passmore, H. A et al. 2012. Data from: Resilient networks of ant-plant mutualists in Amazonian forest fragments. PLoS ONE.  https://doi.org/10.5061/dryad.pf2r0  -->

<!-- 14.	Copeland, S. M. et al. 2012. Data from: Effects of experimentally elevated precipitation and nitrogen on soil fertility and plant growth in a Neotropical savanna. Ecosphere. Dryad Digital Repository: https://doi.org/10.5061/dryad.dg380p9q -->

<!-- 15.	Bruna EM, et al. 2011. Data from: Asymmetric dispersal and colonization success of Amazonian plant-ants queens. PLoS ONE. Dryad Digital Repository. https://doi.org/10.5061/dryad.h6t7g  -->

<!-- 16.	Bruna, EM and A. Segalin de Andrade. 2011. Data from: Phenotypic plasticity in plant responses to edge effects: experimental test with an Amazonian understory herb. American Journal of Botany. Dryad Digital Repository: https://doi.org/10.5061/dryad.553hc134 -->

<!-- 17.	Microsatellite sequences from Cortes, MC et al. 2009. Characterization of 10 microsatellite markers for the understorey Amazonian herb Heliconia acuminata. Molecular Ecology Resources. Genbank. -->

<!-- 18.	Stocks G. et al. 2008. Data from: The geographical and institutional distribution of ecological research in the tropics. Biotropica. Dryad Digital Repository: htttps://doi.org/10.5061/dryad.9097  -->

<!-- 19.	Sequences used in Mitochondrial DNA sequences from Bruna, EM et al. 1996. Morphological and genetic evolution appear decoupled in Pacific skinks (Emoia). Proc. Roy. Soc. Lond. B: Genbank. -->



<!-- ### MANUSCRIPTS IN PREPARATION (N=11, drafts available upon request).  -->

<!-- 1. Mundim,F. M. E. Vieira-Neto, G. F. Schneider, H. Alborn, EM Bruna. _In preparation._ Simultaneous and Sequential herbivory dynamics: exploring plant defenses strategies amidst drought challenges. **_Ecology_** -->

<!-- 1.	E. S. Parsons and EM Bruna. Assessing the magnitude of abiotic edge effects in fragmented landscapes: revisiting the paradigm with data. In preparation for **_Biological Conservation_**[MA Thesis chapter]. -->

<!-- 1. Kortessis, N., R. J. Fletcher, EM Bruna, T. A. H. Smith, and R. D. Holt. Habitat degradation, dispersal, and matrix effects: Interactions between processes at different spatial scales determine species persistence. In preparation for **_The American Naturalist_**. -->

<!-- 1. Smith, T. A. H., EM Bruna, and R. J. Fletcher. Autotomy in a changing world: How landscape structure alters the prevalence and consequences of limb loss. In preparation for **_Functional Ecology_** -->

<!-- 1.	E. H. M. Vieira-Neto, F. M. Mundim, and EM Bruna. Closed and open demographic models give contrasting patterns of inter-population dynamics in a long-lived superorganism. In preparation for **_Ecology_**. -->

<!-- 1.	Goring, S., K.S. Whitney, EM Bruna, A. L. Jacob, and T. Poisot. Ten Simple Rules to Make Scientific Content More Accessible. Preprint at: DOI: [10.22541/au.150844289.92609826](https://doi.org/10.22541/au.150844289.92609826) -->

<!-- 1.	Tolentino,M., M. Anciães, EM Bruna, and M. Uriarte. **_In revision*. Local and landscape factors mediate reproductive success of _Heliconia acuminata_ in a fragmented landscape. **_Landscape Ecology_**.  -->

<!-- 1.	Corrêa Côrtes, M., M. Uriarte, W. J. Kress, J. B. Illian, and EM Bruna. *In revision*. Why are the spatial distribution of individual plants and genotypes decoupled? A framework to assess the contribution of ecological interactions and environmental filtering. **_Ecology Letters_** [Ph.D. Thesis Chapter]. -->

<!-- 1.	Bruna, EM and W. Hahn. In preparation. Is Latin American ecology really on the rise? Cross-national patterns of productivity and collaboration by Latin American scientists. In preparation for **_PLoS Biology_**. -->

<!-- 1.	Rubim, P., D. M. S. Matos, M. Uriarte, and EM Bruna. Local density influences the growth of seedlings of an Amazonian understory herb. In preparation for **_Biotropica_** [Ph.D. Thesis Chapter]. -->

<!-- 1.	Loiselle, B. A., EM Bruna, R. Buschbacher, and S. Athayde. In prep. How can Brazilian Studies and partnerships with Brazilian institutions shape the perception of Tropical Conservation & Development issues at US universities? Lessons from the University of Florida. Brazilian Studies in the United States: The road ahead. Yale University, FGV, and The Ministry of Education in Brazil. -->

<!-- 1.	Rubim, P. R., M. Uriarte, O. Ferreira, and EM Bruna. Neighborhood effects and the reproductive effort of a tropical perennial herb. In preparation for **_Ecology_** [Ph.D. Thesis Chapter]. -->


<!-- ### WEB PAGES AND OTHER FORMS OF DIGITAL COMMUNICATION -->

<!-- 1. Over 30 Wikipedia pages edited with students as part of graduate and undergraduate courses in Plant Animal Interactions and Tropical Ecology. Complete at: http://en.wikipedia.org/wiki/User:Embruna -->

<!-- 2. Biotropica.org: The Editor"s Blog. A forum for sharing information about the journal Biotropica, its ongoing initiatives, announcements debates on recent publications, and advice for authors. >80,000 views since launched in August 2013, http://biotropica.org/ -->

<!-- 3. BrunaLab.org / -Between the 23s- (lab web page and blog): In addition to hosting publications, data, and other information on current lab activities, we provide resources for the broader community on topics (e.g., resources for beginners on making figures in R or version control with GitHub). We also maintain a compendium on smartphone applications for fieldwork, which is responsible for most of the sites 113,000 page views since January 2013. http://brunalab.org/ -->

<!-- 4. -Jean Theodore Descourtilz (1796-1855) and the Ornithologie Bresilienne.- A website created by TCD graduate students in my Tropical Ecology class to provide background information and an analysis of artwork acquired by the Harn Museum of Art on their behalf. https://descourtilz.wordpress.com/ -->


<!-- text size: https://www.overleaf.com/learn/latex/Font_sizes%2C_families%2C_and_styles -->
<!-- xrfill adjusts line position and width -->
\vspace{0.2cm}
\textbf{\textcolor{darkmidnightblue}{\large{COURSE DEVELOPMENT \& TEACHING EXPERIENCE }}}\xrfill[0.5ex]{1pt}

<!-- ## COURSE DEVELOPMENT & TEACHING EXPERIENCE -->
\vspace{-0.4cm}

### Undergraduate Courses, New Mexico State University
\vspace{-0.2cm}
\begingroup 
\setlength{\leftskip}{.2cm}  
FWCE 330: Natural History of the Vertebrates (2024-2026)  
FWCE 355: Wildlife Management and Analysis (2025)  
FWCE 431: Mammalogy (2024, 2026)

\endgroup 

\vspace{-0.2cm}
### Graduate Courses, New Mexico State University
\vspace{-0.2cm}
\begingroup 
\setlength{\leftskip}{.2cm}  
FWCE 531: Mammalogy (2024, 2026)  
Intro to GitHub Workshop (2023)  

\endgroup 
\vspace{-0.2cm}
### Guest Lectures & Participation in Field Courses
\vspace{-0.2cm}
\begingroup 
\setlength{\leftskip}{.2cm}  
**New Mexico State University:** FWCE 464: Management of Aquatic and Terrestrial Ecosystems (2026), FWCE 330: Natural History of the Vertebrates (2023)  
**University of British Columbia, Okanagan:** BIOL_O 370: African Savannah Biology (2025)  
**Colorado State University:** NR 300: Biological Diversity (2018, 2019, 2020), BZ 330: Mammalogy (2020), FW 455: Principles of Conservation Biology (2020), FW 370: Design of Fish and Wildlife Projects (2018), FW 104: Wildlife Ecology and Conservation (2017)  
**University of Montana:** WILD 480: The Upshot: Applied Wildlife Management (2014, 2015), BIOO 101N: Montana Wildlife (2013)  
**Princeton University:** EEB 380: Ecology and Conservation of African Landscapes field course (2012)
 

\endgroup 
\vspace{-0.2cm}
### Teaching Assistant
\vspace{-0.2cm}
\begingroup 
\setlength{\leftskip}{.2cm}
**Colorado State University:** NR 300: Biological Diversity (2018, 2019), FW 370: Design of Fish and Wildlife Projects (2018)  
**University of Montana:** WILD 480: The Upshot: Applied Wildlife Management (2014), BIOO 101N: Montana Wildlife (2013)  
**Princeton University:** EEB 380: Ecology and Conservation of African Landscapes field course (2012)

\endgroup 

<!-- THIS WILL PUT THINGS CENTERED WITH LINE AROUND -->
<!-- \newcommand*\ruleline[1]{\par\noindent\raisebox{.8ex}{\makebox[\linewidth]{\hrulefill\hspace{1ex}\raisebox{-.8ex}{#1}\hspace{1ex}\hrulefill}}} -->
<!-- \ruleline{\textbf{\textcolor{darkmidnightblue}{STUDENT AND POSTDOCTORAL MENTORSHIP}}}    -->

<!-- THIS WILL PUT THINGS LEFT WITH LINE EXTENDING RIGHT -->
<!-- Raised Rule Command: -->
<!-- Arg 1 (Optional) - How high to raise the rule -->
<!-- Arg 2            - Thickness of the rule -->
<!-- \newcommand{\raisedrule}[2][0em]{\leaders\hbox{\rule[#1]{1pt}{#2}}\hfill}   -->
<!-- <!-- into your document's preface, and then you can accomplish what you were hoping to with: --> 

<!-- text size: https://www.overleaf.com/learn/latex/Font_sizes%2C_families%2C_and_styles -->
<!-- xrfill adjusts line position and width -->
\vspace{0.2cm}
\textbf{\textcolor{darkmidnightblue}{\large{STUDENT MENTORSHIP }}}\xrfill[0.5ex]{1pt}

\vspace{-0.4cm}

<!-- ## STUDENT & POSTDOCTORAL MENTORSHIP -->

### GRADUATE STUDENTS ADVISED
```{r GradStudentsA, echo=FALSE,message = FALSE,warning=FALSE}
# library(flextable)
# https://cran.r-project.org/web/packages/flextable/vignettes/format.html#borders
# df <- grad_advised<-read_csv("./cv_files/grad_advised.csv") %>% filter(role=="advisor") %>% select(-role)
# table1<-df %>% regulartable() %>% autofit() %>% 
# width(j=~Degree,width=.3) %>% width(j=~Year,width=.9) %>% width(j=~Instititution,width=.1)
# table1 <- fontsize(table1, part = "header", size = 11) 
# table1 <- border_remove(x = table1)
# table1
```
\vspace{-0.2cm}
```{r GradStudents, echo=FALSE,message = FALSE,warning=FALSE}
grad_advised<-read_csv("./cv_files/grad_advised.csv") %>% filter(role=="advisor") %>% select(-role) %>%
  mutate(
    Student = if_else(
      !is.na(coadvisor) & coadvisor != "",
      str_c("\\makecell[tl]{", Student, " \\\\ \\textit{\\small (Co-advisor: ", coadvisor, ")}}"),
      Student
    )
  ) %>%
  select(-coadvisor) # Remove the standalone column so it doesn't print
kbl(grad_advised, digits = 1,
                  format = "latex",
                  align = "l",
                  col.names = NULL,
                  row.names = FALSE,
                  booktabs = T,
                  longtable = T, 
                  linesep = "",
                  escape = FALSE) %>% # CRITICAL: Allows LaTeX code inside cells to render
  kable_styling(bootstrap_options = c("hover"),
                latex_options = c("scale_down"),
                full_width = F,
                font_size = 10,
                position = "left") %>%
  sub("\\\\toprule", "", .) %>%
  sub("\\\\bottomrule", "", .)
# kbl(grad_advised,digits = 1,
#                   format = "latex",
#                   align="l",
#     col.names = NULL,
#     longtable = T,
#                   row.names = FALSE,
#                   booktabs=T,
#                   linesep = "") %>%
#   kable_styling(bootstrap_options = c("hover"),
#                 latex_options = c("scale_down"),
#                 full_width = F,
#                 font_size = 10,
#                 position = "left") %>%
#   sub("\\\\toprule", "", .) %>%
#   sub("\\\\bottomrule", "", .)
```
\vspace{-0.2cm}
_Students under my supervision have been awarded over $12,000 in grants and awards in support of their research, teaching, and outreach and been the recipients of the following fellowships- and grants to support their studies:_ 

*	JRN-LTER Graduate Research Fellowship (R Moore)
*	Ecological Society of America's Forrest Shreve Student Research Award (T Pichler)
*	Southwestern Association of Naturalists’ Howard McCarley Student Research Grant (H Whited)

<!-- _Research by my students has also been recognized for its excellence:_  -->

<!-- * Priyanka Hariharan: ATBC Navjot Sodhi Conservation Research Award (2024), UF CALS International Student Outstanding Achievement Award (2022) -->
<!-- * Thomas Smith: Best Student Presentation, IALE-NA Annual Meeting (2022) -->
<!-- *	Fabiane Mundim: UF IFAS High-Impact Publication (2021), UF WEC High-Impact Publication (2016), Honorable Mention, OTS Outstanding Paper Competition (2012) -->
<!-- *	Ana Cecilia Silva: Honorable Mention, Charles Wood Thesis Award, UF-LATAM (2015) -->
<!-- *	Christine Lucas: Marianne Schmink Outstanding Dissertation Award, UF-TCD (2012) -->
<!-- *	Ernane Vieira Neto: E. Lucy Braun Outstanding Poster Award, ESA Annual Meeting (2011) -->
<!-- *	Shannon Moore: Best Presentation, UF High School Student Science Training Program Symposium (2010) -->
<!-- *	Matthew Trager: Outstanding Research Award - MS Category, Wildlife Graduate Student Association (2005) -->

\vspace{-0.2cm}
### DOCTORAL COMMITTEE SERVICE
\vspace{-0.2cm}
```{r DocCmte, echo=FALSE,message = FALSE,warning=FALSE}
grad_advised<-read_csv("./cv_files/grad_advised.csv") %>% filter(role=="doc_cmte") %>% select(-role, -coadvisor)
kbl(grad_advised,digits = 1,
                  format = "latex",
                  align="l",
                  col.names = NULL,
                  row.names = FALSE,
                  booktabs=T,
                  longtable = T, 
                  linesep = "") %>%
  kable_styling(bootstrap_options = c("hover"),
                #latex_options = c("scale_down"),
                full_width = T, #F, #change this back when FWCE students are added
                font_size = 10,
                position = "left") %>%
  sub("\\\\toprule", "", .) %>%
  sub("\\\\bottomrule", "", .)
```
\vspace{-0.4cm}
<!-- **Wildlife Ecology & Conservation**: David Mason (in progress), Flavia Montaño Centellas (2018), Margo Stoddard (2017), Rajeev Pillay (2016), Rosalyn Johnson (2016), Gonzalo Rivas (2015), Oscar Gonzalez (2015), Dan Thornton (2010), Ivan Diaz (2008), Christa Zweig (2008), Traci Darnell (2006)  **Forest Resources & Conservation**: Andrea Irons (in progress), Demian Gomez Damiano (2020), Jacy Hyde (2019), Christie Klimas (2010)  **Biology (*formerly Depts. of Botany & Zoology*)**: Maria Beatriz Cortez (in progress), Johanna R. Jantzen (2020), Trevor Caughlin (2013), Kristine Callis (2013), Mollie Brooks (2012), Silvia Alvarez-Clare (2012), C. Seabird McKeon (2010), Joseph Veldman (2010), Jennifer Schafer (2010), Alex Jahn (2009), Silvia Lomáscolo (2008), Bonifacio Mostacedo (2007)  **Interdisciplinary Ecology**: Joanna Tucker (2010), Matt Trager	(2009)  **Entomology**: Jade Badon (2018), Pablo Allen (2017), David Serrano (2006)  **Anthropology**: Scott Macrae (2017)  **Columbia University**: Marina Cortes (Ecology, Evolution, Behavior (2012) -->
\vspace{-0.2cm}
### MASTERS COMMITTEE SERVICE
\vspace{-0.2cm}
```{r MastersCmte, echo=FALSE,message = FALSE,warning=FALSE}
grad_advised<-read_csv("./cv_files/grad_advised.csv") %>% filter(role=="masters_cmte") %>% select(-role, -coadvisor)
kbl(grad_advised,digits = 1,
                  format = "latex",
                  align="l",
        longtable = T, 
    col.names = NULL,
                  row.names = FALSE,
                  booktabs=T,
                  linesep = "") %>%
  kable_styling(bootstrap_options = c("hover"),
                latex_options = c("scale_down"),
                full_width = F,
                font_size = 10,
                position = "left") %>%
  sub("\\\\toprule", "", .) %>%
  sub("\\\\bottomrule", "", .)
```
\vspace{-0.4cm}
<!-- **Wildlife Ecology & Conservation**: Lauren Trotta (2016), Brittany Bankovich (2014), Nathan Marcy (2010), Kristen Aaltonen (2009), Mariano Rodriguez Cabal (2007)  **Forest Resources & Conservation**: Jennifer Holm (2007), Christie Klimas (2006)  **Biology (*formerly Depts. of Botany & Zoology*)**: Tania Kim (2006), Silvia Alvarez Clare (2004)  **Interdisciplinary Ecology**: Jamie Cotta (2006)  **Entomology**: Teresa Cooper (2006)  **Latin American Studies**: Crisne Lebron (2015), Carly Voight (2010), Laura Fonseca (2009), Jennifer Davies (2003), **Anthropology**: Autumn Rose (in progress). -->
\vspace{-0.2cm}
### EXTERNAL THESIS EXAMINER
<!-- ### PROPOSAL REVIEW, THESIS REVIEW, OR THESIS DEFENSE COMMITTEE -->
\vspace{-0.2cm}
```{r external, echo=FALSE,message = FALSE,warning=FALSE}
other<-read_csv("./cv_files/grad_advised.csv") %>% filter(role=="other") %>% select(-role,-coadvisor,-Program)
kbl(other,digits = 1,
                  format = "latex",
                  align="l",
        longtable = T, 
    col.names = NULL,
                  row.names = FALSE,
                  booktabs=T,
                  linesep = "") %>%
  kable_styling(bootstrap_options = c("hover"),
                latex_options = c("scale_down"),
                full_width = F,
                font_size = 10,
                position = "left") %>%
  sub("\\\\toprule", "", .) %>%
  sub("\\\\bottomrule", "", .)
```

<!-- ### POSTDOCTORAL, SABBATICAL, & POSTGRADUATE SCHOLARS  -->
<!-- \vspace{-0.2cm} -->
<!-- ```{r lab_members, echo=FALSE,message = FALSE,warning=FALSE} -->
<!-- lab_members<-read_csv("./cv_files/lab_members.csv")  -->
<!-- lab_members$cat<-as.factor(lab_members$cat) -->
<!-- lab_members$cat <- ordered(lab_members$cat, levels = c("postdoc","sab" ,"postgrad", "vis_grad","ug","hs")) -->
<!-- lab_members2<-lab_members %>%  -->
<!--   filter(cat=="postdoc"|cat=="sab"|cat=="postgrad") %>%  -->
<!--   arrange(order(cat)) %>%  -->
<!--   select(-cat) %>%  -->
<!--   drop_na() -->
<!--   # replace_na(list(Name=" ", Institution=" ", Position=" ", Year=" ")) -->
<!-- kbl(lab_members2,digits = 1, -->
<!--     format = "latex", -->
<!--     align="l", -->
<!--     longtable = T,  -->
<!--     col.names = NULL, -->
<!--     row.names = FALSE, -->
<!--     booktabs=T, -->
<!--     linesep = "") %>% -->
<!--   kable_styling(bootstrap_options = c("hover"), -->
<!--                 latex_options = c("scale_down"), -->
<!--                 full_width = F, -->
<!--                 font_size = 10, -->
<!--                 position = "left") %>% -->
<!--   sub("\\\\toprule", "", .) %>% -->
<!--   sub("\\\\bottomrule", "", .) -->
<!-- ``` -->

<!-- **Doctoral Defense Committee:** Helen Belan (U. Federal de Uberlândia, Brazil, 2020), Alan Nilo da Costa (U. Federal de Uberlândia, Brazil, 2013). **Doctoral Dissertation Examiner:** Talita Reis (UNICAMP, Brazil, 2015), Mauricio Bonesso Sampaio (UNICAMP, Brazil, 2012), Maria Rosa Darrigo (UNICAMP, Brazil, 2011), Sjirk Geerts (Stellenbosch University, South Africa, 2011), Rafael Soares de Arruda (INPA, Brazil, 2008), Thiago Junqueira Izzo (INPA, Brazil, 2005). **Doctoral Proposal Defense**: Laura Vivian Barbosa (U. Federal de Uberlândia, Brazil, 2014) **Master's Thesis Defense Committee**: Jonas B. Marvalhas (U. Federal de Uberlândia, Brazil 2013), Ernane Vieira Neto (U. Federal de Uberlândia, Brazil, 2008), Alan Nilo da Costa (U. Federal de Uberlândia, Brazil, 2007), Daniel Mascia Vieira (Universidade de Brasília, Brazil, 2002) -->

<!-- Carine Emer			INPA, Brazil				M.S. Thesis Examiner, 2011 -->
<!-- Flávio de Oliveira Rodrigues	INPA, Brazil				M.S. Thesis Examiner, 2011 -->
<!-- Waldete Castro Lourenço		INPA, Brazil				M.S. Thesis Examiner, 2009 -->
<!-- Simone B. Fontoura		INPA, Brazil				M.S. Thesis Examiner, 2007 -->
<!-- Paula M. Pedrosa			INPA, Brazil				M.S. Thesis Examiner, 2004 -->

<!-- Carine Emer 			INPA, Brazil				M.S. Proposal Examiner, 2009 -->
<!-- Simone Benedet Founatura	INPA, Brazil				M.S. Proposal Examiner, 2006 -->
<!-- Thiago Junqueira Izzo		INPA, Brazil				M.S. Proposal Examiner, 2002 -->

<!-- Melina Alcalá			UFSCar, Brazil				M.S. Qualifying Exam, 2009 -->
<!-- Raquel Negrão Baldoni		UFSCar, Brazil				M.S. Qualifying Exam, 2009 -->

<!-- Pedro B. Vasconcelos		U. Federal de Uberlândia, Brazil		B.S. Thesis Defense, 2009 -->
<!-- Fabiane M. Mundim 		U. Federal de Uberlândia, Brazil		B.S. Thesis Defense, 2007 -->
<!-- ### UNDERGRADUATE THESIS STUDENTS, HIGH SCHOOL RESEARCHERS, & VISITING GRADUATE STUDENTS  -->
<!-- \vspace{-0.2cm} -->
<!-- ```{r lab_members2, echo=FALSE,message = FALSE,warning=FALSE} -->
<!-- lab_members<-read_csv("./cv_files/lab_members.csv")  -->
<!-- lab_members$cat<-as.factor(lab_members$cat) -->
<!-- lab_members$cat <- ordered(lab_members$cat, levels = c("postdoc","sab" ,"postgrad", "vis_grad","ug","hs")) -->
<!-- lab_members3<-lab_members %>%  -->
<!--   filter(cat=="ug"|cat=="hs"|cat=="vis_grad") %>%  -->
<!--   arrange(order(cat)) %>%  -->
<!--   select(-cat) %>%  -->
<!--   drop_na() -->
<!--   # replace_na(list(Name=" ", Institution=" ", Position=" ", Year=" ")) -->
<!-- kbl(lab_members3,digits = 1, -->
<!--                   format = "latex", -->
<!--                   align="l", -->
<!--         longtable = T,  -->
<!--     col.names = NULL, -->
<!--                   row.names = FALSE, -->
<!--                   booktabs=T, -->
<!--                   linesep = "") %>% -->
<!--   kable_styling(bootstrap_options = c("hover"), -->
<!--                 latex_options = c("scale_down"), -->
<!--                 full_width = F, -->
<!--                 font_size = 10, -->
<!--                 position = "left") %>% -->
<!--   sub("\\\\toprule", "", .) %>% -->
<!--   sub("\\\\bottomrule", "", .) -->
<!-- ``` -->

<!-- \begingroup  -->
<!-- \setlength{\leftskip}{.2cm} -->

<!-- *__Field Research Interns:__* _2007_: Wesley Dáttilo da Cruz, Univ. Estadual do Norte Fluminense (Brazil), Jefferson José Valsko da Silva, Universidade do Norte Fluminense (Brazil). _2005-2007_: Renata Pacheco do Nacimento and Cauê Thomé Lopes, Univ. Federal de Uberlândia (Brazil). _2005-2006_: Polyanna Custodio Duarte, Univ. Federal de Uberlândia (Brazil). _2006_: Waldete Castro Lourenço, Univ. Federal do Amazonas (Brazil), Elisabete Marques da Costa, Univ. Federal de Pernambuco (Brazil). _2005_: Bruno Turbiani, Universidade Estadual Paulista (Brazil). _2004_: Cristina Escate, Universidade Federal do Amazonas (Brazil), Simone Benedet, Universidade do Vale do Rio do Sinos (Brazil). _2003_: Cris Follman Jurinitz, Univ. Federal do Rio Grande do Sul (Brazil), Denise Cruz, Univ. Federal do Rio de Janeiro (Brazil). _2003-2004_: David M. Lapola, Universidade Estadual Paulista (Brazil). _2002_: Maria Beatriz Nogueira, Universidade de São Paulo (Brazil). _2001-2002_: Sylvia Heredia, Universidad de los Andes (Colombia). _2001_: Obed Garcia, INPA (Brazil), Olavo Nardy, UNESP (Brazil). -->

<!-- \endgroup -->


<!-- text size: https://www.overleaf.com/learn/latex/Font_sizes%2C_families%2C_and_styles -->
<!-- xrfill adjusts line position and width -->
<!-- \vspace{0.2cm} -->
\vspace{-0.2cm}
\textbf{\textcolor{darkmidnightblue}{\large{SERVICE TO THE UNIVERSITY \& PROFESSIONAL ORGANIZATIONS }}}\xrfill[0.5ex]{1pt}

<!-- ## INSTITUIONAL AND PROFESSIONAL SERVICE -->
\vspace{-0.4cm}
### NEW MEXICO STATE UNIVERSITY  
\vspace{-0.2cm}
\begingroup 
\setlength{\leftskip}{.2cm}
Institutional Animal Care and Use Committee (2023-present)  

\endgroup 

\vspace{-0.2cm}
### NMSU COLLEGE OF AGRICULTURAL, CONSUMER AND ENVIRONMENTAL SCIENCES
\vspace{-0.2cm}
\begingroup 
\setlength{\leftskip}{.2cm}
Connections & Engagement Initiative Committee (2025-present) 

\endgroup 

<!-- \vspace{-0.2cm} -->
<!-- ### UF TENURE MENTORSHIP & PEER TEACHING EVALUATION COMMITTEES -->
<!-- \vspace{-0.2cm} -->
<!-- \begingroup  -->
<!-- \setlength{\leftskip}{.2cm} -->
<!-- Tenure Mentorship Committee, Dr. Nia Morales (Wildlife Ecology & Conservation, 2021-2025)   -->
<!-- Tenure Mentorship Committee, Dr. Matt Hallet (Wildlife Ecology & Conservation, 2022-2025)   -->
<!-- Tenure Mentorship Committee, Dr. Luke Flory (Agronomy, 2012)   -->
<!-- Chair, Tenure Mentorship Committee, Dr. Robert Fletcher (Wildlife Ecology & Conservation, 2009-2012)   -->
<!-- Chair, Peer Teaching Evaluation Committee, Dr. Ben Baiser (Wildlife Ecology & Conservation, 2025)   -->
<!-- Peer Teaching Evaluation Committee, Dr. Raelene Crandall (Forest Fisheries, & Geomatic Sciences, 2021-2022)   -->
<!-- Chair, Peer Teaching Evaluation Committee, Dr. Ethan White (Wildlife Ecology & Conservation, 2019-2020)   -->
<!-- Chair, Peer Teaching Evaluation Committee, Dr. Morgan Ernest (Wildlife Ecology & Conservation, 2017-2018)   -->
<!-- Peer Teaching Evaluation Committee, Dr. Bill Pine (Wildlife Ecology & Conservation, 2015)   -->
<!-- Chair, Peer Teaching Evaluation Committee, Dr. Lyn Branch (Wildlife Ecology & Conservation, 2011)   -->
<!-- Chair, Peer Teaching Evaluation Committee, Dr. Bill Pine (Wildlife Ecology & Conservation, 2010)   -->
<!-- Chair, Peer Teaching Evaluation Committee, Dr. Robert Fletcher (Wildlife Ecology & Conservation, 2009)   -->
<!-- Peer Teaching Evaluation for Dr. Michelle Mack (Botany, 2006)     -->
<!-- Peer Teaching Evaluation Committee, Dr. Madan Oli (Wildlife Ecology & Conservation, 2003)   -->

<!-- \endgroup  -->

\vspace{-0.2cm}
### NMSU DEPARTMENT OF FISH, WILDLIFE AND CONSERVATION ECOLOGY
\vspace{-0.2cm}
\begingroup 
\setlength{\leftskip}{.2cm}
Undergraduate Curriculum Committee (2024-present)  
Search Committee, Chair of the Department of Fish, Wildlife and Conservation Ecology (2024)  
Diversity, Equity, and Inclusion Committee (2023-2024)  

\endgroup 

<!-- \vspace{-0.2cm} -->

<!-- ### UF CENTER FOR LATIN AMERICAN STUDIES -->
<!-- \vspace{-0.2cm} -->
<!-- \begingroup  -->
<!-- \setlength{\leftskip}{.2cm} -->
<!-- Search Committee, TCD Program Director (2026)     -->
<!-- Graduate Admissions Committee (2026)     -->
<!-- Search & Screen Committee, Asst. Professor of Environmental Justice (2023)   -->
<!-- Post-Tenure Review Committee (2023)   -->
<!-- LAS Faculty Advisory Council (2020-2023)   -->
<!-- LAS Bylaws Revision Committee (2020-2021)   -->
<!-- Chair, LAS Awards & Scholarship Committee (2019)   -->
<!-- LAS Merit Evaluation Committee (2015-2017)   -->
<!-- Chair, Search & Screen Committee, Lecturer in Latin American Studies (2015)   -->
<!-- Search & Screen Committee, LAS Communications Coordinator (2015)   -->
<!-- Bacardi Seminar Series Coordinator (2015)   -->
<!-- Curriculum Committee, Masters Program in Sustainable Development Practice (2010-2012)   -->
<!-- LAS Graduate Admissions & Fellowship Committee (2007)   -->

<!-- \endgroup  -->

<!-- \vspace{-0.2cm} -->
<!-- ### UF TROPICAL CONSERVATION & DEVELOPMENT PROGRAM -->
<!-- \vspace{-0.2cm} -->
<!-- \begingroup  -->
<!-- \setlength{\leftskip}{.2cm} -->
<!-- Chair, Field Research Grants Committee (2026, 2024, 2020-2021, 2015, 2008-2010, 2004)   -->
<!-- Academic Advisor (2021-2022, 2008-11)   -->
<!-- Steering Committee (2018-2024,2010-16, 2002-08)   -->
<!-- Chair, Marianne Schmink Outstanding Dissertation Award Selection Committee (2022, 2013, 2011, 2010)   -->
<!-- Chair, Fellowship Committee (2022, 2012)   -->
<!-- Advisory Committee, Pan-Amazon Community of Practice Project (2018-2020)   -->
<!-- Faculty Advisor, TCD Graduate Student Organization (2008-2016)   -->
<!-- Organizing Committee, TCD Strategic Planning (2007-2008)   -->
<!-- Advisory Council, Amazon Conservation Leadership Initiative (2004-2005)   -->
<!-- Fellowship Committee (2003-2005)   -->
<!-- Field Research Grants Committee (2003-2004)   -->

<!-- \endgroup   -->

<!-- \vspace{-0.2cm} -->
<!-- ### ASSOCIATION FOR TROPICAL BIOLOGY & CONSERVATION  -->
<!-- \vspace{-0.2cm} -->
<!-- \begingroup  -->
<!-- \setlength{\leftskip}{.2cm}   -->

<!-- Search Committee, ATBC Executive Director (2024-2025)   -->
<!-- Search Committee, _Biotropica_ Editor-in-Chief (2024-2025)   -->
<!-- Past-President (2022)    -->
<!-- Chair, Nominations Committee (2022)   -->
<!-- President (2021)   -->
<!-- Annual Meeting Program & Organizing Committees (2021)     -->
<!-- President-Elect (2020)   -->
<!-- Student Field Grants Selection Committee (2020)   -->
<!-- Crowdfunding Workshop, Annual Meeting (2018)   -->
<!-- Annual Meeting Scientific Committee (2017, 2012)   -->
<!-- Councillor (2009-2011)   -->
<!-- Conservation Committee (2004-2006)   -->
<!-- Annual Meeting Symposium Co-organizer (2004)   -->

<!-- \endgroup  -->

<!-- \vspace{-0.2cm} -->
<!-- ### ECOLOGICAL SOCIETY OF AMERICA    -->
<!-- \vspace{-0.2cm} -->
<!-- \begingroup  -->
<!-- \setlength{\leftskip}{.2cm}   -->

<!-- Secretary (2022-2025)   -->
<!-- Governing Board (2022-2025)   -->
<!-- Teller's Committee (2022-2025)     -->
<!-- Governance & Leadership Committee (2022-2025)    -->
<!-- Historical Records Committee (2022-2025)    -->

<!-- \endgroup  -->
\vspace{-0.2cm}
### JORNADA BASIN LTER 
\vspace{-0.2cm}
\begingroup 
\setlength{\leftskip}{.2cm}
Information Management Committee (2025-present)

\endgroup


\vspace{-0.2cm}
### OTHER UNIVERSITIES & PROFESSIONAL ORGANIZATIONS   
\vspace{-0.2cm}
\begingroup 
\setlength{\leftskip}{.2cm}
State Wildlife Action Plan (SWAP) for New Mexico Core Team, New Mexico Department of Wildlife (2025-present)  
<!-- External Tenure & Promotion Evaluations (N=26, 2011-present)   -->
<!-- 2024 (3) 2023 (3), 2022 (4), 2021 (2), 2020 (3), 2018 (2), 2017 (1), 2016 (3), 2015 (1), 2011 (1)   -->
Co-coordinator of Southern African bat assessments for the IUCN Red List (2023-present)  
GBatNet Prioritizing Key Habitats and Species Conservation Working Group (2022-present)  
NABat Emergence/Summer Colony Count Working Group (2020-2024)  
Lesser Long-nosed Bat Post-Delisting Monitoring Plan Participant, U.S. Fish and Wildlife Service (2023)  
Diversity, Equity, & Inclusion Committee, Ecology and Evolution Department, University of Chicago (2021-2022)  
Planning Committee Member, Northern Colorado & Wyoming chapter of Expanding Your Horizons—an annual STEM workshop for 6th through 8th grade girls (2017-2019)  
Dr. David Otis Speaker Series Selection Committee, Colorado State University (2017-2018)  
Panel member, Colorado State University Info Session on the NSF Graduate Research Fellowship Program (2017)  
Graduate Representative, Population Ecology Assistant Professor Search Committee, University of Montana (2013-2014)  
Graduate Representative, Department of Ecosystem & Conservation Sciences Student Evaluation Committee, University of Montana (2014)  


\endgroup 

\vspace{0.2cm}
\textbf{\textcolor{darkmidnightblue}{\large{SERVICE AS A REVIEWER}}}\xrfill[0.5ex]{1pt}
<!-- \textbf{\textcolor{darkmidnightblue}{\large{EDITORIAL POSITIONS \& SERVICE AS A REVIEWER}}}\xrfill[0.5ex]{1pt} -->

\begingroup 
\setlength{\leftskip}{.2cm}

<!-- **Editor-in-Chief**: _Biotropica_ (2013-2019; _>2400 decisions_)   -->

<!-- **Associate Editor:** _Biotropica_ (2010-2013)   -->

<!-- **Subject Editor:** _PLOS Ecosystems_ (2025-Present), _Biotropica_ (2007-2019), _Ecology_ (2008-2012), _Ecological Monographs_ (2008-2012), _Ecological Applications_ (2008-2012). -->

__*Ad hoc* journal referee (No. of reviews):__ *Acta Chiropterologica* (5), *Biological Conservation* (2), *Biotropica* (2), *EcoHealth* (1), *Global Change Biology* (1), *Global Ecology and Biogeography* (3), *Herpetological Monographs* (1), *Human Dimensions of Wildlife* (7), *Journal of Bat Research and Conservation* (1), *Journal of Mammalogy* (5), *Mammal Review* (1), *PLOS ONE* (1), *Proceedings of the Royal Society B: Biological Sciences* (2), *Scientific Reports* (1), *Sustainability* (2), *Urban Ecosystems* (2)

<!-- __US National Science Foundation:__ Committee of Visitors, NSF Office of International Science & Engineering (2023), Proposal Review Panels: DEB Ecology (2005), DEB Population & Evolutionary Processes (2007), DEB Population & Community Ecology (2011, 2013). Ad-hoc proposal review: Division of Environmental Biology Core Programs (9), UMEB Program (2), RUI Program (1), AMERICAS Dissertation Improvement Grants (1), NSF Global Scientists & Engineers Program (1), International Research Fellowship Program (1),  Dimensions of Biodiversity Program (1) -->

__Other reviewer service__: Bat Conservation International (8), National Geographic Society (17), The Red List of Mammals of South Africa, Lesotho and Swaziland (1, species status assessment)  

\endgroup 


<!-- text size: https://www.overleaf.com/learn/latex/Font_sizes%2C_families%2C_and_styles -->
<!-- xrfill adjusts line position and width -->
\vspace{0.2cm}
\textbf{\textcolor{darkmidnightblue}{\large{SEMINARS, OUTREACH, \& MEDIA APPEARANCES }}}\xrfill[0.5ex]{1pt}
\vspace{-0.4cm}
<!-- ## SEMINARS & OUTREACH ACTIVITIES -->

### INVITED SEMINARS
\vspace{-0.2cm}
```{r talks, echo=FALSE,message = FALSE,warning=FALSE}
talks<-read_csv("./cv_files/talks.csv", col_types = cols(.default = "c")) %>% 
  select(-year)  %>% 
  replace_na(list(year2=""))
kbl(talks, 
      format    = "latex", 
      align = "l",
      longtable = T, 
      row.names = FALSE,
      col.names = NULL,
      booktabs=T,
      linesep = "") %>%
  kable_styling(bootstrap_options = c("hover"),
                latex_options = c("scale_down"),
                full_width = F,
                font_size = 10,
                position = "left") %>%
  sub("\\\\toprule", "", .) %>%
  sub("\\\\bottomrule", "", .)
```
\vspace{-0.4cm}
\vspace{-0.2cm}
### INTERVIEWS & MEDIA APPEARANCES
\vspace{-0.2cm}
\begingroup 
\setlength{\leftskip}{.2cm}
2025\hspace{4mm}  [KRQE News 13-Albuquerque](https://www.youtube.com/watch?v=tSvK9wcM-D4)   
2021\hspace{4mm}  Roan NEWS Magazine, Namibian Environment & Wildlife Society  
2020\hspace{4mm}  [Nature Research Ecology & Evolution Community](https://natureecoevocommunity.nature.com/posts/catalyzing-dialogue-about-who-should-be-advancing-equity-diversity-and-inclusion-on-college-campuses-and-beyond)  
2019\hspace{4mm}  [Nature Research Ecology & Evolution Community](https://natureecoevocommunity.nature.com/channels/521-behind-the-paper/posts/49514-how-can-we-achieve-diversity-and-inclusion-in-ecology)  
2018\hspace{4mm}  [Rocky Mountain Public Broadcasting Station](https://www.youtube.com/watch?v=wg_Weynsgf0), [The Conservation](https://theconversation.com/could-bats-guide-humans-to-clean-drinking-water-in-places-where-its-scarce-90311)   
2016\hspace{4mm}  National Geographic Society
<!-- [Vox](https://www.vox.com/policy-and-politics/23593369/ron-desantis-florida-schools-higher-education-woke)    -->
<!-- 2022\hspace{4mm}  [NRC (The Netherlands)](https://www.nrc.nl/nieuws/2022/02/22/systeem-van-open-access-helpt-wetenschappers-uit-arme-landen-niet-a4092976), [Nature](https://www.nature.com/articles/d41586-022-00342-w), WIRED Magazine, [Fox News Business](https://www.foxbusiness.com/politics/university-professors-unhappy-desantis-accountability), [The Scientist](https://www.the-scientist.com/news-opinion/florida-faculty-await-details-on-new-tenure-law-69971),   -->
<!-- &nbsp;\hspace{10mm} [Miami Herald](https://www.miamiherald.com/news/local/education/article263966316.html), [The Independent Florida Alligator](https://www.alligator.org/article/2022/09/blue-and-gold-macaw-uf-campus), [Chronicle for Higher Education](https://www.chronicle.com/article/it-could-have-been-worse-higher-ed-reacts-to-ben-sasse-at-u-of-florida),   -->
<!-- &nbsp;\hspace{10mm} [The Hill](https://thehill.com/changing-america/enrichment/education/3274661-texas-and-florida-take-steps-to-limit-professor-tenure-at-state-schools/), Voice of America, ['How Ecology Works' Podcast](https://www.chemistryworld.com/news/most-scientific-editors-still-come-from-us-or-uk/3008426.article)   -->
<!-- 2021\hspace{4mm} Chronicle for Higher Education, [The Florida Independent Alligator](https://www.alligator.org/article/2021/01/what-happens-if-my-classmate-has-covid-19), [Associated Press](https://apnews.com/article/brazil-rio-de-janeiro-trees-abf6a15a804ea1485fbcec0eeed3d8d7), ABC's Nightline   -->
<!-- 2020\hspace{4mm} [Washington Post](https://www.washingtonpost.com/nation/2020/11/11/kyle-lamb-data-florida-coronavirus-desantis/), [South Florida Sun Sentinel](https://www.sun-sentinel.com/opinion/commentary/fl-op-com-grimm-ohio-sports-blogger-florida-covid-20201113-wikb2iv3gbd2vcby4hcevwyd6i-story.html)    -->
<!-- 2019\hspace{4mm} [USA Today](https://www.usatoday.com/story/news/nation/2019/10/18/amazon-rainforest-still-burning-more-fires-future/4011238002/), CBC News Network, [The Atlantic](https://www.theatlantic.com/science/archive/2019/04/amnh-under-fire-hosting-brazils-jair-bolsonaro/587068/), [The Verge](https://www.theverge.com/2019/9/10/20858834/amazon-fire-brazil-forest-harm-recovery-century), MSNBCs Velshi & Ruhle   -->
<!-- &nbsp;\hspace{10mm} [WIRED Magazine](https://www.wired.com/story/the-horrifying-science-of-the-deforestation-fueling-amazon-fires/), [Mashable](https://mashable.com/article/amazon-rainforest-fires-tipping-point-deforestation), BBC World Service, [BBC News](https://www.bbc.com/mundo/noticias-america-latina-49504502), The Miradas Podcast, [Phys.org](https://phys.org/news/2019-10-amazon-hasnt-outbreaks-month-future.html), [Infobae](https://www.infobae.com/america/america-latina/2019/09/10/las-areas-quemadas-del-amazonas-podrian-tardar-siglos-en-recuperarse-por-completo/)     -->
<!-- 2018\hspace{4mm} Popular Science, Associated Press, [National Geographic](https://www.nationalgeographic.com/science/article/news-museu-nacional-fire-rio-de-janeiro-natural-history), Scholastic Classroom Magazines,   -->
<!-- &nbsp;\hspace{10mm} [Minnesota Public Radio's Climate Cast](https://www.mprnews.org/story/2018/11/15/why-brazils-new-president-may-bring-more-climate-disaster-death), Gainesville Sun, Mongabay.com, [NBC News](https://www.nbcnews.com/news/latino/brazil-s-president-elect-jair-bolsonaro-may-smother-rainforest-scientists-n940326), [La Repubblica](https://www.repubblica.it/ambiente/2018/11/27/news/amazzonia_un_milione_di_campi_di_calcio_perduti_in_un_anno-212784490/),   -->
<!-- &nbsp;\hspace{10mm} [Editage Insights](https://www.editage.com/insights/science-journals-might-be-more-international-but-not-their-editorial-boards)      -->
<!-- 2017\hspace{4mm} [Nature](https://www.nature.com/articles/548271a), [The Wire (India)](https://thewire.in/science/science-journals-might-international-not-editorial-boards), [Chemistry World](https://www.chemistryworld.com/news/most-scientific-editors-still-come-from-us-or-uk/3008426.article), Gainesville Sun, [Indianapolis  Star](https://www.indystar.com/story/opinion/columnists/suzette-hackney/2017/03/24/hackney-robotics-kids-kindness-triumphs-hate/99578116/), [Phys.org](https://phys.org/news/2017-12-science-international-scientific-editorial-boards.html)    -->
<!-- 2016\hspace{4mm} [_Nature_](https://www.nature.com/articles/nature.2016.20830), [Chronicle of Higher Education](https://www.chronicle.com/article/in-donald-trump-phd-scientists-find-catharsis-amid-gags/), [DailyDot](https://www.dailydot.com/unclick/donald-trump-phd-twitter-makes-science-great-again/), [Mongabay.com](https://news.mongabay.com/2016/01/special-issue-of-biotropica-makes-the-case-for-protecting-south-americas-white-sand-forests/) [Mongabay.com](https://news.mongabay.com/2016/04/building-roads-agricultural-expansion-brazil-aiding-spread-insects-harm-crops/)   -->
<!-- 2015\hspace{4mm} [WAMC Public Radio](https://www.wamc.org/2015-01-30/is-there-a-pubradiovoice-that-sounds-like-america)   -->
<!-- 2014\hspace{4mm} NPRs Morning Edition - "Joe's Big Idea", TV Acre - Rede Globo Rio Branco, Bom Dia Amazônia, [SciDevNet](https://www.scidev.net/global/news/armadillo-mascot-world-cup/)   -->
<!-- &nbsp;\hspace{10mm} TV Gazetta 'Entrevista com Rogério Wenceslau' (Rede Record Rio Branco), Estado de São Paulo   -->
<!-- &nbsp;\hspace{10mm} Nature.com, WUFT, Science/Science Insider, Howler Soccer Magazine's Dummy Podcast   -->
<!-- 2013\hspace{4mm} SciDevNet, WSLR Florida-Caribe Radio Show   -->
<!-- 2006\hspace{4mm} The Scientist Magazine, Science Minority Scientists Network   -->
<!-- 2005\hspace{4mm} Science's Next Wave, Televisão Universitária, Uberlândia, Folha de São Paulo   -->
<!-- 2004\hspace{4mm} _Science_ Magazine's Minority Scientists Network   -->
<!-- 2002\hspace{4mm} NPR's Weekend Edition Saturday, BBC World Service Radio 4, Austrian Public Television's Modern Times.   -->

\endgroup   

<!-- text size: https://www.overleaf.com/learn/latex/Font_sizes%2C_families%2C_and_styles -->
<!-- xrfill adjusts line position and width -->
\vspace{0.2cm}
\textbf{\textcolor{darkmidnightblue}{\large{PROFESSIONAL DEVELOPMENT }}}\xrfill[0.5ex]{1pt}

<!-- ## PROFESSIONAL DEVELOPMENT ACTIVITIES   -->
\vspace{-0.2cm}
### TRAINING PROGRAMS
\begingroup 
\setlength{\leftskip}{.2cm}  
2013\hspace{4mm} University of Florida Science Communication Academy  
2004\hspace{4mm} California Academy of Sciences: The Ant Course: Taxonomy & Systematics (Costa Rica)  
1996\hspace{4mm} Organization for Tropical Studies: Tropical Plant Systematics 96-9 (Costa Rica)  
1993\hspace{4mm} University of California Education Abroad Program: Tropical Biology Quarter (Costa Rica) 

\endgroup

\vspace{-0.2cm}

### MEMBERSHIPS IN PROFESSIONAL SOCIETIES  

\begingroup 
\setlength{\leftskip}{.2cm}  
North American Society for Bat Research, Society for Conservation Biology, American Society of Mammalogists, Ecological Society of America, Association for Tropical Biology and Conservation, Global South Bats, Western Bat Working Group, The Wildlife Society 


\endgroup

<!-- text size: https://www.overleaf.com/learn/latex/Font_sizes%2C_families%2C_and_styles -->
<!-- xrfill adjusts line position and width -->
\vspace{0.2cm}
\textbf{\textcolor{darkmidnightblue}{\large{LANGUAGES }}}\xrfill[0.5ex]{1pt}  

\begingroup 
\setlength{\leftskip}{.2cm}  

Fluent in English & Spanish  

\endgroup
